### 2019-01-10 agujournaltemplate.tex
Initial ShareLaTeX Import

%% March 2018
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def, etc.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
% DO NOT USE \psfrag or \subfigure commands.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
% There are two options for article format:
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%

%% To submit your paper:
\documentclass[draft,linenumbers]{agujournal2018}
\usepackage{apacite}
\usepackage{url} %this package should fix any errors with URLs in refs.

%%%%%%%
% \usepackage{trackchanges}
% uncomment the line above to use the TrackChanges package to mark revisions if needed.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

% Now, type in the journal name: \journalname{<Journal Name>}

% ie, \journalname{Journal of Geophysical Research}
%% Choose from this list of Journals:
%
% JGR-Atmospheres
% JGR-Biogeosciences
% JGR-Earth Surface
% JGR-Oceans
% JGR-Planets
% JGR-Solid Earth
% JGR-Space Physics
% Global Biochemical Cycles
% Geophysical Research Letters
% Paleoceanography
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resource Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%

\journalname{Space Weather}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{=enter title here=}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{=list all authors here=}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

\affiliation{=number=}{=Affiliation Address=}
%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{=name=}{=email address=}

%% Keypoints, final entry on title page.

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 100 characters or less with no special characters or punctuation
% \end{keypoints}

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 100 characters or less with no special characters or punctuation

\begin{keypoints}
\item = enter point 1 here =
\item = enter point 2 here =
\item = enter point 3 here =
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
% A good abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a direct measurement or indirect estimate (via a magnetotelluric transfer function) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth’s surface is that the system is resistive. That is, the approximation GIC(t) = $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters at frequencies associated with significant amplitudes of the geoelectric field.  We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10 kilometers away.  We compute frequency dependent network parameters  $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega) = A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are not frequency independent and that this model provides a significantly better representation of measured GIC.
\end{abstract}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
%
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
%
% \section{Data} (Or section title might be a descriptive heading about data)
%
% \section{Results} (Or section title might be a descriptive heading about the
% results)
%
% \section{Conclusions}

\section{Introduction}

Geomagnetically induced currents (GICs) are electric currents in man-made conducting systems due to electric fields induced near Earth’s surface.  These fields are a result of time variations in ionospheric currents on timescales on the order of hours or less and the movement of Earth’s surface relative to slow-varying current systems in Earth’s ionosphere that are near stationary relative to the Sun [REF].  Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure [REF].  Accurate estimation of GICs given E or B is important for power system design [REF] retrospective analysis [REF] and for the mitigation of the impacts of space weather on power systems.

A common assumption made in estimates of GIC using either measured or estimated values of E is that the power system is resistive or quasi-DC so that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.  The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods [e.g., Boteler, 2014a; Boteler and Pirjola, 2014b]. This assumption has been used or stated in many works in the past decade, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012}.

In this work, we use a unique dataset in which measurements of the geoelectric field and geomagnetic field were made near a site where GICs were being measured in a power system.  We first estimate the system parameters $a_o$ and $b_o$ using conventional methods, which assume that they are frequency independent.  Then, we use a model in which the system parameters are frequency dependent, compute the frequency dependence, and compare this model with the traditional resistive model.

\section{Data}

The 1-second-cadence GIC data used in this paper span cover a total of 36 days across the 10 intervals: 2006-04-02 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01 -- 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22 that have been presented previously in \citet{Watari2009} and \cite{Watari2015}.

One-second-cadence geoelectric field and geomagnetic field data were obtained from the Kakioka Magnetic Observatory of the Japan Meteorological Agency data portal [http://www.kakioka-jma.go.jp/obsdata/metadata/en]. Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure 1.

The GIC dataset includes 1-second raw data and a 1 Hz low-pass filtered raw measurements.  The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure 1.

The GIC data contains non-physical spikes that are followed by a relaxation time of approximately 60 seconds. These spikes were removed by identifying times ts when the change in $|GIC|$ was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ was replaced with a linear interpolation of the values from $GIC(t_s-3)$ to $GIC(t_s + 101)$.

Spikes in the electric field measurements were made using a similar procedure - when the absolute value of a component i of the electric field changed by more than 1 mV/km at time ts, the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced with a linear interpolation of the values from $E_i(t_s-3)$ to $E_i(t_s + 5)$.

The only modifications made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values. The values for all components were the same and much larger than the surrounding values.

The subscripts $x$, $y$, and $z$ correspond to North, East, and Downward, respectively.

%Text here ===>>>

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}[h]
% \centering
% when using pdflatex, use pdf file:
% \includegraphics[width=20pc]{figsamp.pdf}
%
% when using dvips, use .eps file:
% \includegraphics[width=20pc]{figsamp.eps}
%
% \caption{Short caption}
% \label{figone}
%  \end{figure}
%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% SIDEWAYS FIGURE and TABLE
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
%
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$}
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% >>>>	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% 	All funding sources related to this work from all authors
%
% 	Any real or perceived financial conflicts of interests for any
%	author
%
% 	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
%
% It is also the appropriate place to thank colleagues and other contributors.
% AGU does not normally allow dedications.

\acknowledgments
 = enter acknowledgments here =

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BibTeX is preferred:
%
\bibliography{agusample.bib}
%
% no need to specify bibliographystyle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
%% Example \citet and \citep:
%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
%  \citet{Ghel00}, and \citet{Leit74}.

%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
%  \citep{Ghel00, Leit74}.

\end{document}

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

### 2019-01-10 agusuppinfosample.tex
Initial ShareLaTeX Import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supporting Information
%% (Optional)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OVERVIEW
%
% Please note that all supporting information will be peer reviewed with your manuscript.
% In general, the purpose of the supporting information is to enable
% authors to provide and archive auxiliary information such as data
% tables, method information, figures, video, or computer software,
% in digital formats so that other scientists can use it.

% The key criteria are that the data:
% 1. supplement the main scientific conclusions of the paper but are not essential to the conclusions (with the exception of
%    including data so the experiment can be reproducible);
% 2. are likely to be usable or used by other scientists working in the field;
% 3. are described with sufficient precision that other scientists can understand them, and
% 4. are not exe files.
%

% All Supporting text and figures should be included in this document.

% Data sets, large tables, movie files,
% and audio files should be uploaded separately, following AGU naming
% conventions. Include their captions in this document and list the
% file name with the caption. You will be prompted to upload these
% files on the Upload Files tab during the submission process, using
% file type “Supporting Information (SI)”

\documentclass{agujournal2018}

% Please type in the journal name: \journalname{<Journal Name>}
% ie,
\journalname{<enter journal name here>}

%% Choose from this list of Journals:
%
% Journal of Geophysical Research
% JGR-Biogeosciences
% JGR-Earth Surface
% JGR-Planets
% JGR-Solid Earth
% JGR-Space Physics
% Global Biochemical Cycles
% Geophysical Research Letters
% Paleoceanography
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resource Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science

\begin{document}

%% This command needs article title as argument to \supportinginfo{}:
\supportinginfo{Trends in the predictive performance of raw ensemble weather forecasts}

 \authors{A. B. Smith\affil{1}\thanks{Current address, McMurdo Station,
 Antartica},
 Eric Brown\affil{1,2}, Rick Williams\affil{3},
 John B. McDougall\affil{4}, and S. Visconti\affil{5}\thanks{Also
 funded by Monsanto.}}

\affiliation{1}{Department of Hydrology and Water Resources,
University of Arizona, Tucson, Arizona, USA.}
\affiliation{2}{Department of Geography, Ohio State University,
Columbus, Ohio, USA.}
\affiliation{3}{Department of Space Sciences, University of
Michigan, Ann Arbor, Michigan, USA.}
\affiliation{4}{Division of Hydrologic Sciences, Desert Research
Institute, Reno, Nevada, USA.}
\affiliation{5}{Dipartimento di Idraulica, Trasporti ed
Infrastrutture Civili, Politecnico di Torino, Turin, Italy.}

%% Corresponding Author
%(include name and email addresses of the corresponding author.  More
%than one corresponding author is allowed in this Word file and for
%publication; but only one corresponding author is allowed in our
%editorial system.)

\correspondingauthor{A. B. Smith}{email@address.edu}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

\section*{Contents}
%%%Remove or add items as needed%%%
\begin{enumerate}
\item Text S1 to Sx
\item Figures S1 to Sx
\item Tables S1 to Sx
\end{enumerate}

\section*{Additional Supporting Information (Files uploaded separately)}

\begin{enumerate}
\item Captions for Datasets S1 to Sx
\item Captions for large Tables S1 to Sx (if larger than 1 page, upload as separate excel file)
\item Captions for Movies S1 to Sx
\item Captions for Audio S1 to Sx
\end{enumerate}

\section*{Introduction}

%Delete all unused file types below. 
%Copy/paste for multiples of each file type as needed.

\section*{Text S1.}

Type or paste text here. This should be additional explanatory text,
such as: extended descriptions of results, full details of models,
extended lists of acknowledgements etc.  It should not be
additional discussion, analysis, interpretation or critique. It
should not be an additional scientific experiment or paper.

Repeat for any additional Supporting Text

%%Enter Data Set, Movie, and Audio captions here
%%EXAMPLE CAPTIONS

\section*{Data Set S1.} %Type or paste caption here.

Upload your dataset(s) to AGU's journal submission site and select
"Supporting Information (SI)" as the file type. Following naming
convention: ds01.

Repeat for any additional Supporting data sets

\section*{Movie S1.}

Type or paste caption here.

Upload your movie(s) to AGU's journal submission site and select,
"Supporting Information (SI)" as the file type. Following naming convention: ms01.

Repeat any additional Supporting movies

\section*{Audio S1.}

Type or paste caption here.

Upload your audio file(s) to AGU's journal submission site and select
Supporting Information ``(SI)" as the file type. Following naming
convention: auds01.

Repeat for any additional Supporting audio files

\end{document}

### 2019-01-11 agujournaltemplate.tex
Updates from ShareLaTeX

@@
-%% March 2018
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-%
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements, including customized APA reference formatting.
-%
-% You may copy this file and give it your
-% article name, and enter your text.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PLEASE DO NOT USE YOUR OWN MACROS
-% DO NOT USE \newcommand, \renewcommand, or \def, etc.
-%
-% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
-% DO NOT USE \psfrag or \subfigure commands.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Step 1: Set the \documentclass
-%
-% There are two options for article format:
-%
-% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
-% The draft option produces double spaced output.
-%
-
-%% To submit your paper:
 \documentclass[draft,linenumbers]{agujournal2018}
@@
 
-% Now, type in the journal name: \journalname{<Journal Name>}
-
-% ie, \journalname{Journal of Geophysical Research}
-%% Choose from this list of Journals:
-%
-% JGR-Atmospheres
-% JGR-Biogeosciences
-% JGR-Earth Surface
-% JGR-Oceans
-% JGR-Planets
-% JGR-Solid Earth
-% JGR-Space Physics
-% Global Biochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resource Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-
 \journalname{Space Weather}
@@
 \begin{document}
-
-%% ------------------------------------------------------------------------ %%
-%  Title
-%
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
-
-% Example: \title{This is a test title}
-
-\title{=enter title here=}
+\title{A Test of the Frequency Independence Assumption of the System Parameters used in Geomagnetically Induced Current Calculations}
 
@@
 
-\authors{=list all authors here=}
+\authors{R.S. Weigel\affil{1} and P.J. Cilliers\affil{2}}
 
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
+\affiliation{1}{Space Weather Lab, George Mason University}
+\affiliation{2}{South African National Space Agency}
 
-\affiliation{=number=}{=Affiliation Address=}
+\affiliation{1}{4400 University Drive, Fairfax VA 22030}
+\affiliation{2}{Hospital Street, Hermanus 7200}
+
 %(repeat as many times as is necessary)
@@
 
-\correspondingauthor{=name=}{=email address=}
-
-%% Keypoints, final entry on title page.
-
-% Example:
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 100 characters or less with no special characters or punctuation
-% \end{keypoints}
+\correspondingauthor{R.S. Weigel}{rweigel@gmu.edu}
 
@@
 \begin{keypoints}
-\item = enter point 1 here =
-\item = enter point 2 here =
-\item = enter point 3 here =
+\item The system parameters derived from GIC and geoelectric field measurements are frequency dependent
+\item A model with frequency dependence significantly outperforms a frequency independent model in predicting GIC
+\item The best predictions are found when the model input is not the measured geoelectric field but rather the geoelectric field predicted by a transfer function model driven by the geomagnetic field
 \end{keypoints}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT
-%
-% A good abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
 
@@
 
-%% ------------------------------------------------------------------------ %%
-%
-%  TEXT
-%
-%% ------------------------------------------------------------------------ %%
-
-%%% Suggested section heads:
-% \section{Introduction}
-%
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
-
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-%
-% \section{Data} (Or section title might be a descriptive heading about data)
-%
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-%
-% \section{Conclusions}
-
 \section{Introduction}
@@
 These fields are a result of time variations in ionospheric currents on timescales on the order of hours or less and the movement of Earth’s surface relative to slow-varying current systems in Earth’s ionosphere that are near stationary relative to the Sun [REF].
-Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure [REF].
-Accurate estimation of GICs given E or B is important for power system design [REF] retrospective analysis [REF] and for the mitigation of the impacts of space weather on power systems.
+Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}.
+Accurate estimation of GICs given direct measurements of the electric field (or estimates based on the geomagnetic field) is important for power system design [REF] retrospective analysis [REF] and for the mitigation of the impacts of space weather on power systems.
 
@@
 The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods [e.g., Boteler, 2014a; Boteler and Pirjola, 2014b].
-This assumption has been used or stated in many works in the past decade, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012}.
+This assumption has been used or stated in many works in the past decade, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.
 
@@
 
-One-second-cadence geoelectric field and geomagnetic field data were obtained from the Kakioka Magnetic Observatory of the Japan Meteorological Agency data portal [http://www.kakioka-jma.go.jp/obsdata/metadata/en].
+One-second-cadence geoelectric field and geomagnetic field data were obtained from the Kakioka Magnetic Observatory of the Japan Meteorological Agency data portal.
 Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure 1.
@@
 \acknowledgments
- = enter acknowledgments here =
+Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal http://www.kakioka-jma.go.jp/obsdata/metadata/en.
 

### 2019-01-11 agujournaltemplate.tex
Updates from ShareLaTeX

@@
 Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure 1.
+
+\begin{figure}[h]
+\centering
+\includegraphics[]{figures/plot_raw_All_20060402.pdf}
+\caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
+(Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) X km from the Memanbetsu substation.
+(Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.}
+\label{figure1}
+\end{figure}
 

### 2019-01-14 paper.tex

\documentclass[draft,linenumbers]{agujournal2018}
\usepackage{apacite}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{amsmath}

%%%%%%%
%\usepackage{trackchanges}
% uncomment the line above to use the TrackChanges package to mark revisions if needed.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

\journalname{Space Weather}

\begin{document}
\title{A Test of the Frequency Independence Assumption of the System Parameters used in Geomagnetically Induced Current Calculations}

\authors{R.S. Weigel\affil{1} and P.J. Cilliers\affil{2}}

\affiliation{1}{Space Weather Lab, George Mason University}
\affiliation{2}{South African National Space Agency}

\affiliation{1}{4400 University Drive, Fairfax VA 22030}
\affiliation{2}{Hospital Street, Hermanus 7200}

\correspondingauthor{R.S. Weigel}{rweigel@gmu.edu}

\begin{keypoints}
\item The power system parameters derived from GIC and geoelectric field measurements are frequency dependent.
\item A model with frequency dependence significantly outperforms a frequency independent model in predicting GIC.
\item The best predictions are found when the model input is not the measured geoelectric field but rather the geoelectric field predicted by a transfer function model driven by the geomagnetic field.
\end{keypoints}

\begin{abstract}
A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive. That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.  We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10 kilometers away.  We compute frequency dependent network parameters  $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are not frequency independent and that this model provides a significantly better representation of measured GIC.
\end{abstract}

\section{Introduction}

Geomagnetically induced currents (GICs) are electric currents in man-made conducting systems due to electric fields induced near Earth's surface. These fields are a result of time variations in ionospheric currents on timescales on the order of hours or less \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}. Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}. Accurate estimation of GICs given direct measurements of the electric field (or estimates based on the geomagnetic field) is important for power system design [REF] retrospective analysis [REF] and for the mitigation of the impacts of space weather on power systems.

A common assumption made in estimates of GIC using either measured or estimated values of the horizontal geoelectric field at Earth's surface, $\mathbf{E}$, is that the power system is resistive or quasi-DC so that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.  The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}). This assumption has been used or stated in many works in the past decade, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.

In this work, we use a unique dataset in which measurements $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$ were made near a site where GICs were being measured in an electric power system.  We first estimate the system parameters $a_o$ and $b_o$ using conventional methods, which assume that they are frequency independent.  Then, we use a model in which the system parameters are frequency dependent, compute the frequency dependence, and compare this model with the traditional frequency independent model.

\section{Data}

The 1-second-cadence GIC data used in this paper span a total of 36 days across the 10 intervals: 2006-04-02 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22 that have been presented previously in \citet{Watari2009} and \cite{Watari2015}.

One-second-cadence geoelectric field and geomagnetic field data from Kakioka Magnetic Observatory were obtained from the Japan Meteorological Agency data portal. Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{figure1}.

Unphysical spikes in the electric field measurements were concluded when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced with a linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.

The only modifications made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values. The values for all components were the same and much larger than the surrounding values.

The GIC dataset includes 1-second raw data and a 1 Hz low-pass filtered raw measurements.  The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure~\ref{figure1}.

The GIC data contains non-physical spikes that are followed by a relaxation time of approximately 60 seconds. These spikes were removed by identifying times $t_s$ when the change in $|GIC|$ was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ was replaced with a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_raw_All_20060402.pdf}
\caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB). (Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$10 km from the Memanbetsu substation. (Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co. The subscripts $x$, $y$, and $z$ correspond to North, East, and Downward, respectively.}
\label{figure1}
\end{figure}

\section{Models and Methods}

To determine the coefficients $a_o$ and $b_o$ in the model (Model 1)

\begin{linenomath*}
\begin{equation}
G_o(t) = a_oE_x(t) + b_oE_y(t)
\label{model1}
\end{equation}
\end{linenomath*}

\noindent
we form the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$, where GIC is a 864000x1 matrix of observations, $\mathbf{p} = [a,b]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations. The values of p are then obtained using a standard linear regression solver. The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ for all values of $t$. \citep[][ provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}

In Model 2, we assume that the system parameters are frequency dependent

\begin{linenomath*}
\begin{equation}
G_E(\omega) = A(\omega)E_x(\omega) + B(\omega)Ey(\omega)
\label{model2}
\tag{2a}
\end{equation}
\end{linenomath*}

\noindent
In the time domain, this is equivalent to

\begin{linenomath*}
\begin{equation}
G_E(t) = \sum_{\tau=-\infty}^{\infty}\big[a(\tau)E_x(t-\tau) + b(\tau)E_y(t-\tau)\big]
\label{model2b}
\tag{2b}
\end{equation}
\end{linenomath*}

\noindent
were the parameters $a$ and $A$ and $b$ and $B$ are inverse fourier transforms of each other.

The third model considered uses the electric field $\mathbf{E}'$ in Equation~\ref{model1}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\mathbf{Z}$: $\mathbf{E}'(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$:

\setcounter{equation}{2}
\begin{linenomath*}
\begin{equation}
G_{E'}(\omega) = A(\omega)E'_x(\omega) + B(\omega)E'_y(\omega)
\end{equation}
\end{linenomath*}

\noindent
where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{yx}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives

\begin{equation*}
G_{E'}(\omega) = \big[A(\omega)Z_{xx}(\omega) + B(\omega)Z_{yx}(\omega) \big] B_x(\omega) + \big[ A(\omega)Z_{xy}(\omega) + B(\omega)Z_{yy}(\omega) \big] B_y(\omega)
\end{equation*}

That is, instead of using the measured electric field directly as in Model 2, we use a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ to provide an estimate of the electric field based on the magnetic field. In this work, we have computed $\mathbf{Z}$ using the method described later in this section; $A(\omega)$ and $B(\omega)$ are taken from the solution of Equation~\ref{model2}, using the method also described later in this section.

A fourth model was also considered

\begin{linenomath*}
\begin{equation}
G_B(\omega) = Z_x(\omega)B_x(\omega) + Z_y(\omega)B_y(\omega)
\end{equation}
\end{linenomath*}

\noindent
In this model, the transfer function components $Z_x$ and $Z_y$ are determined directly from measurements of $GIC$ and $\mathbf{B}$ and is expected to produced predictions of $GIC$ that are similar to that obtained from Model 3.

The parameters in Model 2 and 4 can be computed using either a time or frequency domain method. In the time domain method, fixed values of the upper ($N_c$) and lower ($N_c$) limits are selected in the summations in Equation~\ref{model2b} (and equivalent for Equation 4) and then the equation is repeated for at least ($N_c+N_a+1$) time values and then the resulting over-determined set of equations is inverted to find the unknowns $a(\tau)$ and $b(\tau)$ for $N_a \le \tau \le Nc$.

In the frequency domain method, estimates of $A(\omega)$ and $B(\omega)$ are made at evaluation frequencies, $\omega_e$, by performing a least squares regression using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $\omega_e$. To convert the frequency domain transfer functions $A(\omega)$ and $B(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed. The transfer function magnitudes $|A(\omega)|$ and $|B(\omega)|$ were calculated by linear interpolation of the transfer function magnitudes (instead of linearly interpolating the real and imaginary components and then computing the magnitude). The same approach is used for solving for the transfer function $\mathbf{Z}$ in $\mathbf{E}(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$ - the $xx$ and $yx$ components of $\mathbf{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$, and the $yx$ and $yy$ components of $\mathbf{Z}$ are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$. Additional details on the method are given in \citep[and references therein.][]{Weigel2017}.

It is expected that the time and frequency domain methods should give similar results because the transfer functions in the time- and frequency-domains are mathematically related \citep{Schoukens2004,Ljung2007}, although differences may exist as a result of implementation choices or roundoff \citep{Ljung2004}.  For example, regarding implementation, in the time domain method we choose a finite value of $N_c$ and $N_a$ and in the frequency domain method, we choose the set of evaluation frequencies $\omega_e$, an averaging window function, and $\omega$ values (which depend on the length of the time series used to calculate the spectra).

Because the frequency domain computations are significantly faster, the results presented here are based on the frequency domain results and only spot checks were performed to verify that the time- and frequency-domain methods gave consistent results.

The frequency domain computations were performed on all possible unique 1-day window segments of available data, which, for Models 1, 2, and 4 resulted in 36 transfer functions as a function of the evaluation frequencies. The average transfer function magnitude was obtained by averaging the 36 transfer function magnitudes; the average transfer function phase was obtained using the average real and imaginary components, respectively, of the 36 transfer functions. The average impulse response functions were obtained by averaging the 36 impulse response functions.

There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that were developed to, for example, address leakage, data spikes and shifts, time intervals in which Equation~\ref{model2} is a poor model, bias, and the fact that errors in the fits of the transfer function at each evaluation frequency may not be Gaussian distributed [REFS]. Several of these methods were used to compute $A(\omega)$ and $B(\omega)$ including tapering the signals in the time domain using a Parzen and Slepian window (instead of box-car), robust regression (with either bi-square and Huber weighting) instead of linear regression, using weighted averages of the 36 1-day segments with the weighting at each frequency equal to the in-sample coherence at that frequency, and averaging the 36 time- or frequency-domain transfer functions using the Huber M estimate with a psi-function \citep{Huber2011}. All results were consistent in the sense that the transfer function values were within or near within each other's error bars. To simplify reproducibility, the method used was the most basic one which gave statistically similar results as methods that required additional or more complex calculations.

\section{Results}

In order to compare models, we have computed the out-of-sample prediction error of each model on a given day by computing the parameters in Models 1, 3, and 4 using the remaining 35 days of data. This process was repeated to create 36 out-of-sample predictions for each model. The 95\% confidence limits in the average prediction efficiencys and correlation coefficients were calculated using a 1000 bootstrap samples of the 36 prediction errors.

\begin{table}
\caption{}
\centering
\begin{tabular}{l l}
\hline
Model & PE  \\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 [0.25, 0.44] \\
2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.61 [0.53, 0.64] \\
3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 [0.73, 0.83] \\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.79 [0.75, 0.83] \\
\hline
\end{tabular}
\end{table}

The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as $PE$ = 1 - variance(Prediction Error)/variance(Predicted) and is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the predicted data that predicted by the model.

Based on the results in Table 1, we conclude that Model 1 is inferior to Model 2 in predicting GIC andhe $PEs$ of Model 3 and 4 are statistically equivalent, and Model 3, which uses magnetic field measurements as in input provides better estimates of GIC than Model 2. The last conclusion may be somewhat unexpected because Model 3 uses parameters from two different models. A proposed explanation is that the electric field measurements contain more noise than the magnetic field measurements.

The average parameters for Model 1 are $a_o =$ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 36 parameters calculated using each 1-day interval is shown in Figure 2. The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements. However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.68. In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$ [REFS] despite the fact that physically $\mathbf{E}$ drives $GIC$ and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GIC, but is a poor approximation relative to Models 2 and 3.

The average parameters for Models 2 and 4 are shown in Figure 3. The error bars at each frequency or time is the standard error of the 36 values used to compute the average. Figure 3a shows that the timescale of variations to impulse in E for Model 2 and B for Model 4 are similar.

Figure 3b shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition).

Figure 3c shows the transfer function phases. By definition, that for Model 1 is zero.

%Text here ===>>>

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%\appendix
%\section{Here is a sample appendix}

\acknowledgments
Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BibTeX is preferred:
%
\bibliography{paper.bib}
%
% no need to specify bibliographystyle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
%% Example \citet and \citep:
%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
%  \citet{Ghel00}, and \citet{Leit74}.

%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
%  \citep{Ghel00, Leit74}.

\end{document}

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

### 2019-01-14 figures/update.sh
figs

@@ -1,0 +1,7 @@
+DIR=../../figures/combined
+cp $DIR/plot_TF_aves_aobo_histograms-none-none-rectangular-parzen-ols_on_Z.pdf .
+cp $DIR/plot_TF_aves_GE_H-none-none-rectangular-parzen-ols_on_Z.pdf .
+cp $DIR/plot_TF_aves_GE_Phi-none-none-rectangular-parzen-ols_on_Z.pdf .
+cp $DIR/plot_TF_aves_GE_Z-none-none-rectangular-parzen-ols_on_Z.pdf .
+
+cp $DIR/../20060402/plot_raw_All_20060402.pdf .

### 2019-01-14 paper.tex
Updates from ShareLaTeX

@@
 
-%Text here ===>>>
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-none-none-rectangular-parzen-ols_on_Z.pdf}
+\caption{}
+\label{histogram}
+\end{figure}
 
-%%
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_GE_H-none-none-rectangular-parzen-ols_on_Z.pdf}
+\caption{}
+\label{H}
+\end{figure}
 
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_GE_Z-none-none-rectangular-parzen-ols_on_Z.pdf}
+\caption{}
+\label{Z}
+\end{figure}
 
-%\appendix
-%\section{Here is a sample appendix}
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_GE_Phi-none-none-rectangular-parzen-ols_on_Z.pdf}
+\caption{}
+\label{Phi}
+\end{figure}
 

### 2019-01-30 figures/update.sh
Update figs

@@ -1,7 +1,7 @@
 DIR=../../figures/combined
-cp $DIR/plot_TF_aves_aobo_histograms-none-none-rectangular-parzen-ols_on_Z.pdf .
-cp $DIR/plot_TF_aves_GE_H-none-none-rectangular-parzen-ols_on_Z.pdf .
-cp $DIR/plot_TF_aves_GE_Phi-none-none-rectangular-parzen-ols_on_Z.pdf .
-cp $DIR/plot_TF_aves_GE_Z-none-none-rectangular-parzen-ols_on_Z.pdf .
+cp $DIR/plot_TF_aves_aobo_histograms-options-1.pdf .
+cp $DIR/plot_TF_aves_H-options-1.pdf .
+cp $DIR/plot_TF_aves_Phi-options-1.pdf .
+cp $DIR/plot_TF_aves_Z-options-1.pdf .
 
 cp $DIR/../20060402/plot_raw_All_20060402.pdf .

### 2019-01-30 figures/update.sh
SN plot

@@ -3,5 +3,6 @@ cp $DIR/plot_TF_aves_aobo_histograms-options-1.pdf .
 cp $DIR/plot_TF_aves_H-options-1.pdf .
 cp $DIR/plot_TF_aves_Phi-options-1.pdf .
 cp $DIR/plot_TF_aves_Z-options-1.pdf .
+cp $DIR/plot_TF_aves_SN-options-1.pdf .
 
 cp $DIR/../20060402/plot_raw_All_20060402.pdf .

### 2019-01-30 paper.tex
Updates from ShareLaTeX

@@
 \begin{document}
-\title{A Test of the Frequency Independence Assumption of the System Parameters used in Geomagnetically Induced Current Calculations}
+\title{A Test of the Frequency Independence Assumption of the Power System Parameters used in Geomagnetically Induced Current Calculations}
 
@@
 \item The power system parameters derived from GIC and geoelectric field measurements are frequency dependent.
-\item A model with frequency dependence significantly outperforms a frequency independent model in predicting GIC.
+\item A GIC prediction model with frequency dependence significantly outperforms a frequency independent model.
 \item The best predictions are found when the model input is not the measured geoelectric field but rather the geoelectric field predicted by a transfer function model driven by the geomagnetic field.
@@
 We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10 kilometers away.
-We compute frequency dependent network parameters  $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are not frequency independent and that this model provides a significantly better representation of measured GIC.
+We compute frequency dependent network parameters  $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
 \end{abstract}
@@
 Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}.
-Accurate estimation of GICs given direct measurements of the electric field (or estimates based on the geomagnetic field) is important for power system design [REF] retrospective analysis [REF] and for the mitigation of the impacts of space weather on power systems.
+Accurate estimation of GICs given direct measurements of the electric field (or estimates based on the geomagnetic field) is important for power system design, retrospective analysis, and for the mitigation of the impacts of space weather on power systems [REFS].
 
 A common assumption made in estimates of GIC using either measured or estimated values of the horizontal geoelectric field at Earth's surface, $\mathbf{E}$, is that the power system is resistive or quasi-DC so that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.
-The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}).
+The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}.
 This assumption has been used or stated in many works in the past decade, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.
 
-In this work, we use a unique dataset in which measurements $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$ were made near a site where GICs were being measured in an electric power system.
+In this work, we use a unique dataset in which measurements $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$, were made near a site where GICs were being measured in an electric power system.
 We first estimate the system parameters $a_o$ and $b_o$ using conventional methods, which assume that they are frequency independent.
@@
 Unphysical spikes in the electric field measurements were concluded when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced with a linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
+In the sample shown in Figure~\ref{figure1}, 21 spikes were identified in $E_x$ and zero in $E_y$.
 
@@
 
-The GIC dataset includes 1-second raw data and a 1 Hz low-pass filtered raw measurements.
+The GIC dataset includes 1-second raw data and 1 Hz low-pass filtered raw measurements.
 The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure~\ref{figure1}.
 
-The GIC data contains non-physical spikes that are followed by a relaxation time of approximately 60 seconds.
+The GIC data contains non-physical spikes that are followed by a relaxation time of approximately 80 seconds.
 These spikes were removed by identifying times $t_s$ when the change in $|GIC|$ was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ was replaced with a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
+In the example interval shown in Figure~\ref{figure1}, 493 spikes were identified.
 
@@
 \noindent
-we form the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$, where GIC is a 864000x1 matrix of observations, $\mathbf{p} = [a,b]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations.
-The values of p are then obtained using a standard linear regression solver.
+we form the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$, where GIC is a 864000x1 matrix of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations.
+The values of $\mathbf{p}$ are then obtained using a standard linear regression solver.
 The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ for all values of $t$.
@@
 \begin{equation}
-G_E(\omega) = A(\omega)E_x(\omega) + B(\omega)Ey(\omega)
+G_E(\omega) = A(\omega)E_x(\omega) + B(\omega)E_y(\omega)
 \label{model2}
@@
 The same approach is used for solving for the transfer function $\mathbf{Z}$ in $\mathbf{E}(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$ - the $xx$ and $yx$ components of $\mathbf{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$, and the $yx$ and $yy$ components of $\mathbf{Z}$ are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$.
-Additional details on the method are given in \citep[and references therein.][]{Weigel2017}.
+Additional details on the method are given in \cite[][and references therein.]{Weigel2017}.
 
@@
 
-There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that were developed to, for example, address leakage, data spikes and shifts, time intervals in which Equation~\ref{model2} is a poor model, bias, and the fact that errors in the fits of the transfer function at each evaluation frequency may not be Gaussian distributed [REFS].
-Several of these methods were used to compute $A(\omega)$ and $B(\omega)$ including tapering the signals in the time domain using a Parzen and Slepian window (instead of box-car), robust regression (with either bi-square and Huber weighting) instead of linear regression, using weighted averages of the 36 1-day segments with the weighting at each frequency equal to the in-sample coherence at that frequency, and averaging the 36 time- or frequency-domain transfer functions using the Huber M estimate with a psi-function \citep{Huber2011}.
-All results were consistent in the sense that the transfer function values were within or near within each other's error bars.
-To simplify reproducibility, the method used was the most basic one which gave statistically similar results as methods that required additional or more complex calculations.
+There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that require additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Simpson2005}.
+The additional steps and processing were developed to address leakage, un-removed data spikes and shifts, high levels of noise contamination, time intervals in which the model assumptions are violated, bias, and the fact that errors in the fits of the transfer function at each evaluation frequency may not be Gaussian distributed.
+These works typically involve the estimation of a robust and low bias transfer function.
+In the results presented, we did not use a robust statistical method to calculate the transfer function or a bias correction algorithm - our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating low-bias model parameters.
+As discussed in ~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model.
+
+However, several alternative methods for computing the transfer functions were tested individually by modifying the basic method described above to determine if there was a change in the computed transfer functions and each model's prediction ability.
+When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, and there are differences in the prediction ability that are within the calculated uncertainty, and the transfer function values do not change significantly at frequencies where the signal to noise ratio is above $\sim$5.
+The experiments in this regard are discussed in the following section.
 
@@
 
-In order to compare models, we have computed the out-of-sample prediction error of each model on a given day by computing the parameters in Models 1, 3, and 4 using the remaining 35 days of data.
+In order to compare models, we have computed the out-of-sample prediction error of each model on a given day by computing the parameters in Models 1, 2, and 4 using the remaining 35 days of data.
 This process was repeated to create 36 out-of-sample predictions for each model.
@@
 
-Based on the results in Table 1, we conclude that Model 1 is inferior to Model 2 in predicting GIC andhe $PEs$ of Model 3 and 4 are statistically equivalent, and Model 3, which uses magnetic field measurements as in input provides better estimates of GIC than Model 2.
+Based on the results in Table 1, we conclude that Model 1 is inferior to Model 2 in predicting GIC and the $PEs$ of Model 3 and 4 are statistically equivalent.
+Model 3, which uses magnetic field measurements as in input provides better estimates of GIC than Model 2.
 The last conclusion may be somewhat unexpected because Model 3 uses parameters from two different models.
@@
 
-The average parameters for Models 2 and 4 are shown in Figure 3.
+The average parameters for Models 2 and 4 are shown in Figure~\ref{H}--\ref{Phi}.
 The error bars at each frequency or time is the standard error of the 36 values used to compute the average.
-Figure 3a shows that the timescale of variations to impulse in E for Model 2 and B for Model 4 are similar.
+Figure \ref{H} shows that the timescale of variations to impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar.
 
-Figure 3b shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition).
+Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition).
 
-Figure 3c shows the transfer function phases. By definition, that for Model 1 is zero.
+Figure~\ref{Phi} shows the transfer function phases. By definition, that for Model 1 is zero.
+
+Several of these methods were tested individually by modifying the method described above to determine their impact on the computed transfer functions and their over all results.
+
+\begin{itemize}
+\item Using robust regression (with either bi-square or Huber weighting) instead of linear regression when computing the transfer function estimates at each $\omega_e$
+
+\item Averaging the 36 transfer functions using the Huber M-estimator with a psi-function \citep{Huber2011}.
+
+\item Using weighted averages of the 36 transfer functions with the weighting at each frequency equal to the in-sample coherence at that frequency
+
+\item Windowing the signals in the frequency domain using a rectangular window (instead of Parzen)
+\end{itemize}
 
@@
 \includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-none-none-rectangular-parzen-ols_on_Z.pdf}
-\caption{}
+\caption{Distribution of parameters in Model 1.
+Each set of parameters were computed 36 times using one day of 1-second cadence data.
+The dots indicate the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
 \label{histogram}
@@
 \includegraphics[width=\textwidth]{figures/plot_TF_aves_GE_H-none-none-rectangular-parzen-ols_on_Z.pdf}
-\caption{}
+\caption{Impulse responses for Models 1, 2, and 4.
+The circles at $t = 0$ correspond to the impulse response for Model 1; the red/blue lines are the impulse responses for Model 2 and 4.
+The values of $a_o$ and $b_o$ have been scaled by a factor of 10 for readability.}
 \label{H}
@@
 \includegraphics[width=\textwidth]{figures/plot_TF_aves_GE_Z-none-none-rectangular-parzen-ols_on_Z.pdf}
-\caption{}
+\caption{Frequency domain transfer functions for Models 1-3. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
 \label{Z}
@@
 \includegraphics[width=\textwidth]{figures/plot_TF_aves_GE_Phi-none-none-rectangular-parzen-ols_on_Z.pdf}
-\caption{}
+\caption{Frequency domain phase values for Models 1-3. The phase at each frequency for parameter $p$ is arg$(p)$.}
 \label{Phi}
@@
 
+\section{Discussion and Conclusions}
+
 \acknowledgments
 Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
+GIC data provided by ...
 
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% BibTeX is preferred:
-%
 \bibliography{paper.bib}
-%
-% no need to specify bibliographystyle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-01-30 figures/update.sh
update figs

@@ -1,3 +1,4 @@
+git pull
 DIR=../../figures/combined
 cp $DIR/plot_TF_aves_aobo_histograms-options-1.pdf .
 cp $DIR/plot_TF_aves_H-options-1.pdf .
@@ -6,3 +7,5 @@ cp $DIR/plot_TF_aves_Z-options-1.pdf .
 cp $DIR/plot_TF_aves_SN-options-1.pdf .
 
 cp $DIR/../20060402/plot_raw_All_20060402.pdf .
+git commit -a -m "update figs"
+git push

### 2019-01-30 paper.tex
Updates from ShareLaTeX

@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-none-none-rectangular-parzen-ols_on_Z.pdf}
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-options-1.pdf}
 \caption{Distribution of parameters in Model 1.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_GE_H-none-none-rectangular-parzen-ols_on_Z.pdf}
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_H-options-1.pdf}
 \caption{Impulse responses for Models 1, 2, and 4.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_GE_Z-none-none-rectangular-parzen-ols_on_Z.pdf}
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_Z-options-1.pdf}
 \caption{Frequency domain transfer functions for Models 1-3. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_GE_Phi-none-none-rectangular-parzen-ols_on_Z.pdf}
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_Phi-options-1.pdf}
 \caption{Frequency domain phase values for Models 1-3. The phase at each frequency for parameter $p$ is arg$(p)$.}
+\label{Phi}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
+\caption{}
 \label{Phi}

### 2019-02-01 figures/update.sh
update figs

@@ -1,4 +1,7 @@
 git pull
+
+cp ../figures/20060805/plot_model_predictions-MeanModel-20060806T150000.pdf .
+
 DIR=../../figures/combined
 cp $DIR/plot_TF_aves_aobo_histograms-options-1.pdf .
 cp $DIR/plot_TF_aves_H-options-1.pdf .

### 2019-02-01 paper.tex

@@
 Geomagnetically induced currents (GICs) are electric currents in man-made conducting systems due to electric fields induced near Earth's surface.
-These fields are a result of time variations in ionospheric currents on timescales on the order of hours or less \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
+These fields are a result of time variations in ionospheric currents on timescales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
 Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}.
@@
 The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}.
-This assumption has been used or stated in many works in the past decade, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.
+This assumption has been used or stated in many works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.
 
@@
 
-The 1-second-cadence GIC data used in this paper span a total of 36 days across the 10 intervals: 2006-04-02 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22 that have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
+The 1-second-cadence GIC data used in this paper span a total of 35 days across the 10 intervals: 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22 that have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
 
@@
 
-Unphysical spikes in the electric field measurements were concluded when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced with a linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
+Unphysical spikes in the electric field measurements assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
 In the sample shown in Figure~\ref{figure1}, 21 spikes were identified in $E_x$ and zero in $E_y$.
@@
 The only modifications made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values.
-The values for all components were the same and much larger than the surrounding values.
+The values for $E_x$ and $E_y$ were the same and much larger than the surrounding values.
 
@@
 The GIC data contains non-physical spikes that are followed by a relaxation time of approximately 80 seconds.
-These spikes were removed by identifying times $t_s$ when the change in $|GIC|$ was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ was replaced with a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
-In the example interval shown in Figure~\ref{figure1}, 493 spikes were identified.
+These spikes were removed by identifying times $t_s$ when the change in $|GIC|$ was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ was replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
+In the example interval shown in Figure~\ref{figure1}, 493 spikes were removed.
 
@@
 
-To determine the coefficients $a_o$ and $b_o$ in the model (Model 1)
+To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model 1,
 
@@
 
-A fourth model was also considered
+A fourth model was also considered:
 
@@
 
-The frequency domain computations were performed on all possible unique 1-day window segments of available data, which, for Models 1, 2, and 4 resulted in 36 transfer functions as a function of the evaluation frequencies.
-The average transfer function magnitude was obtained by averaging the 36 transfer function magnitudes; the average transfer function phase was obtained using the average real and imaginary components, respectively, of the 36 transfer functions.
-The average impulse response functions were obtained by averaging the 36 impulse response functions.
+Transfer function parameters were computed for each of the 35 unique 1-day window segments of available data.
+The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the average real and imaginary components, respectively, of the 35 transfer functions.
+The average impulse response functions were obtained by averaging the 35 impulse response functions.
 
-There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that require additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Simpson2005}.
+There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that use additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Simpson2005}.
 The additional steps and processing were developed to address leakage, un-removed data spikes and shifts, high levels of noise contamination, time intervals in which the model assumptions are violated, bias, and the fact that errors in the fits of the transfer function at each evaluation frequency may not be Gaussian distributed.
@@
 However, several alternative methods for computing the transfer functions were tested individually by modifying the basic method described above to determine if there was a change in the computed transfer functions and each model's prediction ability.
-When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, and there are differences in the prediction ability that are within the calculated uncertainty, and the transfer function values do not change significantly at frequencies where the signal to noise ratio is above $\sim$5.
+When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance is within the calculated uncertainty of the used method, and the transfer function values are within the error bars of the used method when the signal to noise ratio is above $\sim$5.
 The experiments in this regard are discussed in the following section.
@@
 
-In order to compare models, we have computed the out-of-sample prediction error of each model on a given day by computing the parameters in Models 1, 2, and 4 using the remaining 35 days of data.
-This process was repeated to create 36 out-of-sample predictions for each model.
-The 95\% confidence limits in the average prediction efficiencys and correlation coefficients were calculated using a 1000 bootstrap samples of the 36 prediction errors.
+The out-of-sample prediction metrics for each model on a given day was determined by computing the average model parameters using the remaining 34 days of data, generating a prediction time series on each of the 34 days using the average model parameters, and then computing the prediction metrics for each of the 34 days.
+This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
+
+The performance of each model in predicting $GIC$ was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.
+
+The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as $PE$ = 1 - variance(Prediction Error)/variance(Predicted) and is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the predicted data that predicted by the model.
+
+The average MSE ratio was computed by averaging the 35 MSE ratios and the average signal-to-noise ratio at a given frequency is the average of the 35 signal-to-noise ratios at that frequency.
+
+The 95\% confidence interval, CI, on the  metrics was calculated using 1000 bootstrap samples of the 35 prediction metrics.
+The typical difference between the limits when the confidence intervals were computed assuming the values were gaussian-distributed was $\sim$2%.
 
 \begin{table}
-\caption{}
+\caption{Model performance metrics for models $m=1-4$.}
 \centering
-\begin{tabular}{l l}
+\begin{tabular}{l c c c c}
 \hline
-Model & PE  \\
+$m$\hspace{1em} Model & PE & 95\% CI & MSE$_m$/MSE$_1$ & 95\% CI\\
 \hline
-1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 [0.25, 0.44] \\
-2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.61 [0.53, 0.64] \\
-3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 [0.73, 0.83] \\
-4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.79 [0.75, 0.83] \\
+1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.37 & [0.31, 0.43] & 1.0 & \\
+2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6,2.0]\\
+3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8,3.4]\\
+4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8,5.3]\\
 \hline
@@
 
-The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as $PE$ = 1 - variance(Prediction Error)/variance(Predicted) and is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the predicted data that predicted by the model.
-
-Based on the results in Table 1, we conclude that Model 1 is inferior to Model 2 in predicting GIC and the $PEs$ of Model 3 and 4 are statistically equivalent.
-Model 3, which uses magnetic field measurements as in input provides better estimates of GIC than Model 2.
+The primary conclusion from the results shown in Table~1 is that the frequency dependent model, Model 2, is significantly better at predicting $GIC$ than the frequency independent model, Model 1.
+Model 3, which uses magnetic field measurements as in input provides better estimates of $GIC$ than Model 2.
 The last conclusion may be somewhat unexpected because Model 3 uses parameters from two different models.
@@
 
-The average parameters for Model 1 are $a_o =$ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 36 parameters calculated using each 1-day interval is shown in Figure 2.
+The average parameters for Model 1 are $a_o =$ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~2.
 The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements.
-However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.68.
-In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$ [REFS] despite the fact that physically $\mathbf{E}$ drives $GIC$ and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GIC, but is a poor approximation relative to Models 2 and 3.
+However, the model does have a good prediction performance - the PE is 0.37 and the data/model correlation coefficient of 0.64.
+In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$ [REFS] despite the fact that physically $\mathbf{E}$ drives $GIC$ and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of $GIC$, but is a poor approximation relative to Models 2 and 3.
 
 The average parameters for Models 2 and 4 are shown in Figure~\ref{H}--\ref{Phi}.
-The error bars at each frequency or time is the standard error of the 36 values used to compute the average.
+The error bars at each frequency or time is the standard error of the 35 values used to compute the average.
 Figure \ref{H} shows that the timescale of variations to impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar.
@@
 
-\item Averaging the 36 transfer functions using the Huber M-estimator with a psi-function \citep{Huber2011}.
+\item Averaging the 35 transfer functions using the Huber M-estimator with a psi-function \citep{Huber2011}.
 
-\item Using weighted averages of the 36 transfer functions with the weighting at each frequency equal to the in-sample coherence at that frequency
+\item Using weighted averages of the 35 transfer functions with the weighting at each frequency equal to the in-sample coherence at that frequency
 
@@
 \centering
+\includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
+\caption{Signal-to-Noise}
+\label{SN}
+\end{figure}
+
+\begin{figure}[h]
+\centering
 \includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-options-1.pdf}
 \caption{Distribution of parameters in Model 1.
-Each set of parameters were computed 36 times using one day of 1-second cadence data.
+Each set of parameters were computed 35 times using one day of 1-second cadence data.
 The dots indicate the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
@@
 \caption{Frequency domain phase values for Models 1-3. The phase at each frequency for parameter $p$ is arg$(p)$.}
-\label{Phi}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
-\caption{}
 \label{Phi}

### 2019-02-01 figures/update.sh
update figs

@@ -1,6 +1,6 @@
 git pull
 
-cp ../figures/20060805/plot_model_predictions-MeanModel-20060806T150000.pdf .
+cp ../../figures/20060805/plot_model_predictions-MeanModel-20060806T150000.pdf .
 
 DIR=../../figures/combined
 cp $DIR/plot_TF_aves_aobo_histograms-options-1.pdf .

### 2019-02-01 paper.tex
Updates from ShareLaTeX

@@
 One-second-cadence geoelectric field and geomagnetic field data from Kakioka Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
-Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{figure1}.
+Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{sample}.
 
 Unphysical spikes in the electric field measurements assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
-In the sample shown in Figure~\ref{figure1}, 21 spikes were identified in $E_x$ and zero in $E_y$.
+In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$.
 
@@
 The GIC dataset includes 1-second raw data and 1 Hz low-pass filtered raw measurements.
-The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure~\ref{figure1}.
+The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure~\ref{sample}.
 
@@
 These spikes were removed by identifying times $t_s$ when the change in $|GIC|$ was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ was replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
-In the example interval shown in Figure~\ref{figure1}, 493 spikes were removed.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{figures/plot_raw_All_20060402.pdf}
-\caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
-(Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$10 km from the Memanbetsu substation.
-(Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
-The subscripts $x$, $y$, and $z$ correspond to North, East, and Downward, respectively.}
-\label{figure1}
-\end{figure}
+In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
 
@@
 
-The average MSE ratio was computed by averaging the 35 MSE ratios and the average signal-to-noise ratio at a given frequency is the average of the 35 signal-to-noise ratios at that frequency.
-
-The 95\% confidence interval, CI, on the  metrics was calculated using 1000 bootstrap samples of the 35 prediction metrics.
-The typical difference between the limits when the confidence intervals were computed assuming the values were gaussian-distributed was $\sim$2%.
+The average PE and MSE ratio was computed by averaging the 35 PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the 35 signal-to-noise ratios at that frequency.
+The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the 35 prediction metrics.
+The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.
 
 \begin{table}
-\caption{Model performance metrics for models $m=1-4$.}
+\caption{Out-of-sample model performance metrics for models $m=1-4$.
+$MSE_m/MSE_1$ is the ratio of the MSE for model $m$ to that of Model 1.
+The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples of these 35 values.}
 \centering
@@
 \hline
-1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.37 & [0.31, 0.43] & 1.0 & \\
+1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
 2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6,2.0]\\
@@
 
-The primary conclusion from the results shown in Table~1 is that the frequency dependent model, Model 2, is significantly better at predicting $GIC$ than the frequency independent model, Model 1.
+Figure~\ref{predictions} shows a 1-day interval for which the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 4 has the highest PE and Model 1 has the lowest.
+
+The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting $GIC$ than the frequency independent model, Model 1.
 Model 3, which uses magnetic field measurements as in input provides better estimates of $GIC$ than Model 2.
-The last conclusion may be somewhat unexpected because Model 3 uses parameters from two different models.
-A proposed explanation is that the electric field measurements contain more noise than the magnetic field measurements.
+Model 4, in which a transfer function that connects $GIC$ to $\mathbf{B}$ was derived directly from the data, provides the best predictions of $GIC$.
+
+Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models.
+The error bars represent one standard error of the 35 SN ratios at each period.
+Consistent with the results in Table~1, the SN ratio for Model 1 is consistently lower than that for Models 2-4.
+Above $1400$~s (24~min), the SN ratio ordering is consistent with the PE and MSE ratios shown in Table~1.
+Below $500$~s (8~min), Model 2 has the largest SN ratio.
+This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} - at approximately 2006-08-07 07:00, the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.
 
@@
 The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements.
-However, the model does have a good prediction performance - the PE is 0.37 and the data/model correlation coefficient of 0.64.
+However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
 In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$ [REFS] despite the fact that physically $\mathbf{E}$ drives $GIC$ and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of $GIC$, but is a poor approximation relative to Models 2 and 3.
+For the data considered in this work, when a model $G(t) = c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.
 
-The average parameters for Models 2 and 4 are shown in Figure~\ref{H}--\ref{Phi}.
-The error bars at each frequency or time is the standard error of the 35 values used to compute the average.
-Figure \ref{H} shows that the timescale of variations to impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar.
+The average parameters for Models 2 and 4 are shown in Figure~\ref{H}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.
+
+Figure \ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar and are different from zero at more than one time value (the impulse response for Model 3 is similar to that of Model 2 and is not shown).
+By definition, the impulse response for Model 1 is zero except at $t=0$.
+Note that the sign of the impulse response at $t=0$ is opposite for Model 1 and Model 2 for the $E_y$ coefficient - at $t=0$, $b_o < 0$ while $b(0)>0$.
 
 Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition).
+For this system, GIC is most dependent on $E_y$, which has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).
 
@@
 
-\item Averaging the 35 transfer functions using the Huber M-estimator with a psi-function \citep{Huber2011}.
+\item Averaging the 35 transfer functions using the Huber M-estimator with a $\psi$-function \citep{Huber2011}.
 
@@
 \centering
+\includegraphics[width=\textwidth]{figures/plot_raw_All_20060402.pdf}
+\caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
+(Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$10 km from the Memanbetsu substation.
+(Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
+The subscripts $x$, $y$, and $z$ correspond to North, East, and Downward, respectively.}
+\label{sample}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-20060806T150000.pdf}
+\caption{Out-of-sample model predictions and predictions errors for a 1-day interval.
+The bottom overlapping traces are the measured GIC (black line) and the model predictions (colored lines).
+The prediction error of each model is shown shifted by an arbitrary offset.}
+\label{predictions}
+\end{figure}
+
+\begin{figure}[h]
+\centering
 \includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
-\caption{Signal-to-Noise}
+\caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
 \label{SN}
@@
 Each set of parameters were computed 35 times using one day of 1-second cadence data.
-The dots indicate the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
+The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
 \label{histogram}
@@
 \includegraphics[width=\textwidth]{figures/plot_TF_aves_Phi-options-1.pdf}
-\caption{Frequency domain phase values for Models 1-3. The phase at each frequency for parameter $p$ is arg$(p)$.}
+\caption{Frequency domain phase values for Models 1-3.}
 \label{Phi}

### 2019-02-01 paper.tex
Updates from ShareLaTeX

@@
 
-The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting $GIC$ than the frequency independent model, Model 1.
+The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1.
 Model 3, which uses magnetic field measurements as in input provides better estimates of $GIC$ than Model 2.
@@
 \includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-20060806T150000.pdf}
-\caption{Out-of-sample model predictions and predictions errors for a 1-day interval.
-The bottom overlapping traces are the measured GIC (black line) and the model predictions (colored lines).
-The prediction error of each model is shown shifted by an arbitrary offset.}
+\caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval. Note that the scales for Models 2-4 differ from that for Model 1.}
 \label{predictions}

### 2019-02-01 figures/update.sh
update figs

@@ -9,6 +9,6 @@ cp $DIR/plot_TF_aves_Phi-options-1.pdf .
 cp $DIR/plot_TF_aves_Z-options-1.pdf .
 cp $DIR/plot_TF_aves_SN-options-1.pdf .
 
-cp $DIR/../20060402/plot_raw_All_20060402.pdf .
+cp $DIR/../20080805/plot_raw_All_20060805.pdf .
 git commit -a -m "update figs"
 git push

### 2019-02-01 figures/update.sh
update figs

@@ -9,6 +9,6 @@ cp $DIR/plot_TF_aves_Phi-options-1.pdf .
 cp $DIR/plot_TF_aves_Z-options-1.pdf .
 cp $DIR/plot_TF_aves_SN-options-1.pdf .
 
-cp $DIR/../20080805/plot_raw_All_20060805.pdf .
+cp $DIR/../20060805/plot_raw_All_20060805.pdf .
 git commit -a -m "update figs"
 git push

### 2019-02-11 paper.tex
update figs

@@
 
+\section{Discussion and Conclusions}
+
+\acknowledgments
+Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
+GIC data provided by ...
+
+\bibliography{paper.bib}
+
+\clearpage
+
 \begin{figure}[h]
 \centering
-\includegraphics[width=\textwidth]{figures/plot_raw_All_20060402.pdf}
+\includegraphics[width=\textwidth]{figures/plot_raw_All_20060805.pdf}
 \caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_SN-options-1.pdf}
 \caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1.pdf}
 \caption{Distribution of parameters in Model 1.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_H-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_H-options-1.pdf}
 \caption{Impulse responses for Models 1, 2, and 4.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_Z-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1.pdf}
 \caption{Frequency domain transfer functions for Models 1-3. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_Phi-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
 \caption{Frequency domain phase values for Models 1-3.}
@@
 \end{figure}
-
-\section{Discussion and Conclusions}
-
-\acknowledgments
-Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
-GIC data provided by ...
-
-\bibliography{paper.bib}
 

### 2019-02-11 figures/update.sh

@@ -3,11 +3,11 @@ git pull
 cp ../../figures/20060805/plot_model_predictions-MeanModel-20060806T150000.pdf .
 
 DIR=../../figures/combined
-cp $DIR/plot_TF_aves_aobo_histograms-options-1.pdf .
-cp $DIR/plot_TF_aves_H-options-1.pdf .
-cp $DIR/plot_TF_aves_Phi-options-1.pdf .
-cp $DIR/plot_TF_aves_Z-options-1.pdf .
-cp $DIR/plot_TF_aves_SN-options-1.pdf .
+cp $DIR/main_plot_summary_aobo_histograms-options-1.pdf .
+cp $DIR/main_plot_summary_H-options-1.pdf .
+cp $DIR/main_plot_summary_Phi-options-1.pdf .
+cp $DIR/main_plot_summary_Z-options-1.pdf .
+cp $DIR/main_plot_summary_SN-options-1.pdf .
 
 cp $DIR/../20060805/plot_raw_All_20060805.pdf .
 git commit -a -m "update figs"

### 2019-02-11 figures/update.sh
update figs

@@ -3,11 +3,11 @@ git pull
 cp ../../figures/20060805/plot_model_predictions-MeanModel-20060806T150000.pdf .
 
 DIR=../../figures/combined
-cp $DIR/main_plot_summary_aobo_histograms-options-1.pdf .
-cp $DIR/main_plot_summary_H-options-1.pdf .
-cp $DIR/main_plot_summary_Phi-options-1.pdf .
-cp $DIR/main_plot_summary_Z-options-1.pdf .
-cp $DIR/main_plot_summary_SN-options-1.pdf .
+cp $DIR/plot_model_summary_aobo_histograms-options-1.pdf .
+cp $DIR/plot_model_summary_H-options-1.pdf .
+cp $DIR/plot_model_summary_Phi-options-1.pdf .
+cp $DIR/plot_model_summary_Z-options-1.pdf .
+cp $DIR/plot_model_summary_SN-options-1.pdf .
 
 cp $DIR/../20060805/plot_raw_All_20060805.pdf .
 git commit -a -m "update figs"

### 2019-02-11 paper.tex
Updates from Overleaf

--- 2019-02-01 paper.tex
@@
 \usepackage{amsmath}
+\usepackage{xcolor}
+\usepackage[colorlinks]{hyperref}
+\usepackage[colorinlistoftodos]{todonotes}
+%\usepackage{framebox}
 
@@
 % complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%
+%%%%%%%https://www.overleaf.com/project/5c3620127117e2522d9e0c30
 
@@
 \begin{document}
+
+% Not working
+% \listoftodos
+
 \title{A Test of the Frequency Independence Assumption of the Power System Parameters used in Geomagnetically Induced Current Calculations}
@@
 That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.
-We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10 kilometers away.
+We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory \todo[color=blue!40]{Shinichi: Is 10 km correct?} $\sim${\color{red}\textbf{10}} kilometers away.
 We compute frequency dependent network parameters  $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
@@
 Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}.
-Accurate estimation of GICs given direct measurements of the electric field (or estimates based on the geomagnetic field) is important for power system design, retrospective analysis, and for the mitigation of the impacts of space weather on power systems [REFS].
+Accurate estimation of GICs given direct measurements of the electric field (or estimates based on the geomagnetic field) is important for power system design, retrospective analysis, and for the mitigation of the impacts of space weather on power systems \todo{Pierre and Trevor - any suggestions?} {\color{red} \textbf{[REFS]}}.
 
@@
 The 1-second-cadence GIC data used in this paper span a total of 35 days across the 10 intervals: 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22 that have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
+This represents all of the data that were available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.
 
-One-second-cadence geoelectric field and geomagnetic field data from Kakioka Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
+One-second-cadence geoelectric field and geomagnetic field data from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
 Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{sample}.
+Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.
 
-Unphysical spikes in the electric field measurements assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
+Spikes in the electric field measurements assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
 In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$.
+For all of the time intervals considered, the fraction of interpolated $\mathbf{E}$ values ranged from 0--0.1\%.
 
@@
 The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure~\ref{sample}.
+For all of the time intervals considered, the fraction of interpolated GIC values ranged from 1--4\%.
 
-The GIC data contains non-physical spikes that are followed by a relaxation time of approximately 80 seconds.
-These spikes were removed by identifying times $t_s$ when the change in $|GIC|$ was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ was replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
+The GIC data contains non-physical spikes that are followed by a relaxation time of  40-80 seconds.
+These spikes were removed by identifying times $t_s$ when the change in 1-second $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
 In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
 
-\section{Models and Methods}
+\section{Models}
 
-To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model 1,
+To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model~1,
 
@@
 \noindent
-we form the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$, where GIC is a 864000x1 matrix of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations.
+the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $GIC$ is a 864000x1 matrix of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations.
 The values of $\mathbf{p}$ are then obtained using a standard linear regression solver.
@@
 
-In Model 2, we assume that the system parameters are frequency dependent
+In Model 2, the system parameters are frequency dependent
 
@@
 
-\noindent
-In the time domain, this is equivalent to
+Note that the scalar quantity $B$ without a subscript refers to the system parameter in Model~2 while the vector quantity $\mathbf{B}$ or its components, $B_x$ and $B_y$ refer to the magnetic field.
 
-\begin{linenomath*}
-\begin{equation}
-G_E(t) = \sum_{\tau=-\infty}^{\infty}\big[a(\tau)E_x(t-\tau) + b(\tau)E_y(t-\tau)\big]
-\label{model2b}
-\tag{2b}
-\end{equation}
-\end{linenomath*}
-
-\noindent
-were the parameters $a$ and $A$ and $b$ and $B$ are inverse fourier transforms of each other.
-
-The third model considered uses the electric field $\mathbf{E}'$ in Equation~\ref{model1}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\mathbf{Z}$: $\mathbf{E}'(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$:
+The third model considered uses the electric field $\mathbf{E}'$ in Equation~\ref{model2}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\mathbf{Z}$: $\mathbf{E}'(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$:
 
@@
 \noindent
-where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{yx}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives
+where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives
 
@@
 That is, instead of using the measured electric field directly as in Model 2, we use a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ to provide an estimate of the electric field based on the magnetic field.
-In this work, we have computed $\mathbf{Z}$ using the method described later in this section; $A(\omega)$ and $B(\omega)$ are taken from the solution of Equation~\ref{model2}, using the method also described later in this section.
+In this work, we have computed $\mathbf{Z}$ using the method described later in this section.
 
-A fourth model was also considered:
+A fourth model considered is:
 
@@
 \noindent
-In this model, the transfer function components $Z_x$ and $Z_y$ are determined directly from measurements of $GIC$ and $\mathbf{B}$ and is expected to produced predictions of $GIC$ that are similar to that obtained from Model 3.
+In this model, the transfer function components $Z_x$ and $Z_y$ are determined directly from measurements of $GIC$ and $\mathbf{B}$ and the model is expected to produce predictions of $GIC$ that are similar to that obtained from Model 3.
 
-The parameters in Model 2 and 4 can be computed using either a time or frequency domain method.
-In the time domain method, fixed values of the upper ($N_c$) and lower ($N_c$) limits are selected in the summations in Equation~\ref{model2b} (and equivalent for Equation 4) and then the equation is repeated for at least ($N_c+N_a+1$) time values and then the resulting over-determined set of equations is inverted to find the unknowns $a(\tau)$ and $b(\tau)$ for $N_a \le \tau \le Nc$.
-
-In the frequency domain method, estimates of $A(\omega)$ and $B(\omega)$ are made at evaluation frequencies, $\omega_e$, by performing a least squares regression using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $\omega_e$.
+For frequency domain method, estimates of $A(\omega)$ and $B(\omega)$ are made at evaluation frequencies, $\omega_e$, by performing a least squares regression using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $\omega_e$.
 To convert the frequency domain transfer functions $A(\omega)$ and $B(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed.
@@
 The same approach is used for solving for the transfer function $\mathbf{Z}$ in $\mathbf{E}(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$ - the $xx$ and $yx$ components of $\mathbf{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$, and the $yx$ and $yy$ components of $\mathbf{Z}$ are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$.
-Additional details on the method are given in \cite[][and references therein.]{Weigel2017}.
+Additional details on the method are given in \cite{Weigel2017} and references therein.
 
-It is expected that the time and frequency domain methods should give similar results because the transfer functions in the time- and frequency-domains are mathematically related \citep{Schoukens2004,Ljung2007}, although differences may exist as a result of implementation choices or roundoff \citep{Ljung2004}.
-For example, regarding implementation, in the time domain method we choose a finite value of $N_c$ and $N_a$ and in the frequency domain method, we choose the set of evaluation frequencies $\omega_e$, an averaging window function, and $\omega$ values (which depend on the length of the time series used to calculate the spectra).
+%It is expected that the time and frequency domain methods should give similar results because the transfer functions in the time- and frequency-domains are mathematically related \citep{Schoukens2004,Ljung2007}, although differences may exist as a result of implementation choices or roundoff \citep{Ljung2004}.
+For example, regarding implementation, in the time domain method we choose a finite value of $N_c$ and $N_a$ and in the frequency domain method, we choose the set of evaluation frequencies $\omega_e$, an averaging window function and width, and $\omega$ values (which depend on the length of the time series used to calculate the spectra).
 
-Because the frequency domain computations are significantly faster, the results presented here are based on the frequency domain results and only spot checks were performed to verify that the time- and frequency-domain methods gave consistent results.
+%Because the frequency domain computations are significantly faster, the results presented here are based on the frequency domain results and only spot checks were performed to verify that the time- and frequency-domain methods gave consistent results.
 
 Transfer function parameters were computed for each of the 35 unique 1-day window segments of available data.
-The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the average real and imaginary components, respectively, of the 35 transfer functions.
+The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components of the 35 transfer functions.
 The average impulse response functions were obtained by averaging the 35 impulse response functions.
 
-There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that use additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Simpson2005}.
-The additional steps and processing were developed to address leakage, un-removed data spikes and shifts, high levels of noise contamination, time intervals in which the model assumptions are violated, bias, and the fact that errors in the fits of the transfer function at each evaluation frequency may not be Gaussian distributed.
-These works typically involve the estimation of a robust and low bias transfer function.
-In the results presented, we did not use a robust statistical method to calculate the transfer function or a bias correction algorithm - our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating low-bias model parameters.
-As discussed in ~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model.
+There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that use additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic least-squares method \citep{Simpson2005}.
+These works typically involve the estimation of a robust and bias-corrected transfer function.
+The results presented used were obtained using the basic least-squares transfer function estimation method as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating low-bias model parameters.
+As discussed in ~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).
 
-However, several alternative methods for computing the transfer functions were tested individually by modifying the basic method described above to determine if there was a change in the computed transfer functions and each model's prediction ability.
-When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance is within the calculated uncertainty of the used method, and the transfer function values are within the error bars of the used method when the signal to noise ratio is above $\sim$5.
-The experiments in this regard are discussed in the following section.
+However, several alternative methods for computing the transfer functions were tested individually to determine if there was a change in the computed transfer functions and, most important for this work, each model's out-of-sample prediction ability.
+When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance are within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$ seconds.
+The experiments in this regard are discussed in the following section and in the Appendix.
 
@@
 
-The performance of each model in predicting $GIC$ was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.
+The performance of each model in predicting GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.
 
@@
 \caption{Out-of-sample model performance metrics for models $m=1-4$.
-$MSE_m/MSE_1$ is the ratio of the MSE for model $m$ to that of Model 1.
+MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1.
 The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples of these 35 values.}
@@
 1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
-2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6,2.0]\\
-3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8,3.4]\\
-4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8,5.3]\\
+2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
+3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
+4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
 \hline
@@
 The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1.
-Model 3, which uses magnetic field measurements as in input provides better estimates of $GIC$ than Model 2.
-Model 4, in which a transfer function that connects $GIC$ to $\mathbf{B}$ was derived directly from the data, provides the best predictions of $GIC$.
+Model 3, which uses magnetic field measurements as in input provides better estimates of GIC than Model 2.
+Model 4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides the best predictions of GIC.
 
 Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models.
-The error bars represent one standard error of the 35 SN ratios at each period.
-Consistent with the results in Table~1, the SN ratio for Model 1 is consistently lower than that for Models 2-4.
-Above $1400$~s (24~min), the SN ratio ordering is consistent with the PE and MSE ratios shown in Table~1.
+The error bars extend by $\pm$ one standard error of the 35 SN ratios at each period.
+The SN ratio for Model 1 is always lower than that for Models 2-4.
+Above $1400$~s (24~min), the SN ratio ordering is consistent with the ordering of the PE and MSE ratios shown in Table~1.
 Below $500$~s (8~min), Model 2 has the largest SN ratio.
@@
 
-The average parameters for Model 1 are $a_o =$ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~2.
+The average parameters for Model 1 are $a_o = $ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~2.
+(Results are similar if 1-minute averages are used; \cite{Watari2015} found \todo[color=blue!40]{Shinichi: I get $a_o=45.8$ and $b_o=-59.9$ in this time interval.} $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).
+
 The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements.
 However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
-In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$ [REFS] despite the fact that physically $\mathbf{E}$ drives $GIC$ and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of $GIC$, but is a poor approximation relative to Models 2 and 3.
-For the data considered in this work, when a model $G(t) = c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.
+In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of GIC \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GIC, but is a poor approximation relative to Models 2 and 3.
+For the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.
 
@@
 Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition).
-For this system, GIC is most dependent on $E_y$, which has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+Note that the ratio of $|\overline{b}_o/||\overline{a}_o|$ is approximately 1 for Model~1 but varies from 1.6 to 6.3 for Model~2.
+In Model~2, GIC is most dependent on $E_y$, however note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}) so that $E_x$ and $E_y$ on average contribute approximately equal amounts to the predictability of GIC.
 
-Figure~\ref{Phi} shows the transfer function phases. By definition, that for Model 1 is zero.
+Figure~\ref{Phi} shows the transfer function phases.
+By definition, that for Model 1 is zero.
+The phases for $Z_x$ and $B$ are offset by 120 degrees \todo{Why?} while the phases for $A$ and $B$ are significantly different \todo{Why?}.
 
-Several of these methods were tested individually by modifying the method described above to determine their impact on the computed transfer functions and their over all results.
+Although the primary objective of this work is the creation of the models for comparison of the predictive ability, a natural outcome is model parameter estimates.
+As noted in the introduction, the motivation for using a basic least squares method is (1) that it reduces the number of adjustable methodological parameters that must be considered and (2) because we are attempting to consider which model performs best in a least-square-error sense.
+It is natural to ask if the model parameter estimates are biased and how sensitive they depend on the statistical method used.
+Several alternative estimation methods have been considered and are discussed in the Appendix; the primary conclusion is that the values shown in Table~1 are not sensitive and the parameter estimates are only sensitive below approximately $T=30$~s (they are unreliable in the sense that the alternative estimation methods give parameter values outside of the error bars shown for the baseline method).
 
-\begin{itemize}
-\item Using robust regression (with either bi-square or Huber weighting) instead of linear regression when computing the transfer function estimates at each $\omega_e$
+\section{Discussion and Conclusions}
 
-\item Averaging the 35 transfer functions using the Huber M-estimator with a $\psi$-function \citep{Huber2011}.
+Several methods have been used in the literature for estimating GIC.
+When electric field measurements are available, a model of the form of Model~1 is used \todo{TODO} {\color{red} \textbf{[REFS]}}.
+We have shown using a data set of contemporaneous GIC and $\mathbf{E}$ measurements that a significantly better estimation of GIC can be obtained by using a model of the form of Model~2.
 
-\item Using weighted averages of the 35 transfer functions with the weighting at each frequency equal to the in-sample coherence at that frequency
+When only a MT impedance tensor that connects $\mathbf{E}$ to $\mathbf{B}$ is available, a model of the form of Model~1, with the electric field estimated using $\mathbf{E}=\mathbf{Z}\mathbf{B}$ \todo{TODO} {\color{red} \textbf{[REFS]}}.
+For the dataset considered, significantly better estimates can be made if frequency dependent system parameters are available as in Model~2.
 
-\item Windowing the signals in the frequency domain using a rectangular window (instead of Parzen)
-\end{itemize}
+When only magnetic field data is available, a model of the form of Model~4 is used.
+In this case, the transfer function contains components of the transfer function that connects $\mathbf{E}$ to $\mathbf{B}$ and GIC to $\mathbf{E}$.
+
+Because Model~4 provided the best prediction performance, the ideal measurements to use for estimating GIC in this dataset are contemporaneous GIC and $\mathbf{B}$ measurements.
+
+\todo{TODO} {\color{red}\textbf{Discuss the origin of the frequency dependence of the system parameters.}}
+
+\acknowledgments
+Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
+GIC data provided by ...
+\todo[color=blue!40]{Shinichi: How should this be referenced?}
+
+\clearpage
 
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_raw_All_20060402.pdf}
+\includegraphics[width=\textwidth]{figures/plot_raw_All_20060805.pdf}
 \caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
@@
 \includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-20060806T150000.pdf}
-\caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval. Note that the scales for Models 2-4 differ from that for Model 1.}
+\caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval.
+Note that the scales for Models 2-4 differ from that for Model 1 so that the error amplitudes for Models 2-4 appear larger than that of Model 1 by a factor of 2.}
 \label{predictions}
@@
 \includegraphics[width=\textwidth]{figures/plot_TF_aves_Z-options-1.pdf}
-\caption{Frequency domain transfer functions for Models 1-3. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
+\caption{Frequency domain transfer functions for Models 1, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
 \label{Z}
@@
 \includegraphics[width=\textwidth]{figures/plot_TF_aves_Phi-options-1.pdf}
-\caption{Frequency domain phase values for Models 1-3.}
+\caption{Frequency domain phase values for Models 1, 2, and 4.}
 \label{Phi}
@@
 
-\section{Discussion and Conclusions}
+\clearpage
 
-\acknowledgments
-Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
-GIC data provided by ...
+\appendix
+\section{Appendix}
+
+Many alternatives to the basic least-squares for the calculation of transfer functions have been proposed and are strongly advocated by their developers who have shown that clearly problematic data will ....
+For example, as it has stated that the least-squares method often fails \citep{Egbert2011} or that failures are sometimes spectacular but more often insidious~\citep{Chave1987}.
+Several of these methods were tested individually by modifying the method described above to determine their impact on the computed transfer functions and their overall results.
+The primary conclusion is that different methods generally give the same results except when the signal to noise ratio is low (below approximately $30$ seconds).
+Although plotted because they were a part of the prediction models and they do contribute somewhat to the prediction efficiency (as discussed in \cite{Weigel2017}, below $30$ seconds, the model parameter estimates should be regarded as highly uncertain due to large error bars and a large sensitivity to the method used to calculate them.
+
+\begin{itemize}
+\item Using robust regression instead of linear regression when computing the transfer function estimates at each $\omega_e$ (using a custom implementation with the $\psi$-function referenced in the next item and also using the \texttt{robustfit} function in MATLAB with bi-square and Huber weighting options).
+The PEs and MSEs shown in Table~1 were the same to two decimal places and the confidence intervals were smaller by $\sim 2$\% (when the confidence interval was calculated assuming a Gaussian distribution instead of using the bootstrap, the confidence intervals varied by $\sim\pm 2$\%).
+
+\item Five methods for averaging were used to compute the magnitudes at each evaluation frequency of model parameter $p(D)$, where $D$ is estimate made using a 1-day.
+If $u(D)$ and $v(D)$ are the real and imaginary components of $p$, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation).
+The PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$.
+Neither averaging method was yielded consistently higher or lower values.
+
+\item Pre-whitening each 1-day segment prior to calculation of the transfer functions.
+This was done by computing the pre-whitening filter using a 10th-order Yule-Walker filter computed using $B_x$ and applying this filter to all time series used in the calculation of the transfer function.
+The PEs were the same to within two decimal places and the MSE ratios varied by $0.05$.
+
+\item Windowing the signals in the frequency domain using a rectangular window (instead of Parzen).
+The PEs for Model 3 and Model 4 were higher by $0.02$ and $0.01$, respectively and the MSE ratios for Models 3 and 4 were higher by $0.4$ and $0.6$, respectively.
+
+\item Using weighted averages of the 35 transfer functions with the weighting at each frequency equal to the in-sample coherence at that frequency (to give higher weight to better fitting intervals) and weighting the collection of values used to compute the model parameters at a given evaluation frequency by a smoothed estimate of the PSD of the predictand (to account for the fact that the signal to noise ratio depends on frequency as evident in ~\ref{SN} and noted by \citep{Egbert1997}).
+
+\item Instead of calculating 35 transfer functions and then averaging, compute one transfer function - for each evaluation frequency, combine the 35 fourier spectra values near that evaluation frequency and use them for one regression.
+\end{itemize}
+
+In all cases, the parameter estimates below $\sim 20$ seconds using the alternative methods were consistently lower, but the offset was of magnitude that was similar to the variation observed between the alternative methods.
+The largest variations were found in relative to the method used to averaging the 35 transfer functions.
+In addition, how the averages were made when computing the magnitudes mattered - if the magnitudes at a given frequency were calculated by taking the modulus of the mean of the 35 real and mean of the 35 imaginary components vs the mean of the modulus of the 35 transfer functions.
+The same was true for the median.
 

### 2019-02-11 paper.tex
Updates from Overleaf

@@
 That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.
-We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory \todo[color=blue!40]{Shinichi: Is 10 km correct?} $\sim${\color{red}\textbf{10}} kilometers away.
+We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10\todo[color=blue!40]{Shinichi: Is 10 km correct?} $\sim${\color{red}\textbf{10}} kilometers away.
 We compute frequency dependent network parameters  $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
@@
 
-One-second-cadence geoelectric field and geomagnetic field data from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
+One-second-cadence geoelectric field and geomagnetic field data from KakiokaMemambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
 Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{sample}.
-Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.
+
+Unphysical s Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.
 
@@
 
-The GIC data contains non-physical spikes that are followed by a relaxation time of  40-80 seconds.
-These spikes were removed by identifying times $t_s$ when the change in 1-second $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
+The GIC data contains non-physical spikes that are followed by a relaxation time of approximately  40-80 seconds.
+These spikes were removed by identifying times $t_s$ when the change in $|GIC|$1-second $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ wasere replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
 In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
 
-\section{Models}
+\section{Models and Methods}
 
-To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model~1,
+To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model ~1,
 
@@
 \noindent
-the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $GIC$ is a 864000x1 matrix of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations.
+we form the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $GIC$ is a 864000x1 matrix of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations.
 The values of $\mathbf{p}$ are then obtained using a standard linear regression solver.
@@
 
-In Model 2, the system parameters are frequency dependent
+In Model 2, we assume that the system parameters are frequency dependent
 
@@
 
-Note that the scalar quantity $B$ without a subscript refers to the system parameter in Model~2 while the vector quantity $\mathbf{B}$ or its components, $B_x$ and $B_y$ refer to the magnetic field.
+\noindent
+In the time domain, this is equivalent to
 
-The third model considered uses the electric field $\mathbf{E}'$ in Equation~\ref{model2}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\mathbf{Z}$: $\mathbf{E}'(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$:
+\begin{linenomath*}
+\begin{equation}
+G_E(t) = \sum_{\tau=-\infty}^{\infty}\big[a(\tau)E_x(t-\tau) + b(\tau)E_y(t-\tau)\big]
+\label{model2b}
+\tag{2b}
+\end{eNote that the scalar quantity $B$ without a subscript refers to the system parameter in Model~2 while the vector quantion}
+\end{linenomath*}
+
+\noindent
+were the parameters $a$ and $A$ and $bty $\mathbf{B}$ or its components, $B_x$ and $B_y$ are inverse fourier transforms of each otherfer to the magnetic field.
+
+The third model considered uses the electric field $\mathbf{E}'$ in Equation~\ref{model12}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\mathbf{Z}$: $\mathbf{E}'(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$:
 
@@
 \noindent
-where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives
+where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xyx}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives
 
@@
 That is, instead of using the measured electric field directly as in Model 2, we use a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ to provide an estimate of the electric field based on the magnetic field.
-In this work, we have computed $\mathbf{Z}$ using the method described later in this section.
+In this work, we have computed $\mathbf{Z}$ using the method described later in this section; $A(\omega)$ and $B(\omega)$ are taken from the solution of Equation~\ref{model2}, using the method also described later in this section.
 
-A fourth model considered is:
+A fourth model was also considered is:
 
@@
 \noindent
-In this model, the transfer function components $Z_x$ and $Z_y$ are determined directly from measurements of $GIC$ and $\mathbf{B}$ and the model is expected to produce predictions of $GIC$ that are similar to that obtained from Model 3.
+In this model, the transfer function components $Z_x$ and $Z_y$ are determined directly from measurements of $GIC$ and $\mathbf{B}$ and the model is expected to produced predictions of $GIC$ that are similar to that obtained from Model 3.
 
-For frequency domain method, estimates of $A(\omega)$ and $B(\omega)$ are made at evaluation frequencies, $\omega_e$, by performing a least squares regression using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $\omega_e$.
+The parameters in Model 2 and 4 can be computed using either a time or frequency domain method.
+In the time domain method, fixed values of the upper ($N_c$) and lower ($N_c$) limits are selected in the summations in Equation~\ref{model2b} (and equivalent for Equation 4) and then the equation is repeated for at least ($N_c+N_a+1$) time values and then the resulting over-determined set of equations is inverted to find the unknowns $a(\tau)$ and $b(\tau)$ for $N_a \le \tau \le Nc$.
+
+In theFor frequency domain method, estimates of $A(\omega)$ and $B(\omega)$ are made at evaluation frequencies, $\omega_e$, by performing a least squares regression using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $\omega_e$.
 To convert the frequency domain transfer functions $A(\omega)$ and $B(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed.
@@
 The same approach is used for solving for the transfer function $\mathbf{Z}$ in $\mathbf{E}(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$ - the $xx$ and $yx$ components of $\mathbf{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$, and the $yx$ and $yy$ components of $\mathbf{Z}$ are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$.
-Additional details on the method are given in \cite{Weigel2017} and references therein.
+Additional details on the method are given in \cite[][{Weigel2017} and references therein.]{Weigel2017}.
 
@@
 Transfer function parameters were computed for each of the 35 unique 1-day window segments of available data.
-The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components of the 35 transfer functions.
+The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components, respectively, of the 35 transfer functions.
 The average impulse response functions were obtained by averaging the 35 impulse response functions.
 
-There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that use additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic least-squares method \citep{Simpson2005}.
-These works typically involve the estimation of a robust and bias-corrected transfer function.
-The results presented used were obtained using the basic least-squares transfer function estimation method as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating low-bias model parameters.
+There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that use additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Simpson2005}.
+The additional steps and processing were developed to address leakage, un-removed data spikes and shifts, high levels of noise contamination, time intervals in which the model assumptions are violated, bias, and the fact that errors in the fits of the transfer function at each evaluation frequency may not be Gaussian distributedleast-squares method \citep{Simpson2005}.
+These works typically involve the estimation of a robust and low bias-corrected transfer function.
+In tThe results presented, we did not use a robust statistical method to calculate the transfer function or a bias correction algorithm - used were obtained using the basic least-squares transfer function estimation method as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating low-bias model parameters.
 As discussed in ~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).
 
-However, several alternative methods for computing the transfer functions were tested individually to determine if there was a change in the computed transfer functions and, most important for this work, each model's out-of-sample prediction ability.
-When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance are within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$ seconds.
+However, several alternative methods for computing the transfer functions were tested individually by modifying the basic method described above to determine if there was a change in the computed transfer functions and each model's, most important for this work, each model's out-of-sample prediction ability.
+When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance isare within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$53 (corresponding to a period of $\sim 30$ seconds.
 The experiments in this regard are discussed in the following section and in the Appendix.
@@
 
-The performance of each model in predicting GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.
+The performance of each model in predicting $GIC$ was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.
 
@@
 \caption{Out-of-sample model performance metrics for models $m=1-4$.
-MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1.
+$MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1.
 The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples of these 35 values.}
@@
 The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1.
-Model 3, which uses magnetic field measurements as in input provides better estimates of GIC than Model 2.
-Model 4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides the best predictions of GIC.
+Model 3, which uses magnetic field measurements as in input provides better estimates of $GIC$ than Model 2.
+Model 4, in which a transfer function that connects $GIC$ to $\mathbf{B}$ was derived directly from the data, provides the best predictions of $GIC$.
 
 Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models.
-The error bars extend by $\pm$ one standard error of the 35 SN ratios at each period.
-The SN ratio for Model 1 is always lower than that for Models 2-4.
+The error bars representextend by $\pm$ one standard error of the 35 SN ratios at each period.
+Consistent with the results in Table~1, tThe SN ratio for Model 1 is consistentlyalways lower than that for Models 2-4.
 Above $1400$~s (24~min), the SN ratio ordering is consistent with the ordering of the PE and MSE ratios shown in Table~1.
@@
 However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
-In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of GIC \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GIC, but is a poor approximation relative to Models 2 and 3.
+In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$  \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives $GIC$ and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of $GIC$, but is a poor approximation relative to Models 2 and 3.
 For the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.
@@
 Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition).
-Note that the ratio of $|\overline{b}_o/||\overline{a}_o|$ is approximately 1 for Model~1 but varies from 1.6 to 6.3 for Model~2.
-In Model~2, GIC is most dependent on $E_y$, however note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}) so that $E_x$ and $E_y$ on average contribute approximately equal amounts to the predictability of GIC.
+For this systemNote that the ratio of $|\overline{b}_o/||\overline{a}_o|$ is approximately 1 for Model~1 but varies from 1.6 to 6.3 for Model~2.
+In Model~2, GIC is most dependent on $E_y$, whichhowever note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+
+Figure~\ref{Phi} shows the transfer function phases. By definition, that for Model 1 is zero.
+
+Several of these methods were tested individually by modifying the method described above to determine their impact on the computed transfer functions and their over all results.
+
+\begin{itemize}
+\item Using robust regression (with either bi-square or Huber weighting) instead of linear regression when comput so that $E_x$ and $E_y$ on average contribute approximately equal amounts to the predictability of GIC.
 
@@
 It is natural to ask if the model parameter estimates are biased and how sensitive they depend on the statistical method used.
-Several alternative estimation methods have been considered and are discussed in the Appendix; the primary conclusion is that the values shown in Table~1 are not sensitive and the parameter estimates are only sensitive below approximately $T=30$~s (they are unreliable in the sense that the alternative estimation methods give parameter values outside of the error bars shown for the baseline method).
+Several alternative estimation methods have been considered and are discussed in the Appendix; the primary conclusion is that the values shown in Table~1 are not sensitive and the parameter estimates are only sensitive below approximately $T=30$~s (they are unreliable ing the transfer function estimates at each $\omega_e$
+
+\item Averaging the 35 transfer functions using the Huber M-estimator with a $\psi$-function \citep{Huber2011}.
+
+\item Using weighted averages of the 35 transfer functions with the weighting at each frequency equal to the in-sample coherence at that frequency
+
+\item Windowing the signals in the frequency dosense that the alternative estimation methods give parameter values outside of the error bars shown for the baseline method).
 
@@
 
-Because Model~4 provided the best prediction performance, the ideal measurements to use for estimating GIC in this dataset are contemporaneous GIC and $\mathbf{B}$ measurements.
+Because Model~4 provided the best prediction performance, the ideal measurements to use for estimatin using a rectangular window (instead of Parzen)
+\end{itemize}
+
+\section{Discussion and Conclusions}
+g GIC in this dataset are contemporaneous GIC and $\mathbf{B}$ measurements.
 
@@
 GIC data provided by ...
+
+\bibliography{paper.bib}
  \todo[color=blue!40]{Shinichi: How should this be referenced?}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_SN-options-1.pdf}
 \caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_aobo_histograms-options-1.pdf}
 \caption{Distribution of parameters in Model 1.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_H-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_H-options-1.pdf}
 \caption{Impulse responses for Models 1, 2, and 4.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_Z-options-1.pdf}
-\caption{Frequency domain transfer functions for Models 1, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
+\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_Z-options-1.pdf}
+\caption{Frequency domain transfer functions for Models 1-3, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
 \label{Z}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_Phi-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
+\caption{Frequency doTF_aves_Phi-options-1.pdf}
 \caption{Frequency domain phase values for Models 1, 2, and 4.}
@@
 \item Five methods for averaging were used to compute the magnitudes at each evaluation frequency of model parameter $p(D)$, where $D$ is estimate made using a 1-day.
-If $u(D)$ and $v(D)$ are the real and imaginary components of $p$, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation).
+If $u(D)$ and $v(D)$ are the real and imagin phase values for Models 1-3.}
+\label{Phi}
+\end{figure}
+ary components of $p$, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation).
 The PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$.

### 2019-02-11 paper.tex
update figs

@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_SN-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_SN-options-1.pdf}
 \caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_aobo_histograms-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1.pdf}
 \caption{Distribution of parameters in Model 1.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_H-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_H-options-1.pdf}
 \caption{Impulse responses for Models 1, 2, and 4.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_Z-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1.pdf}
 \caption{Frequency domain transfer functions for Models 1-3, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}

### 2019-02-14 paper.tex
Updates from Overleaf

--- 2019-02-11 paper.tex
@@
 \item A GIC prediction model with frequency dependence significantly outperforms a frequency independent model.
-\item The best predictions are found when the model input is not the measured geoelectric field but rather the geoelectric field predicted by a transfer function model driven by the geomagnetic field.
+\item A model derived using GIC/geomagnetic field data provides better predictions than a GIC/geoelectric field-based model.
 \end{keypoints}
@@
 That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.
-We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory \todo[color=blue!40]{Shinichi: Is 10 km correct?} $\sim${\color{red}\textbf{10}} kilometers away.
+We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10\todo[color=blue!40]{Shinichi: Is 10 km correct?} kilometers away.
 We compute frequency dependent network parameters $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
@@
 
-Geomagnetically induced currents (GICs) are electric currents in man-made conducting systems due to electric fields induced near Earth's surface.
+Geomagnetically induced currents (GICs) are electric currents in conducting systems due to electric fields induced near Earth's surface.
 These fields are a result of time variations in ionospheric currents on timescales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
 Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}.
-Accurate estimation of GICs given direct measurements of the electric field (or estimates based on the geomagnetic field) is important for power system design, retrospective analysis, and for the mitigation of the impacts of space weather on power systems \todo{Pierre and Trevor - any suggestions?} {\color{red} \textbf{[REFS]}}.
+Accurate estimation of GICs given direct measurements of the electric field (or estimates of the electric field based on the geomagnetic field) is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \todo{Pierre and Trevor - any suggestions?} {\color{red} \textbf{[REFS]}}.
 
-A common assumption made in estimates of GIC using either measured or estimated values of the horizontal geoelectric field at Earth's surface, $\mathbf{E}$, is that the power system is resistive or quasi-DC so that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.
-The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}.
-This assumption has been used or stated in many works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.
+A common assumption made in estimates of GIC using either measured or estimated values of the horizontal geoelectric field at Earth's surface, $\mathbf{E}$, is that the power system is resistive or quasi-DC in the sense that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.
+In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression, when contemporaneous measurements of GIC and $\mathbf{E}$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}.
+This quasi-DC assumption has been used or stated in many works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.
 
@@
 
-The 1-second-cadence GIC data used in this paper span a total of 35 days across the 10 intervals: 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22 that have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
-This represents all of the data that were available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.
+The 1-second-cadence GIC data used span a total of 35 days across 10 intervals: 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22.
+These data have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
+These time intervals cover those for which GIC data was made available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.
+The GIC dataset includes 1-s raw data and 1 Hz low-pass filtered raw measurements.
+The results presented here are for the 1 Hz low-pass filtered GIC values
 
-One-second-cadence geoelectric field and geomagnetic field data from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
-Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{sample}.
+One-second-cadence $\mathbf{E}$ and $\mathbf{B}$ measurements a from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
 Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.
 
-Spikes in the electric field measurements assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
+Data for the longest continuous time span of GIC measurements, 2006-04-02 through 2006-04-10 are shown Figure~\ref{sample}.
+
+Spikes in the electric field measurements were assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
 In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$.
@@
 
-The only modifications made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values.
-The values for $E_x$ and $E_y$ were the same and much larger than the surrounding values.
+The only modification made to any of the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values.
+The values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.
 
-The GIC dataset includes 1-second raw data and 1 Hz low-pass filtered raw measurements.
-The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure~\ref{sample}.
+The GIC data contains non-physical spikes that are followed by a relaxation time of 40-80 s.
+These spikes were removed by identifying times $t_s$ when the change in 1-se $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
+In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
 For all of the time intervals considered, the fraction of interpolated GIC values ranged from 1--4\%.
 
-The GIC data contains non-physical spikes that are followed by a relaxation time of  40-80 seconds.
-These spikes were removed by identifying times $t_s$ when the change in 1-second $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
-In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
+\section{Models and Methods}
 
-\section{Models}
-
-To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model~1,
+To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model ~1,
 
@@
 \noindent
-the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $GIC$ is a 864000x1 matrix of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations.
+the matrix equation $\mathbf{GIC} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $\mathbf{GIC}$ is a 864000x1 matrix of 1 day of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ observations.
 The values of $\mathbf{p}$ are then obtained using a standard linear regression solver.
-The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ for all values of $t$.
-\citep[][ provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}
+The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ over a 1-day interval.
+\citep[][ provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}.
+The parameters $a_o$ and $b_o$ were computed for each of the 35 unique 1-day window segments of available data.
 
-In Model 2, the system parameters are frequency dependent
+In Model 2 the system parameters are frequency dependent
 
@@
 \begin{equation}
-G_E(\omega) = A(\omega)E_x(\omega) + B(\omega)E_y(\omega)
+G_E(\omega) = a(\omega)E_x(\omega) + b(\omega)E_y(\omega)\,.
 \label{model2}
-\tag{2a}
 \end{equation}
@@
 
-Note that the scalar quantity $B$ without a subscript refers to the system parameter in Model~2 while the vector quantity $\mathbf{B}$ or its components, $B_x$ and $B_y$ refer to the magnetic field.
-
-The third model considered uses the electric field $\mathbf{E}'$ in Equation~\ref{model2}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\mathbf{Z}$: $\mathbf{E}'(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$:
+The third model considered uses an electric field $\mathbf{E}'$ in Equation~\ref{model2}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$:
 
@@
 \begin{equation}
-G_{E'}(\omega) = A(\omega)E'_x(\omega) + B(\omega)E'_y(\omega)
+G_{E'}(\omega) = a(\omega)E'_x(\omega) + b(\omega)E'_y(\omega)
+\label{model3}
 \end{equation}
@@
 \begin{equation*}
-G_{E'}(\omega) = \big[A(\omega)Z_{xx}(\omega) + B(\omega)Z_{yx}(\omega) \big] B_x(\omega) + \big[ A(\omega)Z_{xy}(\omega) + B(\omega)Z_{yy}(\omega) \big] B_y(\omega)
+G_{E'}(\omega) = \big[a(\omega)Z_{xx}(\omega) + b(\omega)Z_{yx}(\omega) \big] B_x(\omega) + \big[ a(\omega)Z_{xy}(\omega) + b(\omega)Z_{yy}(\omega) \big] B_y(\omega)\,.
 \end{equation*}
 
-That is, instead of using the measured electric field directly as in Model 2, we use a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ to provide an estimate of the electric field based on the magnetic field.
-In this work, we have computed $\mathbf{Z}$ using the method described later in this section.
+That is, instead of using the measured electric field directly as in Model 2, a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.
+In this work, $\boldsymbol{Z}$ was calculated using the method described later in this section; $a(\omega)$ and $b(\omega)$ are taken from the solution of Equation~\ref{model2}, using the method also described later in this section.
 
-A fourth model considered is:
+The fourth model is
 
@@
 \begin{equation}
-G_B(\omega) = Z_x(\omega)B_x(\omega) + Z_y(\omega)B_y(\omega)
+G_B(\omega) = z_x(\omega)B_x(\omega) + z_y(\omega)B_y(\omega)\,.
+\label{model4}
 \end{equation}
@@
 \noindent
-In this model, the transfer function components $Z_x$ and $Z_y$ are determined directly from measurements of $GIC$ and $\mathbf{B}$ and the model is expected to produce predictions of $GIC$ that are similar to that obtained from Model 3.
+In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of GIC and $\mathbf{B}$ and the model is expected to produced predictions of GICs that are similar to that obtained from Model 3.
 
-For frequency domain method, estimates of $A(\omega)$ and $B(\omega)$ are made at evaluation frequencies, $\omega_e$, by performing a least squares regression using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $\omega_e$.
-To convert the frequency domain transfer functions $A(\omega)$ and $B(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed.
-The transfer function magnitudes $|A(\omega)|$ and $|B(\omega)|$ were calculated by linear interpolation of the transfer function magnitudes (instead of linearly interpolating the real and imaginary components and then computing the magnitude).
-The same approach is used for solving for the transfer function $\mathbf{Z}$ in $\mathbf{E}(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$ - the $xx$ and $yx$ components of $\mathbf{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$, and the $yx$ and $yy$ components of $\mathbf{Z}$ are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$.
-Additional details on the method are given in \cite{Weigel2017} and references therein.
+Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e = 2\pi f_e$, by performing a least squares regression using Equation~\ref{model2} and the set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $f_e$ from the fourier transform of 24 hours of 1-second data.
+The values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$; if $f_e$ value is not a multiple of $1/86400$, the nearest multiple is chosen.
+The regression for $f_e$ uses all $f$ values between the next higher and lower value of $f_e$.
+
+To convert the frequency domain transfer functions $a(\omega)$ and $b(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed.
+The transfer function magnitudes $|a(\omega)|$ and $|b(\omega)|$ were calculated by linear interpolation of the transfer function magnitudes (instead of linearly interpolating the real and imaginary components and then computing the magnitude).
+The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}.
+For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$, the $xx$ and $yx$ components of $\boldsymbol{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$; the $yx$ and $yy$ components are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$.
 
@@
 Transfer function parameters were computed for each of the 35 unique 1-day window segments of available data.
-The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components of the 35 transfer functions.
+The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components, respectively, of the 35 transfer functions.
 The average impulse response functions were obtained by averaging the 35 impulse response functions.
 
-There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that use additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic least-squares method \citep{Simpson2005}.
+There are many statistical methods that have been developed for estimating transfer functions in the magnetotelluric literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Simpson2005}.
 These works typically involve the estimation of a robust and bias-corrected transfer function.
-The results presented used were obtained using the basic least-squares transfer function estimation method as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating low-bias model parameters.
-As discussed in ~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).
+In the results presented, we did not use a robust statistical method or a bias correction algorithm as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters.
+As discussed in~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).
 
-However, several alternative methods for computing the transfer functions were tested individually to determine if there was a change in the computed transfer functions and, most important for this work, each model's out-of-sample prediction ability.
-When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance are within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$ seconds.
-The experiments in this regard are discussed in the following section and in the Appendix.
+However, several alternative methods were tested individually by modifying the basic method described above to determine if there was a change in the computed transfer functions and, most important for this work, the out-of-sample prediction performance of each model.
+When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance is within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s).
+The experiments in this regard are discussed in the Appendix.
 
@@
 
-The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as $PE$ = 1 - variance(Prediction Error)/variance(Predicted) and is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the predicted data that predicted by the model.
+The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the predicted data that predicted by the model.
 
-The average PE and MSE ratio was computed by averaging the 35 PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the 35 signal-to-noise ratios at that frequency.
+The average PE and MSE ratio was computed by averaging the 35 out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the out-of-sample 35 signal-to-noise ratios at that frequency.
 The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the 35 prediction metrics.
@@
 
-Figure~\ref{predictions} shows a 1-day interval for which the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 4 has the highest PE and Model 1 has the lowest.
+Figure~\ref{predictions} shows a 1-day interval for which the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the lowest.
 
@@
 Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models.
-The error bars extend by $\pm$ one standard error of the 35 SN ratios at each period.
-The SN ratio for Model 1 is always lower than that for Models 2-4.
-Above $1400$~s (24~min), the SN ratio ordering is consistent with the ordering of the PE and MSE ratios shown in Table~1.
+The error bars represent $\pm$ one standard error of the 35 SN ratios at each period.
+Consistent with the results in Table~1, The SN ratio for Model 1 is lower than that for Models 2-4 at all frequencies.
+Above $1400$~s (24~min), the SN ratio ordering is consistent with the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio.
 Below $500$~s (8~min), Model 2 has the largest SN ratio.
-This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} - at approximately 2006-08-07 07:00, the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.
+This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.
 
-The average parameters for Model 1 are $a_o = $ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~2.
+The average parameters for Model 1 are $a_o = $ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~\ref{histogram}.
 (Results are similar if 1-minute averages are used; \cite{Watari2015} found \todo[color=blue!40]{Shinichi: I get $a_o=45.8$ and $b_o=-59.9$ in this time interval.} $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).
 
-The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements.
+The wide distribution and large error bars are suggestive that Model~\ref{model1} is a poor representation of the measurements.
 However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
-In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of GIC \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GIC, but is a poor approximation relative to Models 2 and 3.
+In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$  \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GICs, but is a poor approximation relative to Models 2 and 3.
 For the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.
@@
 
-Figure \ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar and are different from zero at more than one time value (the impulse response for Model 3 is similar to that of Model 2 and is not shown).
+Figure~\ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar and are different from zero at more than one time value (the impulse response for Model 3 is similar to that of Model 2 and is not shown).
 By definition, the impulse response for Model 1 is zero except at $t=0$.
@@
 Note that the ratio of $|\overline{b}_o/||\overline{a}_o|$ is approximately 1 for Model~1 but varies from 1.6 to 6.3 for Model~2.
-In Model~2, GIC is most dependent on $E_y$, however note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}) so that $E_x$ and $E_y$ on average contribute approximately equal amounts to the predictability of GIC.
+In Model~2, GIC is most dependent on $E_y$, however note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).
 
@@
 By definition, that for Model 1 is zero.
-The phases for $Z_x$ and $B$ are offset by 120 degrees \todo{Why?} while the phases for $A$ and $B$ are significantly different \todo{Why?}.
-
-Although the primary objective of this work is the creation of the models for comparison of the predictive ability, a natural outcome is model parameter estimates.
-As noted in the introduction, the motivation for using a basic least squares method is (1) that it reduces the number of adjustable methodological parameters that must be considered and (2) because we are attempting to consider which model performs best in a least-square-error sense.
-It is natural to ask if the model parameter estimates are biased and how sensitive they depend on the statistical method used.
-Several alternative estimation methods have been considered and are discussed in the Appendix; the primary conclusion is that the values shown in Table~1 are not sensitive and the parameter estimates are only sensitive below approximately $T=30$~s (they are unreliable in the sense that the alternative estimation methods give parameter values outside of the error bars shown for the baseline method).
+The phases for $z_x$ and $b$ are offset by 120 degrees \todo{Why?} while the phases for $a$ and $b$ are significantly different \todo{Why?}.
 
@@
 
-Because Model~4 provided the best prediction performance, the ideal measurements to use for estimating GIC in this dataset are contemporaneous GIC and $\mathbf{B}$ measurements.
-
 \todo{TODO} {\color{red}\textbf{Discuss the origin of the frequency dependence of the system parameters.}}
@@
 Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
-GIC data provided by ...
-\todo[color=blue!40]{Shinichi: How should this be referenced?}
+\todo[color=blue!40]{Shinichi: How should this be referenced?} GIC data provided by ...
 
-\clearpage
+\bibliography{paper.bib}
+
+%\clearpage
 
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_Z-options-1.pdf}
-\caption{Frequency domain transfer functions for Models 1, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1.pdf}
+\caption{Frequency domain transfer functions for Models 1-3, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
 \label{Z}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_TF_aves_Phi-options-1.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
 \caption{Frequency domain phase values for Models 1, 2, and 4.}
@@
 
-Many alternatives to the basic least-squares for the calculation of transfer functions have been proposed and are strongly advocated by their developers who have shown that clearly problematic data will ....
-For example, as it has stated that the least-squares method often fails \citep{Egbert2011} or that failures are sometimes spectacular but more often insidious~\citep{Chave1987}.
-Several of these methods were tested individually by modifying the method described above to determine their impact on the computed transfer functions and their overall results.
+Many ``robust'' alternatives to the basic least-squares for the calculation of transfer functions have been proposed and are strongly advocated by their developers.
+For example, \citep{Egbert2011} notes that the least-squares method often fails and ~\citep{Chave1987} that failures are sometimes spectacular but more often insidious.
+~\cite{Chave2017} has noted that the model assumed for the calculation of ``robust'' transfer functions appears to be invalid because of the pervasiveness of non-gaussian residuals.
+
+%Using an exemplar data set, the regression residuals from both an ordinary least squares and robust estimator are shown to be systematically long-tailed rather than comprising a Gaussian core contaminated by outliers.
+This observation is not new and appears to be pervasive.
+
+Several alternative approaches to calculating the transfer functions in this work were tested individually to determine their impact on the computed transfer functions and their overall results.
 The primary conclusion is that different methods generally give the same results except when the signal to noise ratio is low (below approximately $30$ seconds).
@@
 \item Five methods for averaging were used to compute the magnitudes at each evaluation frequency of model parameter $p(D)$, where $D$ is estimate made using a 1-day.
-If $u(D)$ and $v(D)$ are the real and imaginary components of $p$, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation).
+If $u(D)$ and $v(D)$ are the real and imaginary phase values for Models 1-3, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation).
 The PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$.

### 2019-02-23 paper.tex
Updates from Overleaf

@@
 \begin{keypoints}
-\item The power system parameters derived from GIC and geoelectric field measurements are frequency dependent.
+\item Power system parameters derived from a set of GIC and geoelectric field measurements are frequency dependent.
 \item A GIC prediction model with frequency dependence significantly outperforms a frequency independent model.
@@
 \begin{abstract}
-A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive.
+A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a nearby direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive.
 That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.
-We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10\todo[color=blue!40]{Shinichi: Is 10 km correct?} kilometers away.
-We compute frequency dependent network parameters $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
+We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10~km \todo[color=blue!40]{Shinichi: Is 10 km correct?} away.
+We compute frequency dependent network parameters $a(\omega)$ and $b(\omega)$ in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
 \end{abstract}
@@
 Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}.
-Accurate estimation of GICs given direct measurements of the electric field (or estimates of the electric field based on the geomagnetic field) is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \todo{Pierre and Trevor - any suggestions?} {\color{red} \textbf{[REFS]}}.
+Accurate estimation of GICs given direct measurements of the electric field (or estimates of the electric field based on the geomagnetic field) is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}.
+\todo{Pierre and Trevor - any other reference suggestions?}
 
@@
 In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression, when contemporaneous measurements of GIC and $\mathbf{E}$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}.
-This quasi-DC assumption has been used or stated in many works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.
+This quasi-DC assumption has been used or stated in many works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015,Bonner2017}.
+
+\cite{Oyedokun2013a} developed a method to account for the finite inductive response time of transformers by taking the GIC predicted based on the quasi-DC assumption and then applying a piecewise-linear L/R filter derived from bench-scale transformer measurements, with a time constant dependent on load, to obtain an alternative estimate of GIC.
+
+\todo[inline]{What can I conclude from \cite{Oyedokun2013a}? Mostly the thesis documents visual differences in time series traces with and without accounting for transformers.}
 
@@
 
-The 1-second-cadence GIC data used span a total of 35 days across 10 intervals: 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22.
+The 1-second-cadence GIC data used span a total of 35 days across 10 intervals that start and end at midnight Japan Standard Time (JST): 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22.
 These data have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
 These time intervals cover those for which GIC data was made available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.
-The GIC dataset includes 1-s raw data and 1 Hz low-pass filtered raw measurements.
-The results presented here are for the 1 Hz low-pass filtered GIC values
+The GIC dataset includes 1~s raw data and 1~Hz low-pass filtered raw measurements.
+The results presented here are for the 1~Hz low-pass filtered GIC values
 
 One-second-cadence $\mathbf{E}$ and $\mathbf{B}$ measurements a from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
-Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.
 
-Data for the longest continuous time span of GIC measurements, 2006-04-02 through 2006-04-10 are shown Figure~\ref{sample}.
+Data for the time span of 2006-08-04T15:00 through 2006-08-08T15:00 are shown Figure~\ref{sample}.
+The GIC data were provided as 1-day files that start at midnight JST and the times shown in Figure~\ref{sample} are Universal Time.
 
@@
 
-The only modification made to any of the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values.
+The only modification made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values.
 The values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.
@@
 The GIC data contains non-physical spikes that are followed by a relaxation time of 40-80 s.
-These spikes were removed by identifying times $t_s$ when the change in 1-se $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
+These spikes were removed by identifying times $t_s$ when the change in 1~s $|GIC|$ measurements was greater than $0.05$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
 In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
@@
 
-To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model ~1,
+To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model~1,
 
@@
 The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ over a 1-day interval.
-\citep[][ provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}.
-The parameters $a_o$ and $b_o$ were computed for each of the 35 unique 1-day window segments of available data.
+\citep[][provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}.
+The averages $\overline{a}_o$ and $\overline{b}_o$ are the average of the 35 $a_o$ and $b_o$ values computed using the 35 unique 1-day window segments of available data.
 
-In Model 2 the system parameters are frequency dependent
+In Model 2, the system parameters $a$ and $b$ are frequency dependent according to
 
@@
 \begin{equation}
-G_E(\omega) = a(\omega)E_x(\omega) + b(\omega)E_y(\omega)\,.
+G_E(\omega) = a(\omega)E_x(\omega) + b(\omega)E_y(\omega)
 \label{model2}
@@
 \end{linenomath*}
+\noindent
+and are obtained using contemporaneous measurements of GIC and $\mathbf{E}$ using the method described later in this section.
+Note that fourier transformed variables are indicated by explicitly listing their dependence on $\omega$.
 
-The third model considered uses an electric field $\mathbf{E}'$ in Equation~\ref{model2}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$:
+In Model~3, the parameters $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)/\mu_o$:
 
@@
 \noindent
-where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives
+where, explicitly, the electric field components are $\mu_oE'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $\mu_oE'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives
 
+\begin{linenomath*}
 \begin{equation*}
-G_{E'}(\omega) = \big[a(\omega)Z_{xx}(\omega) + b(\omega)Z_{yx}(\omega) \big] B_x(\omega) + \big[ a(\omega)Z_{xy}(\omega) + b(\omega)Z_{yy}(\omega) \big] B_y(\omega)\,.
+G_{E'}(\omega) = \big[a(\omega)Z_{xx}(\omega) + b(\omega)Z_{yx}(\omega) \big] B_x(\omega)/\mu_o + \big[ a(\omega)Z_{xy}(\omega) + b(\omega)Z_{yy}(\omega) \big] B_y(\omega)/\mu_o\,.
 \end{equation*}
+\end{linenomath*}
 
 That is, instead of using the measured electric field directly as in Model 2, a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.
-In this work, $\boldsymbol{Z}$ was calculated using the method described later in this section; $a(\omega)$ and $b(\omega)$ are taken from the solution of Equation~\ref{model2}, using the method also described later in this section.
+In this work, $\boldsymbol{Z}$ was calculated using the method described later in this section.
 
@@
 \noindent
-In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of GIC and $\mathbf{B}$ and the model is expected to produced predictions of GICs that are similar to that obtained from Model 3.
+In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of GIC and $\mathbf{B}$ and the model is expected to produce predictions of GICs that are similar to that obtained from Model 3.
 
-Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e = 2\pi f_e$, by performing a least squares regression using Equation~\ref{model2} and the set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $f_e$ from the fourier transform of 24 hours of 1-second data.
+Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a least squares regression using Equation~\ref{model2} and the set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $f_e$ from the fourier transform of 24 hours of 1-s cadence data.
 The values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$; if $f_e$ value is not a multiple of $1/86400$, the nearest multiple is chosen.
@@
 
-To convert the frequency domain transfer functions $a(\omega)$ and $b(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed.
-The transfer function magnitudes $|a(\omega)|$ and $|b(\omega)|$ were calculated by linear interpolation of the transfer function magnitudes (instead of linearly interpolating the real and imaginary components and then computing the magnitude).
+To convert the frequency domain transfer functions $a(\omega)$ and $b(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $h_a(\tau)$ and $h_b(\tau)$, their real and imaginary components are linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed.
+
 The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}.
-For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$, the $xx$ and $yx$ components of $\boldsymbol{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$; the $yx$ and $yy$ components are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$.
+For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $yx$ components of $\boldsymbol{Z}$ are solved for using $E_x(\omega) = Z_{xx}(\omega)B_x(\omega) + Z_{yx}(\omega)B_{y}(\omega)$; the $yx$ and $yy$ components are solved for using $E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
 
@@
 
-There are many statistical methods that have been developed for estimating transfer functions in the magnetotelluric literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Simpson2005}.
-These works typically involve the estimation of a robust and bias-corrected transfer function.
-In the results presented, we did not use a robust statistical method or a bias correction algorithm as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters.
-As discussed in~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).
+There are many statistical methods and variations thereof that have been developed for estimating transfer functions in the magnetotelluric (MT) literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Sims1971,Simpson2005}.
+These objective of these works is typically the estimation of a robust and low bias transfer function.
+In the results presented, we did not use a robust statistical method or attempt to correct for bias in estimates as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters.
+As discussed and evaluated in~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).
 
-However, several alternative methods were tested individually by modifying the basic method described above to determine if there was a change in the computed transfer functions and, most important for this work, the out-of-sample prediction performance of each model.
-When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance is within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s).
+However, several additional processing methods and steps used in the MT literature were tested individually by modifying the basic least-squares method used to determine if there was a change in the computed transfer functions and, most relevant for this work, the out-of-sample prediction performance of each model.
+When using these additional steps, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance is within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s).
 The experiments in this regard are discussed in the Appendix.
@@
 The performance of each model in predicting GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio. 
-
 The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the predicted data that predicted by the model.
 
-The average PE and MSE ratio was computed by averaging the 35 out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the out-of-sample 35 signal-to-noise ratios at that frequency.
+The average PE and MSE ratio was computed by averaging the 35 out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the 35  out-of-sample signal-to-noise ratios at that frequency.
 The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the 35 prediction metrics.
@@
 The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1.
-Model 3, which uses magnetic field measurements as in input provides better estimates of GIC than Model 2.
+Model 3, which uses horizontal magnetic field measurements as its input provides better estimates of GIC than Model 2.
 Model 4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides the best predictions of GIC.
@@
 The error bars represent $\pm$ one standard error of the 35 SN ratios at each period.
+At each evaluation frequency the signal-to-noise ratio is is the ratio of the average of the 35 segment PSDs of GIC to the average of the 35 segment PSDs of the prediction error.
+The PSD for each of the segments at a given evaluation frequency was obtained by averaging the raw PSD using values between the next higher and lower evaluation frequency.
+
 Consistent with the results in Table~1, The SN ratio for Model 1 is lower than that for Models 2-4 at all frequencies.
-Above $1400$~s (24~min), the SN ratio ordering is consistent with the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio.
+Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio.
 Below $500$~s (8~min), Model 2 has the largest SN ratio.
@@
 
-The wide distribution and large error bars are suggestive that Model~\ref{model1} is a poor representation of the measurements.
+The wide distribution of parameter values and large error bars are suggestive that Model~\ref{model1} is a poor representation of the measurements.
 However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
-In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$  \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GICs, but is a poor approximation relative to Models 2 and 3.
-For the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.
+In a similar way that $|\mathbf{B}|(t)$ or $d|\mathbf{B}|(t)/dt$ are often good predictors of $GIC(t)$ \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GICs, but is a poor approximation relative to Models 2 and 3.
 
-The average parameters for Models 2 and 4 are shown in Figure~\ref{H}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.
+%However, for the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.
+
+The average parameters for Models 2 and 4 are shown in Figure~\ref{H}--\ref{Phi}.
+The error bars at each time step or period is the standard error of the 35 values used to compute the average.
+The transfer function magnitudes were calculated by linear interpolation of the transfer function magnitudes at the evaluation frequencies (instead of linearly interpolating the real and imaginary components and then computing the magnitude).
 
@@
 By definition, the impulse response for Model 1 is zero except at $t=0$.
-Note that the sign of the impulse response at $t=0$ is opposite for Model 1 and Model 2 for the $E_y$ coefficient - at $t=0$, $b_o < 0$ while $b(0)>0$.
+Note that the sign of the impulse response at $t=0$ is opposite for Model 1 and Model 2 for the $E_y$ coefficient - at $t=0$, $b_o < 0$ while $h_b(t=0)>0$.
+An explanation for the acausality of the impulse response is given in \cite{Kelbert2017}.
 
 Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition).
-Note that the ratio of $|\overline{b}_o/||\overline{a}_o|$ is approximately 1 for Model~1 but varies from 1.6 to 6.3 for Model~2.
-In Model~2, GIC is most dependent on $E_y$, however note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+In Model~2, GIC more sensitive to $E_y$ due to $b(\omega) > a(\omega)$, but note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+The ratio of $|\overline{b}_o/|/|\overline{a}_o|$ is approximately 1 for Model~1 and this can be compared with $|b(\omega)|/|a(\omega)|$ calculated using the lines for $|a(\omega)|$ and $|b(\omega)|$ shown in Figure~\ref{Z}.
+The ratio varies from 2.7 at $T=30$~s to 1.6 at $T=12$~h.
+At $T=12$~h, the highest period considered, the phases of $a(\omega)$ and $b(\omega)$ shown in Figure~\ref{Phi} differ by 180 degrees and the signs are consistent with $\overline{a}_o$ and $\overline{b}_o$.
 
@@
 By definition, that for Model 1 is zero.
-The phases for $z_x$ and $b$ are offset by 120 degrees \todo{Why?} while the phases for $a$ and $b$ are significantly different \todo{Why?}.
+The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120 degrees while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.
+\todo{Why?} At a period of $60$~s, the parameters in Equation~\ref{model2} are most consistent with the assumptions of frequency independent system parameters of Model~1.
+The phases of $a(\omega)$ and $b(\omega)$ are near zero and the $b(\omega)$ curve is the flattest.
+
+%\cite{Pulkkinen2010} noted that one could estimate the ratio of $a_o/b_o$ using GIC and local geomagnetic field measurements even if $\mathbf{Z}(\omega)$ was unknown if one assumed the ground conductivity was 1-D so that $Z\equiv Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$ so that Equation~\ref{model2} can be written
+
+%\begin{linenomath*}
+%\begin{equation}
+%G(\omega) = a_oZ(\omega)B_y(\omega)/\mu_o - b_oZ(\omega)B_x(\omega)/\mu_o\,.
+%\label{pulkkinen2010}
+%\end{equation}
+%\end{linenomath*}
+
+%\noindent
+%If the above equations are solved for with unknowns $Z_a\equiv a_oZ$ and $Z_b\equiv b_oZ$,  then $a_o/b_o = Z_a(\omega)/Z_b(\omega)$.
+\cite{Pulkkinen2010} assumed that this ratio was frequency independent and provided a closed-form equation for $a_o/b_o$ by multiplying Equation~\ref{pulkkinen2010} by the complex conjugates of $B_x(\omega)$ and $B_y(\omega)$ so as to form two equations from which $Z$ can be eliminated.
+With the additional assumption that the cross- and auto-spectra of all variables have the same functional form, they showed that the ratio can be written in a way that requires only evaluation of correlations among $G(t)$, $B_x(t)$, and $B_y(t)$.
 
@@
 
-Several methods have been used in the literature for estimating GIC.
-When electric field measurements are available, a model of the form of Model~1 is used \todo{TODO} {\color{red} \textbf{[REFS]}}.
-We have shown using a data set of contemporaneous GIC and $\mathbf{E}$ measurements that a significantly better estimation of GIC can be obtained by using a model of the form of Model~2.
+The primary implications of the results presented are (1) Significantly improved predictions of GICs can be made if frequency dependent system parameters are used and (2) A direct comparison of $a_o$ and $b_o$ estimated using Equation~\ref{model1} and computed using power system configuration information is problematic because the data are not consistent with the assumptions of Equation~\ref{model1}.
+Although the use of $a_o$ and $b_o$ values estimated from the data give a reasonable out-of-sample prediction model, it is not known how much these data estimated values will differ from those directly calculated.
 
-When only a MT impedance tensor that connects $\mathbf{E}$ to $\mathbf{B}$ is available, a model of the form of Model~1, with the electric field estimated using $\mathbf{E}=\mathbf{Z}\mathbf{B}$ \todo{TODO} {\color{red} \textbf{[REFS]}}.
-For the dataset considered, significantly better estimates can be made if frequency dependent system parameters are available as in Model~2.
+Also note that $\sim 10$ days of data are required to give values of $a_o$ and $b_o$ that provide the out-of-sample prediction quality reported here for Model~1 (a PE of 0.35 and CC of 0.64).
+If only a few days of data are used for their estimation, the out-of-sample prediction quality is very poor due to the high degree of uncertainty of in their estimates.
+Predictions of GIC using Model~1 in the literature using only a few days of data are expected to be significantly worse if out-of-sample data are used.
 
-When only magnetic field data is available, a model of the form of Model~4 is used.
+Several methods have been used in the literature for estimating or predicting GIC depending on the data or models that are available.
+
+When electric field measurements are available, a \todo{TODO} {\color{red} \textbf{[REFS]}} model of the form of Model~1, with frequency independent system parameters derived from power system configuration, has been used.
+These works generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
+However, we have shown using a data set of contemporaneous GIC and $\mathbf{E}$ measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
+In addition, we have found that the estimates of the system parameters $a_o$ and $b_o$ are highly sensitive to the interval of data used, and have a high degree of uncertainty.
+
+When only a MT impedance tensor that connects $\mathbf{E}$ to $\mathbf{B}$ is available, a model of the form of Model~1, with the electric field estimated using $\mathbf{E}=\mathbf{Z}\mathbf{B}$ has been used~\citep{Weigel2017,Bonner2017,Kelbert2017}.
+For the dataset considered, significantly better GIC estimates can be made if frequency dependent system parameters are available as in Model~3.
+
+Ideally, frequency dependent system parameters could be calculated analytically so that GIC measurements are not needed for their estimates.
+Because the details of the power system under consideration in this work are not available, we were not able make this comparison.
+Note that frequency independent system parameters calculated using the approach in this work may not be directly comparable to those found from a network analysis for several reasons: (1) At periods less than $\sim 30$~s, the signal-to-noise ratio is low and the parameter estimates are expected to be biased; as shown in the Appendix, the transfer function estimates are highly sensitive to the method used for their calculation at (2) Also at periods less than $\sim 30$~s, effects of the instrument response may be significant.
+\cite{Fujii2015} has noted that the 1-s geoelectric field data from the Memambetsu Observatory were low-pass filtered with a 3.3-s cutoff-period prior to 2010; (3) The GIC observed in a power system is based on the integration of the electric field along a power line.
+All of the models assume that the electric field is constant along the line where the GIC is measured, but many works have shown that the electric field is expected to have variations on length scales comparable to the line considered ($\sim$100~km).
+\todo{REFS}
+
+When GIC and magnetic field measurements are available, a model of the form of Model~4 can be used.
 In this case, the transfer function contains components of the transfer function that connects $\mathbf{E}$ to $\mathbf{B}$ and GIC to $\mathbf{E}$.
+For this particular dataset, we have found that such a model provides the best estimates of GIC.
+A power system operator could use local ground magnetic field estimates along with GIC measurements to derive a transfer function model that is updated when the system configuration changes, and this model could be used to estimate GIC under extreme or simulated geomagnetic conditions.
+A key advantage of this approach is that geomagnetic field measurements are generally easier to make in the long-term and are more readily available.
 
-\todo{TODO} {\color{red}\textbf{Discuss the origin of the frequency dependence of the system parameters.}}
+Based on the results in this work, the model of the form $GIC(t)$ $=a_oE_x(t)+b_oE_y(t)$, where the electric field is either measured or estimated using a transfer function, should be regarded as an approximation similar to that assumed when an equation of the form $GIC(t)$ $=c_oB_x(t)+d_oB_y(t)$ or $GIC(t)$ $=c_odB_x(t)/dt+d_odB_y(t)/dt$ is used, where $a_o, b_o, c_o, $and $d_o$ are constants (e.g., {\color{red} \textbf{[REFS]}}).
+
+%~\ref{pulkkinen2010} used a variation of this approach, but used an assumed conductivity model to estimate the frequency independent system parameters.
+
+%The frequency dependence of $a(\omega)$ and $b(\omega)$ shown in Figure~\ref{model2} is not consistent with a simple model of an ideal resistor $R$ in series with an ideal inductor $L$, for which the transfer function magnitude is $R^{-1}(1+\omega^2\tau^2)^{-1}$ and phase is $\tan^{-1}\omega\tau$, where $\tau=L/R$.
+
+\todo{All - Need help in discussing the origin of the frequency dependence of the system parameters.}
 
@@
 \todo[color=blue!40]{Shinichi: How should this be referenced?} GIC data provided by ...
-
-\bibliography{paper.bib}
 
@@
 (Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
-The subscripts $x$, $y$, and $z$ correspond to North, East, and Downward, respectively.}
+The subscripts $x$ and $y$ correspond to Geographic North and East, respectively.}
 \label{sample}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_SN-options-1.pdf}
-\caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
+\caption{Out-of-sample signal-to-noise ratios at the evaluation frequencies for the four models.}
 \label{SN}
@@
 \caption{Distribution of parameters in Model 1.
-Each set of parameters were computed 35 times using one day of 1-second cadence data.
+Each set of parameters was computed 35 times using one day of 1-second cadence data.
 The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1.pdf}
-\caption{Frequency domain transfer functions for Models 1-3, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
+\caption{Frequency domain transfer functions for Models 1-3, 2, and 4. The frequency domain transfer functions for $a_o$ and $b_o$ in Model~1 are constant and equal to $a_o$ and $b_o$.}
 \label{Z}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
-\caption{Frequency domain phase values for Models 1, 2, and 4.}
+\caption{Frequency domain phase values for Models 1, 2, and 4. The phase for $a_o$ and $b_o$ in Model~1 are constant and equal to zero.}
 \label{Phi}
@@
 
-Many ``robust'' alternatives to the basic least-squares for the calculation of transfer functions have been proposed and are strongly advocated by their developers.
-For example, \citep{Egbert2011} notes that the least-squares method often fails and ~\citep{Chave1987} that failures are sometimes spectacular but more often insidious.
-~\cite{Chave2017} has noted that the model assumed for the calculation of ``robust'' transfer functions appears to be invalid because of the pervasiveness of non-gaussian residuals.
+%Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.
 
-%Using an exemplar data set, the regression residuals from both an ordinary least squares and robust estimator are shown to be systematically long-tailed rather than comprising a Gaussian core contaminated by outliers.
-This observation is not new and appears to be pervasive.
+In the body of the work, a basic ordinary least squares method was used to compute transfer functions.
+Of course, there are many problems that one must address before the parameter estimates from a linear regression model can be used for a physical interpretation because the estimates may be biased, have high uncertainty, and be sensitive to outliers.
+In addition, to make inferences about ground conductivity, corrections must be made for distortion effects and violations of the plane wave assumption \citep{Simpson2005}.
 
-Several alternative approaches to calculating the transfer functions in this work were tested individually to determine their impact on the computed transfer functions and their overall results.
-The primary conclusion is that different methods generally give the same results except when the signal to noise ratio is low (below approximately $30$ seconds).
-Although plotted because they were a part of the prediction models and they do contribute somewhat to the prediction efficiency (as discussed in \cite{Weigel2017}, below $30$ seconds, the model parameter estimates should be regarded as highly uncertain due to large error bars and a large sensitivity to the method used to calculate them.
+Many robust alternatives to the basic least-squares for the calculation of transfer functions have been proposed and are strongly advocated by their developers to reduce uncertainty and sensitivity to outliers.
+For example, \cite{Egbert2011} notes that the least-squares method often fails and \cite{Chave1987} that failures are sometimes spectacular but more often insidious.
+However, ~\cite{Chave2017} has noted that the commonly used robust regression models are also problematic because the error model assumed for the calculation of robust transfer functions is often invalid due to the pervasiveness of non-gaussian residuals, and as a result it is not unconditionally unbiased.
 
-\begin{itemize}
-\item Using robust regression instead of linear regression when computing the transfer function estimates at each $\omega_e$ (using a custom implementation with the $\psi$-function referenced in the next item and also using the \texttt{robustfit} function in MATLAB with bi-square and Huber weighting options).
-The PEs and MSEs shown in Table~1 were the same to two decimal places and the confidence intervals were smaller by $\sim 2$\% (when the confidence interval was calculated assuming a Gaussian distribution instead of using the bootstrap, the confidence intervals varied by $\sim\pm 2$\%).
+There are also many approaches that are taken in pre-processing the data before regression is performed, for example, windowing in the time domain prior to the calculation of the fourier transform, windowing in the frequency domain to down-weight values far from the evaluation frequency or to down-weight frequencies that have low coherence, weighting values by an empirically estimated variance, pre-whitening, and using multi-taper methods to estimate spectra.
 
-\item Five methods for averaging were used to compute the magnitudes at each evaluation frequency of model parameter $p(D)$, where $D$ is estimate made using a 1-day.
-If $u(D)$ and $v(D)$ are the real and imaginary phase values for Models 1-3, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation).
-The PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$.
+In general, the approach taken in the MT literature is to select a given variant of least squares and a set of pre-processing steps and to report the transfer functions along with an uncertainty (several approaches to estimating uncertainty exist as well).
+Absent an understanding of which particular regression alternative to use, the appropriate pre-processing approach, and the appropriate uncertainty estimate method, and an understanding of the appropriate adjustable parameters of each option and how they interact, the approach taken here is to investigate the influence of a few of the common regression alternatives, pre-processing steps, and stack-averaging methods on the transfer function estimate.
+
+The primary conclusion is that different methods generally give the statistically similar results in both the prediction metrics and parameter estimates except when the signal to noise ratio is low (below approximately $30$ seconds).
+Although parameter estimates were plotted below $30$ seconds because they were a part of the prediction models and they do contribute somewhat to the prediction efficiency, the estimates should be regarded as highly uncertain due to large error bars and a large sensitivity to the method used to calculate them.
+
+The various tests that were performed to determine the influence of each method are summarized below and the associated results are shown in Figures~\ref{a_variations}-~\ref{b_variations}.
+
+\begin{enumerate}
+
+\item Robust regression was used instead of linear regression when computing the transfer function estimates at each $\omega_e$ (using a custom implementation with the $\psi$-function referenced in the next item and also using the \texttt{robustfit} function in MATLAB with either the bi-square and Huber weighting options).
+The motivation for the use of robust regression is to reduce the influence of outliers on the parameter estimates; ordinary least squares parameter estimates have a high sensitivity to outliers.
+When using robust regression, PEs and MSEs shown in Table~1 were the same to two decimal places and the confidence intervals were smaller by $\sim 2$\%.
+Figures~\ref{a_variations}-~\ref{b_variations} shows that the robust and used method give near identical estimates to the least square regression estimates.
+
+\item Each 1-day segment was pre-whitened prior to calculation of the transfer functions in order to reduce the spectral range and to remove long-period drifts.
+This was done by computing the pre-whitening filter using a 10th-order Yule-Walker filter using $E_y$ for Model~2 and $B_x$ for Model~4 and applying this filter to all time series used in the calculation of the transfer function.
+The out-of-sample PEs were the same to within two decimal places and the MSE ratios varied by $0.05$.
+Figures~\ref{a_variations}-~\ref{b_variations} show that pre-whitening has a larger impact on the estimates than robust regression.
+
+\item A Parzen window was used to weight the fourier-transformed values instead of a rectangular window.
+This reduces the influence of values used in the transfer function parameter estimate as a function of distance from the evaluation frequency.
+The out-of-sample PEs for Model 3 and Model 4 were higher by $0.02$ and $0.01$, respectively and the MSE ratios for Models 3 and 4 were higher by $0.4$ and $0.6$, respectively.
+Figures~\ref{a_variations}-~\ref{b_variations} show that the pre-whitening has about the same amount of influence as the use of a Parzen window.
+
+\item Five methods for averaging were used to compute magnitudes at each evaluation frequency of a complex-valued model parameter $p(D)$, where $D$ is estimate made using a 1-day of data.
+If $u(D)$ and $v(D)$ are the real and imaginary phase values for Models 2-4, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation).
+The out-of-sample PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$.
 Neither averaging method was yielded consistently higher or lower values.
+Figures~\ref{a_variations}-~\ref{b_variations} shows that most of the sensitivity is below $T=20$ s, as expected, due to the signal-to-noise being near equal to 1 below this period.
 
-\item Pre-whitening each 1-day segment prior to calculation of the transfer functions.
-This was done by computing the pre-whitening filter using a 10th-order Yule-Walker filter computed using $B_x$ and applying this filter to all time series used in the calculation of the transfer function.
-The PEs were the same to within two decimal places and the MSE ratios varied by $0.05$.
+\end{enumerate}
 
-\item Windowing the signals in the frequency domain using a rectangular window (instead of Parzen).
-The PEs for Model 3 and Model 4 were higher by $0.02$ and $0.01$, respectively and the MSE ratios for Models 3 and 4 were higher by $0.4$ and $0.6$, respectively.
+Several addition experiments were performed that we summarize with text only.
 
-\item Using weighted averages of the 35 transfer functions with the weighting at each frequency equal to the in-sample coherence at that frequency (to give higher weight to better fitting intervals) and weighting the collection of values used to compute the model parameters at a given evaluation frequency by a smoothed estimate of the PSD of the predictand (to account for the fact that the signal to noise ratio depends on frequency as evident in ~\ref{SN} and noted by \citep{Egbert1997}).
+\begin{enumerate}
+\item Instead computing an average transfer function by a simple average of the 35 transfer functions, weighted averages were considered.
+The weighting included using the in-sample coherence or prediction efficiency (ad-hoc approaches to give higher weight to better fitting intervals) and weighting the collection of values used to compute the model parameters at a given evaluation frequency by a smoothed estimate of the PSD of the predictand (to account for the fact that the signal to noise ratio depends on frequency as evident in ~\ref{SN} and noted by \citep{Egbert1997}; accounting for this can reduce the variance in the parameter estimates).
+An approach that is more similar to that used in \citep{Egbert1997} was also used - prior to regression, the the values were weighted with a weight that is inversely proportional to an estimate of of the spectra of the vector magnitude of the input.
+
+\item At a given frequency, we have found that there is a correlation between the parameter estimate and the PSD of the input, output, and prediction error that varies between +0.6 and -0.2, which is one of the many ways in which the data violate the assumptions for OLS estimates to be the best linear unbiased estimator.
+To determine how this violation could effect the parameter estimates, at each frequency, the top 17 and lowest 18 transfer functions were averaged and plotted as a function of frequency.
+The parameter estimates from the high input PSD intervals tended to be higher, but within the range of the displayed error bars.
 
 \item Instead of calculating 35 transfer functions and then averaging, compute one transfer function - for each evaluation frequency, combine the 35 fourier spectra values near that evaluation frequency and use them for one regression.
-\end{itemize}
+\end{enumerate}
 
-In all cases, the parameter estimates below $\sim 20$ seconds using the alternative methods were consistently lower, but the offset was of magnitude that was similar to the variation observed between the alternative methods.
+%In all cases, the parameter estimates below $\sim 20$ seconds using the alternative methods were consistently lower, but the offset was of magnitude that was similar to the variation observed between the alternative methods.
 The largest variations were found in relative to the method used to averaging the 35 transfer functions.
@@
 The same was true for the median.
+
+\clearpage
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_options_comparison_a.pdf}
+\caption{The lower group of lines represents the estimate of $a(\omega)$ made using different pre-processing or regression methods.
+The upper group of lines has been shifted upwards by a factor of 10 and represents estimates made using different averaging methods.
+The 1 standard error error bars are only shown for the lines associated with the method used in the body of the text (labeled as ``Default'').}
+\label{a_variations}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_options_comparison_a.pdf}
+\caption{The lower group of lines represents the estimate of $a(\omega)$ made using different pre-processing or regression methods.
+The upper group of lines has been shifted upwards by a factor of 10 and represents estimates made using different averaging methods.
+The 1 standard error error bars are only shown for the lines associated with the method used in the body of the text (labeled as ``Default'').}
+\label{b_variations}
+\end{figure}
 

### 2019-02-23 figures/update.sh
update figs

@@ -8,6 +8,10 @@ cp $DIR/plot_model_summary_H-options-1.pdf .
 cp $DIR/plot_model_summary_Phi-options-1.pdf .
 cp $DIR/plot_model_summary_Z-options-1.pdf .
 cp $DIR/plot_model_summary_SN-options-1.pdf .
+cp $DIR/plot_options_comparison_a.pdf .
+cp $DIR/plot_options_comparison_b.pdf .
+cp $DIR/plot_options_comparison_zx.pdf .
+cp $DIR/plot_options_comparison_zy.pdf .
 
 cp $DIR/../20060805/plot_raw_All_20060805.pdf .
 git commit -a -m "update figs"

### 2019-02-23 paper.tex
Updates from Overleaf

@@
 \begin{figure}[h]
-\centering
+\begin{minipage}[t]{.48\linewidth}
 \includegraphics[width=\textwidth]{figures/plot_options_comparison_a.pdf}
-\caption{The lower group of lines represents the estimate of $a(\omega)$ made using different pre-processing or regression methods.
-The upper group of lines has been shifted upwards by a factor of 10 and represents estimates made using different averaging methods.
-The 1 standard error error bars are only shown for the lines associated with the method used in the body of the text (labeled as ``Default'').}
-\label{a_variations}
+\label{plot_options_comparison_a}
+\end{minipage}
+\begin{minipage}[t]{.48\linewidth}
+\includegraphics[width=\textwidth]{figures/plot_options_comparison_b.pdf}
+\end{minipage}
+\begin{minipage}[t]{.48\linewidth}
+\includegraphics[width=\textwidth]{figures/plot_options_comparison_zx.pdf}
+\label{plot_options_comparison_a}
+\end{minipage}
+\begin{minipage}[t]{.48\linewidth}
+\includegraphics[width=\textwidth]{figures/plot_options_comparison_zy.pdf}
+\end{minipage}
+
 \end{figure}
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{figures/plot_options_comparison_a.pdf}
-\caption{The lower group of lines represents the estimate of $a(\omega)$ made using different pre-processing or regression methods.
-The upper group of lines has been shifted upwards by a factor of 10 and represents estimates made using different averaging methods.
-The 1 standard error error bars are only shown for the lines associated with the method used in the body of the text (labeled as ``Default'').}
-\label{b_variations}
-\end{figure}
+\clearpage
 

### 2019-02-26 paper.tex
Updates from Overleaf

@@
 That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.
-We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10~km \todo[color=blue!40]{Shinichi: Is 10 km correct?} away.
+We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.
 We compute frequency dependent network parameters $a(\omega)$ and $b(\omega)$ in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
@@
 Accurate estimation of GICs given direct measurements of the electric field (or estimates of the electric field based on the geomagnetic field) is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}.
-\todo{Pierre and Trevor - any other reference suggestions?}
 
@@
 
-\cite{Oyedokun2013a} developed a method to account for the finite inductive response time of transformers by taking the GIC predicted based on the quasi-DC assumption and then applying a piecewise-linear L/R filter derived from bench-scale transformer measurements, with a time constant dependent on load, to obtain an alternative estimate of GIC.
+%\cite{Oyedokun2013a} developed a method to account for the finite inductive response time of transformers by taking the GIC predicted based on the quasi-DC assumption and then applying a piecewise-linear L/R filter derived from bench-scale transformer measurements, with a time constant dependent on load, to obtain an alternative estimate of GIC.
 
-\todo[inline]{What can I conclude from \cite{Oyedokun2013a}? Mostly the thesis documents visual differences in time series traces with and without accounting for transformers.}
+%\todo[inline]{What can I conclude from \cite{Oyedokun2013a}? Mostly the thesis documents visual differences in time series traces with and without accounting for transformers.}
 
 In this work, we use a unique dataset in which measurements $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$, were made near a site where GICs were being measured in an electric power system.
-We first estimate the system parameters $a_o$ and $b_o$ using conventional methods, which assume that they are frequency independent.
+We first estimate the system parameters $a_o$ and $b_o$ from data using conventional methods, which assume that they are frequency independent.
 Then, we use a model in which the system parameters are frequency dependent, compute the frequency dependence, and compare this model with the traditional frequency independent model.
@@
 The only modification made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values.
-The values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.
+The 36 values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.
 
@@
 \noindent
-the matrix equation $\mathbf{GIC} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $\mathbf{GIC}$ is a 864000x1 matrix of 1 day of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ observations.
+the matrix equation $\mathbf{GIC} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $\mathbf{GIC}$ is a 864000x1 matrix of 1 day of observations with zero mean, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ observations with each column having zero mean.
 The values of $\mathbf{p}$ are then obtained using a standard linear regression solver.
@@
 
-In Model 2, the system parameters $a$ and $b$ are frequency dependent according to
+In Model 2, the system parameters $a(\omega)$ and $b(\omega)$ are frequency dependent according to
 
@@
 The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}.
-For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $yx$ components of $\boldsymbol{Z}$ are solved for using $E_x(\omega) = Z_{xx}(\omega)B_x(\omega) + Z_{yx}(\omega)B_{y}(\omega)$; the $yx$ and $yy$ components are solved for using $E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
+For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ are solved for using $E_x(\omega) = Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_{y}(\omega)$; the $yx$ and $yy$ components are solved for using $E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
 
@@
 
-There are many statistical methods and variations thereof that have been developed for estimating transfer functions in the magnetotelluric (MT) literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Sims1971,Simpson2005}.
-These objective of these works is typically the estimation of a robust and low bias transfer function.
+There are many statistical methods and variations thereof that have been developed for estimating transfer functions in the magnetotelluric (MT) literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Sims1971} and \cite{Simpson2005}.
+The objective of these works is typically the estimation of a robust and low bias transfer function.
 In the results presented, we did not use a robust statistical method or attempt to correct for bias in estimates as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters.
@@
 1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
-2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
-3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
+2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
+3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
 4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
@@
 
-Figure~\ref{predictions} shows a 1-day interval for which the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the lowest.
+Figure~\ref{predictions} shows a 1-day interval for selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the lowest.
 
 The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1.
-Model 3, which uses horizontal magnetic field measurements as its input provides better estimates of GIC than Model 2.
+Model 3, which uses an estimate of the electric field based on horizontal magnetic field measurements provides better estimates of GIC than Model 2.
 Model 4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides the best predictions of GIC.
+
+Although inferior, the Model~1 does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
+The top panel of  Figure~\ref{predictions} shows that its predicted GIC generally follows the overall trend of the measured GIC.
+
+%In a similar way that $|\mathbf{B}|(t)$ or $d|\mathbf{B}|(t)/dt$ are often good predictors of $GIC(t)$ \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GICs, but is a poor approximation relative to Models 2 and 3.
 
@@
 Consistent with the results in Table~1, The SN ratio for Model 1 is lower than that for Models 2-4 at all frequencies.
-Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio.
-Below $500$~s (8~min), Model 2 has the largest SN ratio.
+
+Below $600$~s (10~min), Model 2 has the largest SN ratio.
 This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.
 
-The average parameters for Model 1 are $a_o = $ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~\ref{histogram}.
-(Results are similar if 1-minute averages are used; \cite{Watari2015} found \todo[color=blue!40]{Shinichi: I get $a_o=45.8$ and $b_o=-59.9$ in this time interval.} $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).
+Below $600$~s (10~min), the predictive ability of each model is in the ordering that may be expected based on the data used to derive the model: Model 2, which uses direct measurements of the electric field, has a higher SN ratio than Model 3, which uses an indirect estimate of the electric field.
+Model 4 has a slightly higher SN than Model 3, and this is expected because they both use the same input of the magnetic field, but Model 3 has an additional source of uncertainty from the parameters in $\mathbf{Z}$.
+Above $600$~s, the ordering of the SN is not consistent with this expectation and is possibly explained by two factors (1) the measured GIC is based on an integrated electric field and the electric field used in Model 2 is a point measurement - the result is that implicit in the model is an assumption that the electric field is uniform over the length of the line on which GIC is measured and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
 
-The wide distribution of parameter values and large error bars are suggestive that Model~\ref{model1} is a poor representation of the measurements.
-However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
-In a similar way that $|\mathbf{B}|(t)$ or $d|\mathbf{B}|(t)/dt$ are often good predictors of $GIC(t)$ \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GICs, but is a poor approximation relative to Models 2 and 3.
+Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio.
+
+The average parameters and 95\% CI for Model 1 are $a_o = $ $73 [62,64]$ A/(V/km) and $b_o =$ $-72.0 \pm [-98,-45]$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~\ref{histogram}.
+(Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).
 
@@
 
-\section{Discussion and Conclusions}
+\section{Discussion}
+
+There are at least three possible explanations of the frequency dependence of the data-derived system parameters: (1) the quasi-DC approximation is invalid, (2) invalid model assumptions, and (3) artifacts of the measurement process.
+
+(1) When the quasi-DC approximation is used, ~\cite{Albertson1981} is usually cited; the justification for the quasi-DC approximation is given the Appendix, where it is stated that ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects.
+The correction factor varies from 0.95 to 1.0 depending on conductor size.'' \cite{Lehtinen1985} notes justifies the use of the quasi-DC approximation because GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency.
+Carson's equations are equations for the frequency-dependent impedance associated with a wire above a conducting plane.
+The justification of \cite{Lehtinen1985} is essentially a statement that the frequency dependent component of Carson's impedance is proportional to frequency, the frequency dependent impedance of a power line is small at 50~Hz and so it should be even smaller at GIC frequencies of < 1~Hz.
+\todo{Need a reference here to 50~Hz values} Thus it appears that the quasi-DC approximation is justified.
+
+(2) The electric field was measured at a single location $\sim$9 km away from the GIC instrument.
+The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the earth potential at the grounding points of the system, and simulations of GIC flow \citep{Albertson1981,Lehtinen1985} typically start with an assumed electric field along the line to calculate these potential differences.
+The potential differences are then used along with the system component resistances and connectivity to calculate GIC in the different parts of the system.
+In both Model~1 and Model~2, the input electric field is assumed to be spatially homogeneous along the line in which GIC was measured.
+If it is not, the parameters derived from the data in Model~1 may vary significantly from that expected.
+The same should be true of Model~2, but this modeling error is not expected to introduce a frequency dependence.
+
+(3) A final consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence.
+Given the response characteristics of these instruments are essentially frequency independent above a period of tens of seconds, this does not explain the frequency dependence observed over the range of 10-10,000~s.
+
+\section{Conclusions}
 
 The primary implications of the results presented are (1) Significantly improved predictions of GICs can be made if frequency dependent system parameters are used and (2) A direct comparison of $a_o$ and $b_o$ estimated using Equation~\ref{model1} and computed using power system configuration information is problematic because the data are not consistent with the assumptions of Equation~\ref{model1}.
-Although the use of $a_o$ and $b_o$ values estimated from the data give a reasonable out-of-sample prediction model, it is not known how much these data estimated values will differ from those directly calculated.
+
+Although the use of $a_o$ and $b_o$ values estimated from the data give a reasonable out-of-sample prediction model, it is not known how much these data-estimated values will differ from those directly calculated.
+(Because the details of the power system under consideration in this work are not available, we were not able make this comparison.)
 
@@
 
-Ideally, frequency dependent system parameters could be calculated analytically so that GIC measurements are not needed for their estimates.
-Because the details of the power system under consideration in this work are not available, we were not able make this comparison.
-Note that frequency independent system parameters calculated using the approach in this work may not be directly comparable to those found from a network analysis for several reasons: (1) At periods less than $\sim 30$~s, the signal-to-noise ratio is low and the parameter estimates are expected to be biased; as shown in the Appendix, the transfer function estimates are highly sensitive to the method used for their calculation at (2) Also at periods less than $\sim 30$~s, effects of the instrument response may be significant.
-\cite{Fujii2015} has noted that the 1-s geoelectric field data from the Memambetsu Observatory were low-pass filtered with a 3.3-s cutoff-period prior to 2010; (3) The GIC observed in a power system is based on the integration of the electric field along a power line.
-All of the models assume that the electric field is constant along the line where the GIC is measured, but many works have shown that the electric field is expected to have variations on length scales comparable to the line considered ($\sim$100~km).
-\todo{REFS}
-
 When GIC and magnetic field measurements are available, a model of the form of Model~4 can be used.
@@
 
+%Ideally, frequency dependent system parameters could be calculated analytically for comparison with those estimated using data in this work.
+Because the details of the power system under consideration in this work are not available, we were not able make this comparison.
+Note that frequency independent system parameters calculated using the approach in this work may not be directly comparable to those found from a network analysis for several reasons: (1) At periods less than $\sim 30$~s, the signal-to-noise ratio is low and the parameter estimates are expected to be biased; as shown in the Appendix, the transfer function estimates are highly sensitive to the method used for their calculation at (2) Also at periods less than $\sim 30$~s, effects of the instrument response may be significant.
+\cite{Fujii2015} has noted that the 1-s geoelectric field data from the Memambetsu Observatory were low-pass filtered with a 3.3-s cutoff-period prior to 2010; (3) The GIC observed in a power system is based on the integration of the electric field along a power line.
+All of the models assume that the electric field is constant along the line where the GIC is measured, but many works have shown that the electric field is expected to have variations on length scales comparable to the line considered ($\sim$100~km).
+\todo{REFS}
+
 \acknowledgments
 Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
-\todo[color=blue!40]{Shinichi: How should this be referenced?} GIC data provided by ...
-
+We would like to thank the Hokkaido Electric Power Co.
+for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the data.
 %\clearpage
@@
 \caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
-(Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$10 km from the Memanbetsu substation.
+(Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$9 km from the Memanbetsu substation.
 (Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-20060806T150000.pdf}
+\includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-2006-08-06.pdf}
 \caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval.
-Note that the scales for Models 2-4 differ from that for Model 1 so that the error amplitudes for Models 2-4 appear larger than that of Model 1 by a factor of 2.}
+Note that the scales for Models~2-4 differ from that for Model 1 so that the error amplitudes for Models~2-4 are scaled by a factor of 2 relative to that for Model~1.}
 \label{predictions}
@@
 
-The various tests that were performed to determine the influence of each method are summarized below and the associated results are shown in Figures~\ref{a_variations}-~\ref{b_variations}.
+The various tests that were performed to determine the influence of each method are summarized below and the associated results are shown in Figure~\ref{model_2_and_4_variations}.
 
@@
 Neither averaging method was yielded consistently higher or lower values.
-Figures~\ref{a_variations}-~\ref{b_variations} shows that most of the sensitivity is below $T=20$ s, as expected, due to the signal-to-noise being near equal to 1 below this period.
+Figure~\ref{model_2_and_4_variations} shows that most of the sensitivity is below $T=20$ s, as expected, due to the signal-to-noise being near equal to 1 below this period.
 
@@
 \end{minipage}
-
+\caption{Caption}
+\label{model_2_and_4_variations}
 \end{figure}

### 2019-02-26 figures/update.sh
update figs

@@ -1,6 +1,6 @@
 git pull
 
-cp ../../figures/20060805/plot_model_predictions-MeanModel-20060806T150000.pdf .
+cp ../../figures/predictions/plot_model_predictions-MeanModel-2006-08-06.pdf .
 
 DIR=../../figures/combined
 cp $DIR/plot_model_summary_aobo_histograms-options-1.pdf .

### 2019-03-07 paper.tex
Updates from Overleaf

@@
 
-\title{A Test of the Frequency Independence Assumption of the Power System Parameters used in Geomagnetically Induced Current Calculations}
+\title{A Test of the Frequency Independence Assumption of Power System Parameters used in Geomagnetically Induced Current Estimates}
 
@@
 \begin{keypoints}
-\item Power system parameters derived from a set of GIC and geoelectric field measurements are frequency dependent.
+\item Power system parameters derived from a time series of GIC and geoelectric field measurements are frequency dependent.
 \item A GIC prediction model with frequency dependence significantly outperforms a frequency independent model.
@@
 We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.
-We compute frequency dependent network parameters $a(\omega)$ and $b(\omega)$ in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
+We compute frequency dependent network parameters $a(\omega)$ and $b(\omega)$ from data in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better estimates of the measured GIC.
+A third model, which uses the electric field estimated from a transfer function convolved with the geomagnetic field, provides significantly better estimates to Model~2.
+It is suggested that the measurement-derived frequency dependence of the system parameters is explained by spatial variations in the spectra of the geoelectric field rather than by an actual frequency dependence of the system parameters.
 \end{abstract}
@@
 
-Geomagnetically induced currents (GICs) are electric currents in conducting systems due to electric fields induced near Earth's surface.
-These fields are a result of time variations in ionospheric currents on timescales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
+Geomagnetically induced currents (GICs) are electric currents in conducting systems due to electric fields near Earth's surface.
+Time-varying electric and magnetic fields at Earth's surface are driven by time variations in ionospheric currents on time scales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
 Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}.
-Accurate estimation of GICs given direct measurements of the electric field (or estimates of the electric field based on the geomagnetic field) is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}.
+Accurate estimation of GICs magnitudes is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}.
 
@@
 In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression, when contemporaneous measurements of GIC and $\mathbf{E}$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}.
-This quasi-DC assumption has been used or stated in many works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015,Bonner2017}.
+This quasi-DC assumption has been used or stated in many GIC-related works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015,Bonner2017}.
 
@@
 We first estimate the system parameters $a_o$ and $b_o$ from data using conventional methods, which assume that they are frequency independent.
-Then, we use a model in which the system parameters are frequency dependent, compute the frequency dependence, and compare this model with the traditional frequency independent model.
+Then, we use a model in which the system parameters are allowed to be frequency dependent, compute the frequency dependence, and compare this model with the traditional frequency independent model.
 
@@
 
-The 1-second-cadence GIC data used span a total of 35 days across 10 intervals that start and end at midnight Japan Standard Time (JST): 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22.
+The 1-second-cadence GIC data used span a total of 35 days across 10 continuous intervals that start and end at midnight Japan Standard Time (JST): 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22.
 These data have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
-These time intervals cover those for which GIC data was made available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.
-The GIC dataset includes 1~s raw data and 1~Hz low-pass filtered raw measurements.
-The results presented here are for the 1~Hz low-pass filtered GIC values
+These time intervals are those for which GIC data was made available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.
+The GIC dataset includes 1~s raw data and 1~Hz low-pass-filtered raw measurements.
+The results presented here are for the 1~Hz low-pass-filtered GIC values
 
-One-second-cadence $\mathbf{E}$ and $\mathbf{B}$ measurements a from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
+One-second-cadence $\mathbf{E}$ and $\mathbf{B}$ measurements from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
 
@@
 
-Spikes in the electric field measurements were assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
+Spikes in the electric field measurements were assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1~mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
 In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$.
@@
 
-The only modification made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values.
+The only modification made to the magnetic field measurements was the replacement of 36 data points in all components on 2006-08-06 with linearly interpolated values.
 The 36 values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.
@@
 \noindent
-the matrix equation $\mathbf{GIC} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $\mathbf{GIC}$ is a 864000x1 matrix of 1 day of observations with zero mean, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ observations with each column having zero mean.
-The values of $\mathbf{p}$ are then obtained using a standard linear regression solver.
+the matrix equation $\underline{\mathbf{GIC}} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $\underline{\mathbf{GIC}}$ is a zero mean 864000x1 matrix of 1 day observations with rows of $GIC(t)$; $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ observations with each column having zero mean.
+The values of $\mathbf{p}$ are then obtained using linear regression.
 The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ over a 1-day interval.
 \citep[][provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}.
-The averages $\overline{a}_o$ and $\overline{b}_o$ are the average of the 35 $a_o$ and $b_o$ values computed using the 35 unique 1-day window segments of available data.
+The notation $\overline{a}_o$ and $\overline{b}_o$ is used for the average of the $a_o$ and $b_o$ values computed using the 35 unique 1-day window segments of available data.
 
@@
 \noindent
-and are obtained using contemporaneous measurements of GIC and $\mathbf{E}$ using the method described later in this section.
+and are computed using contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ using the method described later in this section.
 Note that fourier transformed variables are indicated by explicitly listing their dependence on $\omega$.
 
-In Model~3, the parameters $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)/\mu_o$:
+In Model~3, the parameters $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$:
 
@@
 \noindent
-where, explicitly, the electric field components are $\mu_oE'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $\mu_oE'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives
-
-\begin{linenomath*}
-\begin{equation*}
-G_{E'}(\omega) = \big[a(\omega)Z_{xx}(\omega) + b(\omega)Z_{yx}(\omega) \big] B_x(\omega)/\mu_o + \big[ a(\omega)Z_{xy}(\omega) + b(\omega)Z_{yy}(\omega) \big] B_y(\omega)/\mu_o\,.
-\end{equation*}
-\end{linenomath*}
-
+where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. 
 That is, instead of using the measured electric field directly as in Model 2, a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.
 In this work, $\boldsymbol{Z}$ was calculated using the method described later in this section.
+
+%Substitution gives
+
+%\begin{linenomath*}
+%\begin{equation*}
+%G_{E'}(\omega) = %\big[a(\omega)Z_{xx}(\omega) + b(\omega)Z_{yx}(\omega) \big] B_x(\omega) + \big[ a(\omega)Z_{xy}(\omega) + b(\omega)Z_{yy}(\omega) \big] B_y(\omega)\,.
+%\end{equation*}
+%\end{linenomath*}
 
@@
 \noindent
-In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of GIC and $\mathbf{B}$ and the model is expected to produce predictions of GICs that are similar to that obtained from Model 3.
+In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$, and the model is expected to produce predictions of $GIC(t)$ that are similar to that obtained from Model 3 because they both use the same input variable.
 
@@
 The values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$; if $f_e$ value is not a multiple of $1/86400$, the nearest multiple is chosen.
-The regression for $f_e$ uses all $f$ values between the next higher and lower value of $f_e$.
+The regression for $f_e$ uses all $f$ values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$.
 
@@
 
-Transfer function parameters were computed for each of the 35 unique 1-day window segments of available data.
-The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components, respectively, of the 35 transfer functions.
-The average impulse response functions were obtained by averaging the 35 impulse response functions.
+Transfer function parameters were computed for each of the 35 1-day segments of available data.
+The average transfer function magnitude and phase was obtained by calculating their values each segment and then averaging the segments values.
 
-There are many statistical methods and variations thereof that have been developed for estimating transfer functions in the magnetotelluric (MT) literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Sims1971} and \cite{Simpson2005}.
-The objective of these works is typically the estimation of a robust and low bias transfer function.
-In the results presented, we did not use a robust statistical method or attempt to correct for bias in estimates as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters.
+There are many statistical methods and variations thereof that have been developed for estimating transfer functions in the magnetotelluric (MT) literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used in this work that follow the basic method used in \cite{Sims1971} and \cite{Simpson2005}.
+Several additional processing methods and steps beyond the basic method that have been used in the MT literature were tested individually by modifying the basic least-squares method used to determine if there was a change in the computed transfer functions and, most relevant for this work, the out-of-sample prediction performance of each model.
+When using these additional steps, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance is within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the basic method used when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s).
+The experiments in this regard are discussed in the Appendix.
+
+%The objective of these works is typically the estimation of a robust and low bias transfer function.
+
+%In the results presented, we did not use a robust statistical method or attempt to correct for bias in estimates as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters.
 As discussed and evaluated in~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).
-
-However, several additional processing methods and steps used in the MT literature were tested individually by modifying the basic least-squares method used to determine if there was a change in the computed transfer functions and, most relevant for this work, the out-of-sample prediction performance of each model.
-When using these additional steps, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance is within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s).
-The experiments in this regard are discussed in the Appendix.
 
@@
 
-The out-of-sample prediction metrics for each model on a given day was determined by computing the average model parameters using the remaining 34 days of data, generating a prediction time series on each of the 34 days using the average model parameters, and then computing the prediction metrics for each of the 34 days.
+The out-of-sample prediction metrics for each model on day $D$ were determined by computing the average model parameters using the remaining 34 days of data, using these averaged parameters to generate a prediction time series on day $D$.
 This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
@@
 The performance of each model in predicting GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.
-The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the predicted data that predicted by the model.
 
-The average PE and MSE ratio was computed by averaging the 35 out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the 35  out-of-sample signal-to-noise ratios at that frequency.
-The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the 35 prediction metrics.
+The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the data that predicted by the model.
+
+The average PE and MSE ratios were computed by averaging the 35 out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the out-of-sample signal-to-noise ratios at that frequency.
+The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the out-of-sample prediction metrics.
 The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.
@@
 
-Figure~\ref{predictions} shows a 1-day interval for selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the lowest.
+Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the lowest.
 
-The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1.
-Model 3, which uses an estimate of the electric field based on horizontal magnetic field measurements provides better estimates of GIC than Model 2.
-Model 4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides the best predictions of GIC.
+The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides significantly better estimates of GIC than Model~3.
+
+In terms of statistical significance, using $10^5$ bootstrap samples, the null hypothesis that the PE of Model~2 is equal to that of Model~1 can be rejected at a significance level of less than $10^{-5}$, with the same result for Model~3 vs.
+Model~2.
+The null hypothesis that the PE of Model~4 is equal to that of Model~3 can be rejected at a significance level of $10^{-2}$.
 
@@
 
-Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models.
-The error bars represent $\pm$ one standard error of the 35 SN ratios at each period.
-At each evaluation frequency the signal-to-noise ratio is is the ratio of the average of the 35 segment PSDs of GIC to the average of the 35 segment PSDs of the prediction error.
-The PSD for each of the segments at a given evaluation frequency was obtained by averaging the raw PSD using values between the next higher and lower evaluation frequency.
+Figure~\ref{SN} shows the average signal-to-noise (i.e., signal to prediction error) ratios as a function of period for the four models.
+At each evaluation frequency, the SN ratio is the ratio of the average of the SN ratios for each of the 35 segments and the error bars represent 95\% CIs based on 1000 bootstrap samples of the segment SN ratios at each period.
+The spectra for the signal and prediction error for each of the segments at a given evaluation frequency was obtained by averaging the raw spectra using values between the next higher and lower evaluation frequency.
 
-Consistent with the results in Table~1, The SN ratio for Model 1 is lower than that for Models 2-4 at all frequencies.
+Consistent with the results in Table~1, The SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies. However, ordering of the SN ratio for Models~2-4 is dependent on period.
 
-Below $600$~s (10~min), Model 2 has the largest SN ratio.
-This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.
-
-Below $600$~s (10~min), the predictive ability of each model is in the ordering that may be expected based on the data used to derive the model: Model 2, which uses direct measurements of the electric field, has a higher SN ratio than Model 3, which uses an indirect estimate of the electric field.
-Model 4 has a slightly higher SN than Model 3, and this is expected because they both use the same input of the magnetic field, but Model 3 has an additional source of uncertainty from the parameters in $\mathbf{Z}$.
-Above $600$~s, the ordering of the SN is not consistent with this expectation and is possibly explained by two factors (1) the measured GIC is based on an integrated electric field and the electric field used in Model 2 is a point measurement - the result is that implicit in the model is an assumption that the electric field is uniform over the length of the line on which GIC is measured and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
+Below $600$~s (10~min), Model~2 has the largest SN ratio.
+This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high frequency fluctuations in the prediction error for Model~2 are the smallest among the models.
+For these periods, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a higher SN ratio than Model~3, which uses an indirect estimate of the electric field.
+Model~4 has a slightly higher SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the parameters in $\mathbf{Z}$.
+Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors (1) the measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement - the result is that implicit in the model is an assumption that the electric field is uniform over the length of the line on which GIC is measured and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
+This issue is addressed in the following section.
 
@@
 
-The average parameters and 95\% CI for Model 1 are $a_o = $ $73 [62,64]$ A/(V/km) and $b_o =$ $-72.0 \pm [-98,-45]$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~\ref{histogram}.
+The average parameters and 95\% CI for Model 1 are $a_o = $ $73 [62,64]$ A/(V/km) and $b_o =$ $-72.0 \pm [-98,-45]$ A/(V/km) and the distribution of the parameters for the segments is shown in Figure~\ref{histogram}.
 (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).
@@
 
-The average parameters for Models 2 and 4 are shown in Figure~\ref{H}--\ref{Phi}.
+The average parameters for Models~2 and~4 are shown in Figure~\ref{H}--\ref{Phi}.
 The error bars at each time step or period is the standard error of the 35 values used to compute the average.
@@
 
-Figure~\ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar and are different from zero at more than one time value (the impulse response for Model 3 is similar to that of Model 2 and is not shown).
-By definition, the impulse response for Model 1 is zero except at $t=0$.
-Note that the sign of the impulse response at $t=0$ is opposite for Model 1 and Model 2 for the $E_y$ coefficient - at $t=0$, $b_o < 0$ while $h_b(t=0)>0$.
-An explanation for the acausality of the impulse response is given in \cite{Kelbert2017}.
+Figure~\ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model~2 and $\mathbf{B}$ for Model~4 are similar and are different from zero at more than one time value (the impulse response for Model~3 is similar to that of Model~2 and is not shown).
+By definition, the impulse response for Model~1 is zero except at $t=0$.
+Note that the sign of the impulse response at $t=0$ is opposite for Model~1 and Model~2 for the $E_y$ coefficient (at $t=0$, $b_o < 0$ while $h_b(t=0)>0$).
+An explanation for the acausality of the impulse response is given in \cite{Egbert1992,Kelbert2017}.
 
-Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition).
-In Model~2, GIC more sensitive to $E_y$ due to $b(\omega) > a(\omega)$, but note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition).
+For this system, GIC more sensitive to $E_y$ due to $b(\omega) > a(\omega)$, but note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).
 The ratio of $|\overline{b}_o/|/|\overline{a}_o|$ is approximately 1 for Model~1 and this can be compared with $|b(\omega)|/|a(\omega)|$ calculated using the lines for $|a(\omega)|$ and $|b(\omega)|$ shown in Figure~\ref{Z}.
@@
 
-Figure~\ref{Phi} shows the transfer function phases.
-By definition, that for Model 1 is zero.
-The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120 degrees while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.
-\todo{Why?} At a period of $60$~s, the parameters in Equation~\ref{model2} are most consistent with the assumptions of frequency independent system parameters of Model~1.
-The phases of $a(\omega)$ and $b(\omega)$ are near zero and the $b(\omega)$ curve is the flattest.
+Figure~\ref{Phi} shows the transfer function phases (that for Model~1 is 0$^o$ for positive parameters and 180$^o$ for negative parameters by definition).
+The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120$^o$ while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.
 
@@
 
-There are at least three possible explanations of the frequency dependence of the data-derived system parameters: (1) the quasi-DC approximation is invalid, (2) invalid model assumptions, and (3) artifacts of the measurement process.
+There are at least three possible explanations for the frequency dependence of the data-derived system parameters: (1) the quasi-DC approximation is invalid, (2) it is an artifact of the measurement process, and (3) the electric field is not spatially and spectrally constant across the region of the power system.
 
-(1) When the quasi-DC approximation is used, ~\cite{Albertson1981} is usually cited; the justification for the quasi-DC approximation is given the Appendix, where it is stated that ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects.
-The correction factor varies from 0.95 to 1.0 depending on conductor size.'' \cite{Lehtinen1985} notes justifies the use of the quasi-DC approximation because GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency.
-Carson's equations are equations for the frequency-dependent impedance associated with a wire above a conducting plane.
-The justification of \cite{Lehtinen1985} is essentially a statement that the frequency dependent component of Carson's impedance is proportional to frequency, the frequency dependent impedance of a power line is small at 50~Hz and so it should be even smaller at GIC frequencies of < 1~Hz.
-\todo{Need a reference here to 50~Hz values} Thus it appears that the quasi-DC approximation is justified.
+(1) When the quasi-DC approximation is used, ~\cite{Albertson1981} is usually cited; their justification for the quasi-DC approximation is given the Appendix where it is stated that ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects.
+The correction factor varies from 0.95 to 1.0 depending on conductor size.'' (Carson's equations  \citep{Carson1926,Kersting2002} are equations for the impedance associated with a wire above an infinite and uniform conducting slab.) \cite{Lehtinen1985} justifies the use of the quasi-DC approximation using the argument that GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency.
+Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala200}, the reactance will be $\sim$10 times smaller or less than the resistance at significant GIC frequencies of $\sim$1~Hz and less.
+Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half space conducting slab Earth also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
 
-(2) The electric field was measured at a single location $\sim$9 km away from the GIC instrument.
-The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the earth potential at the grounding points of the system, and simulations of GIC flow \citep{Albertson1981,Lehtinen1985} typically start with an assumed electric field along the line to calculate these potential differences.
-The potential differences are then used along with the system component resistances and connectivity to calculate GIC in the different parts of the system.
-In both Model~1 and Model~2, the input electric field is assumed to be spatially homogeneous along the line in which GIC was measured.
-If it is not, the parameters derived from the data in Model~1 may vary significantly from that expected.
-The same should be true of Model~2, but this modeling error is not expected to introduce a frequency dependence.
+(2) A second consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence.
+However, the response characteristics of these instruments are essentially frequency independent relative the observed frequency dependence for periods in the range considered of $1-10^{5}$~s \citep{Simpson2005}.
+\todo{Need reference to GIC instrument}
 
-(3) A final consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence.
-Given the response characteristics of these instruments are essentially frequency independent above a period of tens of seconds, this does not explain the frequency dependence observed over the range of 10-10,000~s.
+(3) The equation of Model~1 assumes that the electric field is spatially uniform at all points in the systems.
+The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the earth potential at the grounding points of the system, and simulations of GIC flow \citep{Albertson1981,Lehtinen1985} typically start with an assumed electric field along the lines to calculate these potential differences.
+The potential differences are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system.
+In both Model~1 and Model~2, the input electric field is assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field.
+If it is not, the parameters derived from the data in both models may vary significantly from those calculated using system configuration and resistance information.
+However, this will not explain a frequency dependence of system parameters as both Models~1 and~2 make this assumption.
+
+A second assumption of Model~1 is that that the spectra of the electric field is uniform.
+Violation of this assumption can lead to an apparent frequency dependence in its parameters.
+Suppose the electric field and GIC are measured at one end of a long line segment and the electric field is non-zero only at a frequency of $\omega_1$.
+If the electric field is spectrally uniform along the full line segment, then from the quasi-static approximation it follows that the measured GIC will also only be non-zero at a frequency of $\omega_1$.
+If the electric field on the far half of the line changes so that it is non-zero only at a frequency of $\omega_2$, the observed GIC will appear to be inconsistent with the quasi-static approximation because it will have non-zero variations at $\omega_1$ and $\omega_2$.
+
+To simulate the effect of different points on the line having different electric field spectra, we created a simulated GIC signal using electric field measurements made in Memambetsu (MMB) and Kakioka (KAK), Japan: $G_s(t)=a_o(E^{MMB}_x(t) + fE^{KAK}_x(t))+b_o(E^{MMB}_y(t) + fE^{KAK}_y(t))$ with $a_o=b_o=1$ and $f=0.2$.
+$G_s$ is thus a simulated GIC in which 17\% of the line has a different electric field.
+The model $G_s(\omega)=a(\omega)E^{MMB}_x+b(\omega)E^{MMB}_y$ was then used to compute $a(\omega)$ and $b(\omega)$, which are both found to exhibit a frequency dependence of similar magnitude to those shown in Figure~\ref{Z}.
+
+Although the correlation is -0.40 for $E^{MMB}_x(t)$ and $E^{KAK}_x(t)$ and 0.64 for $E^{MMB}_y$ and $E^{KAK}_y$, this simulation has the flaw that Kakioka is $\sim$10x farther away than the length of the power line ($100$~km) on which GIC was measured.
+A more realistic simulation would use electric field measurements that were both near the line on which GIC was measured and separated from MMB by a distance of $\sim$50~km.
+
+%In the case that $E^{MMB}_x(t) = E^{KAK}_x(t)$, our calculation would give a frequency independent $a$.
+
+%The violation of the spectral uniformity assumption may contribute to the explanation of the finding that Model~4 provides significantly better estimates of GIC than Model~1.
+On average, the magnetic field is more uniform spectrally than the electric field and so the parameters of Model~4 may be capturing both a MT response and spectral averaging effect.
+
+%This explanation may also explain the finding that Model 4 provides significantly better predictions than Model 2 at long periods (but not at short periods).
+The magnetic field spectra typically exhibits significantly less spatial variability so the computed model parameters are able to capture a spatial averaging effect in addition to the effect of the ground impedance.
+
+%One way to test this hypothesis is to consider time intervals in which the spatial variability in the spectra of $\mathbf{E}$ is low -- the frequency dependence of $a(\omega)$ and $b(\omega)$ should be less than average.
+The frequency dependence of $a(\omega)$ and $b(\omega)$ should depend on the average spatial gradient in $\mathbf{E}_x(\omega)$ and $\mathbf{E}_y(\omega)$, respectively.
+(A reasonable test of this is not possible with this dataset because the nearest electric field instrument is $\sim$10x farther away from MMB, in Kakioka, Japan, than the length of the power line on which GIC was measured.)
+
+%The finding that at long periods, the ratio of $a(\omega)/b(\omega)$ approaches $|\overline{b}_o/|/|\overline{a}_o|$
 
@@
 
-The primary implications of the results presented are (1) Significantly improved predictions of GICs can be made if frequency dependent system parameters are used and (2) A direct comparison of $a_o$ and $b_o$ estimated using Equation~\ref{model1} and computed using power system configuration information is problematic because the data are not consistent with the assumptions of Equation~\ref{model1}.
+The primary implications of the results presented are (1) Significantly improved predictions of GICs using data-derived system parameters $a$ and $b$ can be made if they are allowed to be frequency dependent and (2) A direct comparison of data-derived $a$ and $b$ parameters and those computed using power system configuration information will be problematic if the data are not consistent with the assumptions of Equation~\ref{model1} as was the case for the data set considered.
 
-Although the use of $a_o$ and $b_o$ values estimated from the data give a reasonable out-of-sample prediction model, it is not known how much these data-estimated values will differ from those directly calculated.
-(Because the details of the power system under consideration in this work are not available, we were not able make this comparison.)
+Although the use of frequency independent $a$ and $b$ parameter values estimated from data give a reasonable out-of-sample prediction model, it is not known how much data-estimated parameter values will differ from those directly calculated.
+(Because the details of the power system under consideration in this work are not available, we were not able make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
+However, we have shown using a data set of contemporaneous GIC and $\mathbf{E}$ measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
 
-Also note that $\sim 10$ days of data are required to give values of $a_o$ and $b_o$ that provide the out-of-sample prediction quality reported here for Model~1 (a PE of 0.35 and CC of 0.64).
-If only a few days of data are used for their estimation, the out-of-sample prediction quality is very poor due to the high degree of uncertainty of in their estimates.
-Predictions of GIC using Model~1 in the literature using only a few days of data are expected to be significantly worse if out-of-sample data are used.
-
-Several methods have been used in the literature for estimating or predicting GIC depending on the data or models that are available.
-
-When electric field measurements are available, a \todo{TODO} {\color{red} \textbf{[REFS]}} model of the form of Model~1, with frequency independent system parameters derived from power system configuration, has been used.
-These works generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
-However, we have shown using a data set of contemporaneous GIC and $\mathbf{E}$ measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
-In addition, we have found that the estimates of the system parameters $a_o$ and $b_o$ are highly sensitive to the interval of data used, and have a high degree of uncertainty.
-
-When only a MT impedance tensor that connects $\mathbf{E}$ to $\mathbf{B}$ is available, a model of the form of Model~1, with the electric field estimated using $\mathbf{E}=\mathbf{Z}\mathbf{B}$ has been used~\citep{Weigel2017,Bonner2017,Kelbert2017}.
-For the dataset considered, significantly better GIC estimates can be made if frequency dependent system parameters are available as in Model~3.
+%The out-of-sample prediction quality reported here for Model~1 (a PE of 0.35 and CC of 0.64).
+If only a few days of data are used for their estimation, the out-of-sample prediction quality using all data is very poor due to the high degree of uncertainty of in their estimates \todo{Give example}.
+As a result, predictions of GIC using Model~1 in the literature using only a few days of data are expected to be significantly worse if out-of-sample predictions had been used.
 
 When GIC and magnetic field measurements are available, a model of the form of Model~4 can be used.
-In this case, the transfer function contains components of the transfer function that connects $\mathbf{E}$ to $\mathbf{B}$ and GIC to $\mathbf{E}$.
-For this particular dataset, we have found that such a model provides the best estimates of GIC.
-A power system operator could use local ground magnetic field estimates along with GIC measurements to derive a transfer function model that is updated when the system configuration changes, and this model could be used to estimate GIC under extreme or simulated geomagnetic conditions.
-A key advantage of this approach is that geomagnetic field measurements are generally easier to make in the long-term and are more readily available.
+For this particular data set, we have found that such a model provides the best estimates of GIC in terms of the PE and MSE metrics.
+Local ground magnetic field estimates along with GIC measurements could be used to derive a transfer function model that is updated when the system configuration changes, and this model could be used to estimate GIC under extreme or simulated geomagnetic conditions.
+A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and are more readily available.
+A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect $\mathbf{E}$ to $\mathbf{B}$ and GIC to $\mathbf{E}$ and frequency dependent artifacts due to violations of the assumption that that the field is spectrally uniform.
 
-Based on the results in this work, the model of the form $GIC(t)$ $=a_oE_x(t)+b_oE_y(t)$, where the electric field is either measured or estimated using a transfer function, should be regarded as an approximation similar to that assumed when an equation of the form $GIC(t)$ $=c_oB_x(t)+d_oB_y(t)$ or $GIC(t)$ $=c_odB_x(t)/dt+d_odB_y(t)/dt$ is used, where $a_o, b_o, c_o, $and $d_o$ are constants (e.g., {\color{red} \textbf{[REFS]}}).
+%Based on the results in this work, the model of the form $GIC(t)$ $=a_oE_x(t)+b_oE_y(t)$, where the electric field is either measured or estimated using a transfer function at a given point near where GIC was measured, should be regarded as an approximation similar to that assumed when an equation of the form $GIC(t)$ $=c_oB_x(t)+d_oB_y(t)$ or $GIC(t)$ $=c_odB_x(t)/dt+d_odB_y(t)/dt$ is used, where $a_o, b_o, c_o, $and $d_o$ are constants (e.g., {\color{red} \textbf{[REFS]}}).
+
+%Several methods have been used in the literature for estimating or predicting GIC depending on the data or models that are available.
+
+%In addition, we have found that the estimates of the system parameters $a_o$ and $b_o$ are highly sensitive to the interval of data used, and have a high degree of uncertainty.
+
+%When only a MT impedance tensor that connects $\mathbf{E}$ to $\mathbf{B}$ is available, a model of the form of Model~3, with the electric field estimated using $\mathbf{E}=\mathbf{Z}\mathbf{B}$ has been used~\citep{Marshall2013,Weigel2017,Bonner2017,Kelbert2017}.
+For the data set considered, significantly better GIC estimates can be made if frequency dependent system parameters are available.
+When the electric field $\mathbf{E}'$ is use in Model~1, the PE is X.
+\todo{Calculate X}
 
@@
 
-\todo{All - Need help in discussing the origin of the frequency dependence of the system parameters.}
+%\todo{All - Need help in discussing the origin of the frequency dependence of the system parameters.}
 
@@
 \caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval.
-Note that the scales for Models~2-4 differ from that for Model 1 so that the error amplitudes for Models~2-4 are scaled by a factor of 2 relative to that for Model~1.}
+Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled by a factor of 2.}
 \label{predictions}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_SN-options-1.pdf}
-\caption{Out-of-sample signal-to-noise ratios at the evaluation frequencies for the four models.}
+\caption{Out-of-sample signal-to-noise (signal to prediction error) ratios at the evaluation frequencies for the four models.}
 \label{SN}
@@
 Each set of parameters was computed 35 times using one day of 1-second cadence data.
-The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
+The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 73 A/(V/km) and $b_o$ = -72.0 A/(V/km).}
 \label{histogram}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_H-options-1.pdf}
-\caption{Impulse responses for Models 1, 2, and 4.
-The circles at $t = 0$ correspond to the impulse response for Model 1; the red/blue lines are the impulse responses for Model 2 and 4.
-The values of $a_o$ and $b_o$ have been scaled by a factor of 10 for readability.}
+\caption{Impulse responses for the parameters in Models 1, 2, and 4.
+The circles at $t$ $=0$ correspond to the impulse response for Model 1; the red/blue lines are the impulse responses for Model 2 and 4.
+The values of $\overline{a}_o$ and $\overline{b}_o$ reduced by a factor of 10.}
 \label{H}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1.pdf}
-\caption{Frequency domain transfer functions for Models 1-3, 2, and 4. The frequency domain transfer functions for $a_o$ and $b_o$ in Model~1 are constant and equal to $a_o$ and $b_o$.}
+\caption{Frequency domain transfer functions for the parameters in Models 1, 2, and 4.
+The frequency domain transfer functions for $a_o$ and $b_o$ in Model~1 are constant and equal to $a_o$ and $b_o$.}
 \label{Z}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
-\caption{Frequency domain phase values for Models 1, 2, and 4. The phase for $a_o$ and $b_o$ in Model~1 are constant and equal to zero.}
+\caption{Frequency domain phase values for the parameters in Models 1, 2, and 4.
+The phase for $a_o$ and $b_o$ are constant either 0 or 180 degrees depending on the sign of their value, with positive values having a phase of 0.}
 \label{Phi}

### 2019-03-29 paper.tex
Updates from Overleaf

@@
 \item Power system parameters derived from a time series of GIC and geoelectric field measurements are frequency dependent.
-\item A GIC prediction model with frequency dependence significantly outperforms a frequency independent model.
+\item A GIC prediction model with frequency dependent system parameters significantly outperforms a frequency independent model.
 \item A model derived using GIC/geomagnetic field data provides better predictions than a GIC/geoelectric field-based model.
@@
 \begin{abstract}
-A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a nearby direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive.
+A common assumption used when estimating geomagnetically induced currents (GICs) in a power system given a nearby direct measurement or indirect estimate of the horizontal geoelectric field components $E_x(t)$ and $E_y(t)$ on Earth's surface is that the system is resistive.
 That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.
-We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.
-We compute frequency dependent network parameters $a(\omega)$ and $b(\omega)$ from data in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better estimates of the measured GIC.
-A third model, which uses the electric field estimated from a transfer function convolved with the geomagnetic field, provides significantly better estimates to Model~2.
-It is suggested that the measurement-derived frequency dependence of the system parameters is explained by spatial variations in the spectra of the geoelectric field rather than by an actual frequency dependence of the system parameters.
+A first test of this assumption is made using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.
+Frequency-dependent network parameters $a(\omega)$ and $b(\omega)$ computed from data in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$,  are shown to be frequency dependent and this model provides a significantly better estimates of the measured GIC over Model~1.
+A third model, which uses the electric field estimated by convolving a magnetotelluric transfer function with the geomagnetic field, provides significantly better estimates over Model~2.
+It is suggested that the measurement-derived frequency dependence of the system parameters is explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters.
 \end{abstract}
@@
 
-Geomagnetically induced currents (GICs) are electric currents in conducting systems due to electric fields near Earth's surface.
+Geomagnetically induced currents (GICs) are electric currents in conducting systems that are due to electric fields near the Earth's surface.
 Time-varying electric and magnetic fields at Earth's surface are driven by time variations in ionospheric currents on time scales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
-Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}.
-Accurate estimation of GICs magnitudes is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}.
+Of particular interest are currents induced in electric power systems because they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012,Gaunt2014}.
+Accurate estimation of GIC magnitudes is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}.
 
-A common assumption made in estimates of GIC using either measured or estimated values of the horizontal geoelectric field at Earth's surface, $\mathbf{E}$, is that the power system is resistive or quasi-DC in the sense that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.
-In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression, when contemporaneous measurements of GIC and $\mathbf{E}$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}.
+A common assumption made in estimates of GIC in an electric power system using either a measured or estimated time series of the horizontal geoelectric field at Earth's surface, $\mathbf{E}(t)$, is that the power system is resistive or quasi-DC in the sense that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.
+In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression when contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.,][]{Boteler2014a,Boteler2014b}.
 This quasi-DC assumption has been used or stated in many GIC-related works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015,Bonner2017}.
@@
 
-In this work, we use a unique dataset in which measurements $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$, were made near a site where GICs were being measured in an electric power system.
-We first estimate the system parameters $a_o$ and $b_o$ from data using conventional methods, which assume that they are frequency independent.
-Then, we use a model in which the system parameters are allowed to be frequency dependent, compute the frequency dependence, and compare this model with the traditional frequency independent model.
+In this work, we use a unique dataset in which time series measurements of $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$, were made near a site where GIC was being measured in an electric power system.
+We first estimate the system parameters $a_o$ and $b_o$ from data using conventional methods, in which they are modeled as frequency independent.
+Then, we use a model in which the system parameters are allowed to be frequency dependent, compute the frequency dependence, and compare this model's ability to represent GIC with the traditional frequency-independent model.
 
@@
 
-The 1-second-cadence GIC data used span a total of 35 days across 10 continuous intervals that start and end at midnight Japan Standard Time (JST): 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22.
+The 1-second-cadence GIC data used span a total of 35 days across 10 intervals that start and end at midnight Japan Standard Time (JST): 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22.
 These data have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
@@
 The GIC dataset includes 1~s raw data and 1~Hz low-pass-filtered raw measurements.
-The results presented here are for the 1~Hz low-pass-filtered GIC values
+The results presented here are for the 1~Hz low-pass-filtered GIC measurements
 
-One-second-cadence $\mathbf{E}$ and $\mathbf{B}$ measurements from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
+One-second-cadence $\mathbf{E}(t)$ and $\mathbf{B}(t)$ measurements from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
 
@@
 
-The only modification made to the magnetic field measurements was the replacement of 36 data points in all components on 2006-08-06 with linearly interpolated values.
+The only modification made to the magnetic field measurements was the replacement of 36 data points on 2006-08-06 with linearly interpolated values.
 The 36 values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.
@@
 The GIC data contains non-physical spikes that are followed by a relaxation time of 40-80 s.
-These spikes were removed by identifying times $t_s$ when the change in 1~s $|GIC|$ measurements was greater than $0.05$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
+These spikes were removed by identifying times $t_s$ when the change in the 1~s absolute value of $GIC(t)$ was greater than $0.05$~Amps and then values in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
 In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
@@
 \noindent
-the matrix equation $\underline{\mathbf{GIC}} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $\underline{\mathbf{GIC}}$ is a zero mean 864000x1 matrix of 1 day observations with rows of $GIC(t)$; $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ observations with each column having zero mean.
-The values of $\mathbf{p}$ are then obtained using linear regression.
+the matrix equation $\underline{\mathbf{GIC}} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is solved in a least squares sense, where $\underline{\mathbf{GIC}}$ is a zero mean 86400x1 matrix of 1 day observations with rows of $GIC(t)$; $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ with zero mean columns.
 The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ over a 1-day interval.
-\citep[][provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}.
-The notation $\overline{a}_o$ and $\overline{b}_o$ is used for the average of the $a_o$ and $b_o$ values computed using the 35 unique 1-day window segments of available data.
+\citep[][provided the mathematically equivalent closed-form equations for solving the matrix equation.]{Pulkkinen2007} The notation $\overline{a}_o$ and $\overline{b}_o$ will be used for the average of the $a_o$ and $b_o$ values computed using the 35 1-day window segments of available data.
 
@@
 \noindent
-and are computed using contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ using the method described later in this section.
-Note that fourier transformed variables are indicated by explicitly listing their dependence on $\omega$.
+and are computed using contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$.
+The method of parameter estimation and generating predictions for Models 2-4 are given at the end of this section and fourier transformed variables are indicated by explicitly listing their dependence on $\omega$.
 
-In Model~3, the parameters $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$:
+In Model~3, the parameters $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field, $\mathbf{B}$, and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ and
 
@@
 where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. 
-That is, instead of using the measured electric field directly as in Model 2, a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.
-In this work, $\boldsymbol{Z}$ was calculated using the method described later in this section.
+That is, instead of using the measured electric field directly as in Model 2, a magnetotelluric transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.
 
@@
 \noindent
-In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$, and the model is expected to produce predictions of $GIC(t)$ that are similar to that obtained from Model 3 because they both use the same input variable.
+In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$ and no electric field measurements are used; this model is expected to produce predictions of GIC that are similar to that obtained from Model 3 because they both use the same input variable.
 
-Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a least squares regression using Equation~\ref{model2} and the set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $f_e$ from the fourier transform of 24 hours of 1-s cadence data.
-The values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$; if $f_e$ value is not a multiple of $1/86400$, the nearest multiple is chosen.
-The regression for $f_e$ uses all $f$ values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$.
-
-To convert the frequency domain transfer functions $a(\omega)$ and $b(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $h_a(\tau)$ and $h_b(\tau)$, their real and imaginary components are linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed.
+Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $f_e$ from the DFT (discrete fourier transform) of 24 hours of 1-s cadence data.
+The 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if the $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
+The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation window has a lower bound of $1/86400$~Hz and highest evaluation window has an upper bound of $0.5$~Hz.
 
@@
 
-%It is expected that the time and frequency domain methods should give similar results because the transfer functions in the time- and frequency-domains are mathematically related \citep{Schoukens2004,Ljung2007}, although differences may exist as a result of implementation choices or roundoff \citep{Ljung2004}.
-For example, regarding implementation, in the time domain method we choose a finite value of $N_c$ and $N_a$ and in the frequency domain method, we choose the set of evaluation frequencies $\omega_e$, an averaging window function and width, and $\omega$ values (which depend on the length of the time series used to calculate the spectra).
+Transfer function parameters were computed for each of the segments.
+The average transfer function magnitude and phase at a given evaluation frequency was obtained by averaging the segment values at that evaluation frequency.
 
-%Because the frequency domain computations are significantly faster, the results presented here are based on the frequency domain results and only spot checks were performed to verify that the time- and frequency-domain methods gave consistent results.
+To convert the frequency domain transfer function parameters estimated at the evaluation frequencies to impulse responses, their real and imaginary components are linearly interpolated individually onto the original DFT frequency grid (with spacing of $1/86400$~Hz), and then the inverse DFT is computed.
+The interpolated transfer function parameters were set to zero at frequencies higher than $0.25$~Hz and the transfer function estimate at $f=0$ was set to zero before interpolation.
 
-Transfer function parameters were computed for each of the 35 1-day segments of available data.
-The average transfer function magnitude and phase was obtained by calculating their values each segment and then averaging the segments values.
+Prediction time series were calculated by convolution of the interpolated transfer function parameters with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(f)=a^I(f)E_x(f)+b^I(f)E_y(f)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(f)$ was calculated to give the prediction $G_E(t)$ on a 1~s time grid).
 
-There are many statistical methods and variations thereof that have been developed for estimating transfer functions in the magnetotelluric (MT) literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used in this work that follow the basic method used in \cite{Sims1971} and \cite{Simpson2005}.
-Several additional processing methods and steps beyond the basic method that have been used in the MT literature were tested individually by modifying the basic least-squares method used to determine if there was a change in the computed transfer functions and, most relevant for this work, the out-of-sample prediction performance of each model.
-When using these additional steps, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance is within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the basic method used when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s).
-The experiments in this regard are discussed in the Appendix.
-
-%The objective of these works is typically the estimation of a robust and low bias transfer function.
-
-%In the results presented, we did not use a robust statistical method or attempt to correct for bias in estimates as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters.
-As discussed and evaluated in~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).
+Several modifications to this basic procedure were considered.
+Robust regression \citep{Egbert1986} with a Huber weighting constant of $1.345$ and an upper limit of 50 iterations and bi-square weighting with a constant of $4.685$ (both constants provide 95\% efficiency for normally distributed errors), both with no hard-cutoff step; time domain pre-whitening and tapering; and weighting the frequencies around each evaluation frequency using Parzen window weights.
+When using these additional steps individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
+Transfer function parameter values found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)
 
@@
 
-The out-of-sample prediction metrics for each model on day $D$ were determined by computing the average model parameters using the remaining 34 days of data, using these averaged parameters to generate a prediction time series on day $D$.
+The out-of-sample prediction metrics for each model on day $D$ were determined by computing the average model parameters using the remaining 34 days of data and then using these averaged parameters to generate a prediction time series and associated metrics on day $D$.
 This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
+In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the effect of transients that exist in the prediction at the start of the interval for the amount of time that the causal part of the impulse response is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.
 
@@
 
-The average PE and MSE ratios were computed by averaging the 35 out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the out-of-sample signal-to-noise ratios at that frequency.
+The average PE and MSE ratios were computed by averaging the out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the out-of-sample signal-to-noise ratios at that frequency.
 The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the out-of-sample prediction metrics.
-The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.
+(The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.) The spectra for the signal and prediction error for each of the segments at a given evaluation frequency was obtained by averaging the raw spectra using the same frequency windows used for the regression.
 
@@
 MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1.
-The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples of these 35 values.}
+The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
 \centering
@@
 
-Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the lowest.
+Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the highest.
 
-The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides significantly better estimates of GIC than Model~3.
+The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
 
-In terms of statistical significance, using $10^5$ bootstrap samples, the null hypothesis that the PE of Model~2 is equal to that of Model~1 can be rejected at a significance level of less than $10^{-5}$, with the same result for Model~3 vs.
-Model~2.
+To determine the statistical significance, $10^5$ bootstrap averages of the PEs of each model were created and their differences were computed.
+From this, the null hypothesis that the PE of Model~2 is equal to that of Model~1 can be rejected at a significance level of less than $10^{-5}$ because the bootstrap average of Model~2 was not found to be lower than that for Model~1.
+The null hypothesis that the PE of Model~3 is equal to that of Model~2 can also be rejected at the same significance level.
 The null hypothesis that the PE of Model~4 is equal to that of Model~3 can be rejected at a significance level of $10^{-2}$.
 
-Although inferior, the Model~1 does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
+%Although inferior, the Model~1 does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
 The top panel of  Figure~\ref{predictions} shows that its predicted GIC generally follows the overall trend of the measured GIC.
 
-%In a similar way that $|\mathbf{B}|(t)$ or $d|\mathbf{B}|(t)/dt$ are often good predictors of $GIC(t)$ \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GICs, but is a poor approximation relative to Models 2 and 3.
-
 Figure~\ref{SN} shows the average signal-to-noise (i.e., signal to prediction error) ratios as a function of period for the four models.
-At each evaluation frequency, the SN ratio is the ratio of the average of the SN ratios for each of the 35 segments and the error bars represent 95\% CIs based on 1000 bootstrap samples of the segment SN ratios at each period.
-The spectra for the signal and prediction error for each of the segments at a given evaluation frequency was obtained by averaging the raw spectra using values between the next higher and lower evaluation frequency.
+At each evaluation frequency, the SN ratio is the ratio of the average of the out-of-sample SN ratios of the segments and the error bars represent a 95\% CIs based on 1000 bootstrap samples of the segment SN ratios.
 
@@
 Below $600$~s (10~min), Model~2 has the largest SN ratio.
-This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high frequency fluctuations in the prediction error for Model~2 are the smallest among the models.
+This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high-frequency fluctuations in the prediction error for Model~2 are the smallest among the models.
 For these periods, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a higher SN ratio than Model~3, which uses an indirect estimate of the electric field.
@@
 Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors (1) the measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement - the result is that implicit in the model is an assumption that the electric field is uniform over the length of the line on which GIC is measured and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
-This issue is addressed in the following section.
+Issues related to this are discussed in the following section.
 
@@
 
-The average parameters and 95\% CI for Model 1 are $a_o = $ $73 [62,64]$ A/(V/km) and $b_o =$ $-72.0 \pm [-98,-45]$ A/(V/km) and the distribution of the parameters for the segments is shown in Figure~\ref{histogram}.
+The average parameters and 95\% CI for Model 1 are $a_o = 73$ $[62,83]$ A/(V/km) and $b_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the parameters for the segments is shown in Figure~\ref{histogram}.
 (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).
@@
 
-The average parameters for Models~2 and~4 are shown in Figure~\ref{H}--\ref{Phi}.
-The error bars at each time step or period is the standard error of the 35 values used to compute the average.
-The transfer function magnitudes were calculated by linear interpolation of the transfer function magnitudes at the evaluation frequencies (instead of linearly interpolating the real and imaginary components and then computing the magnitude).
+The average parameters for Models~2 and~4 are shown in Figure~\ref{H}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.
 
-Figure~\ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model~2 and $\mathbf{B}$ for Model~4 are similar and are different from zero at more than one time value (the impulse response for Model~3 is similar to that of Model~2 and is not shown).
+%Figure~\ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model~2 and $\mathbf{B}$ for Model~4 are similar and are different from zero at more than one time value (the impulse response for Model~3 is similar to that of Model~2 and is not shown).
 By definition, the impulse response for Model~1 is zero except at $t=0$.
-Note that the sign of the impulse response at $t=0$ is opposite for Model~1 and Model~2 for the $E_y$ coefficient (at $t=0$, $b_o < 0$ while $h_b(t=0)>0$).
-An explanation for the acausality of the impulse response is given in \cite{Egbert1992,Kelbert2017}.
+Note that the sign of the $E_y$ impulse response at $t=0$ for Model~1 is opposite that for Model~2 (at $t=0$, $b_o < 0$ while $h_b(t=0)>0$).
+(An explanation for the acausality of the impulse response is given in \cite{Egbert1992,Kelbert2017}).
 
 Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition).
-For this system, GIC more sensitive to $E_y$ due to $b(\omega) > a(\omega)$, but note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).
-The ratio of $|\overline{b}_o/|/|\overline{a}_o|$ is approximately 1 for Model~1 and this can be compared with $|b(\omega)|/|a(\omega)|$ calculated using the lines for $|a(\omega)|$ and $|b(\omega)|$ shown in Figure~\ref{Z}.
+For this system, GIC more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+The ratio of $|\overline{b}_o|/|\overline{a}_o|$ is approximately 1 for Model~1 and this can be compared with $|b(\omega)|/|a(\omega)|$ calculated using the lines for $|a(\omega)|$ and $|b(\omega)|$ shown in Figure~\ref{Z}.
 The ratio varies from 2.7 at $T=30$~s to 1.6 at $T=12$~h.
@@
 
-(1) When the quasi-DC approximation is used, ~\cite{Albertson1981} is usually cited; their justification for the quasi-DC approximation is given the Appendix where it is stated that ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects.
-The correction factor varies from 0.95 to 1.0 depending on conductor size.'' (Carson's equations  \citep{Carson1926,Kersting2002} are equations for the impedance associated with a wire above an infinite and uniform conducting slab.) \cite{Lehtinen1985} justifies the use of the quasi-DC approximation using the argument that GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency.
-Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala200}, the reactance will be $\sim$10 times smaller or less than the resistance at significant GIC frequencies of $\sim$1~Hz and less.
-Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half space conducting slab Earth also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
+(1) When the quasi-DC approximation is used, \cite{Albertson1981} is usually cited; their justification for the quasi-DC approximation is given the Appendix where it is stated that ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects.
+The correction factor varies from 0.95 to 1.0 depending on conductor size.'' (Carson's equations  \citep{Carson1926,Grigsby2007} are equations for the impedance associated with a wire above an infinite and uniform half-space conducting slab.) \cite{Lehtinen1985} justifies the use of the quasi-DC approximation using the argument that GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency.
+Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala2005}, the reactance will be $\sim$10 times smaller or less than the resistance at GIC frequencies of $\sim$1~Hz and lower.
+Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
 
 (2) A second consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence.
-However, the response characteristics of these instruments are essentially frequency independent relative the observed frequency dependence for periods in the range considered of $1-10^{5}$~s \citep{Simpson2005}.
-\todo{Need reference to GIC instrument}
+However, the response characteristics of electric and magnetic field instruments are essentially frequency independent relative the observed frequency dependence for periods in the range considered of $10-10^{5}$~s \citep{Simpson2005} and the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range.
 
@@
 In both Model~1 and Model~2, the input electric field is assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field.
-If it is not, the parameters derived from the data in both models may vary significantly from those calculated using system configuration and resistance information.
-However, this will not explain a frequency dependence of system parameters as both Models~1 and~2 make this assumption.
+If it is not, the parameters derived from the data in both models may differ significantly from those calculated using system configuration and resistance information.
+However, the data-derived parameters will not have a frequency dependence that results from a variation in the direction of the electric field along the line.
 
-A second assumption of Model~1 is that that the spectra of the electric field is uniform.
-Violation of this assumption can lead to an apparent frequency dependence in its parameters.
+The second assumption of Model~1 is that the spectrum of the electric field is uniform.
+Violation of this assumption can lead to an apparent violation of the quasi-static approximation.
 Suppose the electric field and GIC are measured at one end of a long line segment and the electric field is non-zero only at a frequency of $\omega_1$.
 If the electric field is spectrally uniform along the full line segment, then from the quasi-static approximation it follows that the measured GIC will also only be non-zero at a frequency of $\omega_1$.
-If the electric field on the far half of the line changes so that it is non-zero only at a frequency of $\omega_2$, the observed GIC will appear to be inconsistent with the quasi-static approximation because it will have non-zero variations at $\omega_1$ and $\omega_2$.
+If the electric field on the far half of the line is non-zero only at a frequency of $\omega_2$, the observed GIC and electric field will appear to be inconsistent with the quasi-static approximation because GIC will have non-zero variations at both $\omega_1$ and $\omega_2$.
 
-To simulate the effect of different points on the line having different electric field spectra, we created a simulated GIC signal using electric field measurements made in Memambetsu (MMB) and Kakioka (KAK), Japan: $G_s(t)=a_o(E^{MMB}_x(t) + fE^{KAK}_x(t))+b_o(E^{MMB}_y(t) + fE^{KAK}_y(t))$ with $a_o=b_o=1$ and $f=0.2$.
-$G_s$ is thus a simulated GIC in which 17\% of the line has a different electric field.
-The model $G_s(\omega)=a(\omega)E^{MMB}_x+b(\omega)E^{MMB}_y$ was then used to compute $a(\omega)$ and $b(\omega)$, which are both found to exhibit a frequency dependence of similar magnitude to those shown in Figure~\ref{Z}.
+To test the effect of different points on the line having different electric field spectra on the parameter estimates, we created a simulated GIC signal using electric field measurements made in Memambetsu (indicated with a superscript M) and Kakioka (superscript K), Japan:
 
-Although the correlation is -0.40 for $E^{MMB}_x(t)$ and $E^{KAK}_x(t)$ and 0.64 for $E^{MMB}_y$ and $E^{KAK}_y$, this simulation has the flaw that Kakioka is $\sim$10x farther away than the length of the power line ($100$~km) on which GIC was measured.
-A more realistic simulation would use electric field measurements that were both near the line on which GIC was measured and separated from MMB by a distance of $\sim$50~km.
+\begin{equation}
+G_s(t)=a_o\big[(1-r)E^{\mbox{\tiny M}}_x(t) + rE^{\mbox{\tiny K}}_x(t)\big]+b_o\big[(1-r)E^{\mbox{\tiny M}}_y(t) + rE^{\mbox{\tiny K}}_y(t)\big]
+\end{equation}
+
+\noindent
+with $a_o=1$, $b_o=-1$ and $r=0.2$.
+$G_s$ represents a simulated GIC in which 20\% of the line has a different electric field.
+(The zero-lag correlation is -0.40 for $E^{\mbox{\tiny M}}_x(t)$ and $E^{\mbox{\tiny K}}_x(t)$ and 0.64 for $E^{\mbox{\tiny M}}_y(t)$ and $E^{\mbox{\tiny K}}_y(t)$).
+The model equation
+\begin{equation}
+G_s(\omega)=a(\omega)E^{\mbox{\tiny M}}_x(\omega)+b(\omega)E^{\mbox{\tiny M}}_y(\omega)
+\end{equation}
+\noindent
+was then used to compute $a(\omega)$ and $b(\omega)$ and they were both found to exhibit a frequency dependence similar to that shown in Figure~\ref{Z} in that they both have an increase in magnitude by a factor of 2-6 for periods in the range of $2$~s to $12$~hr.
+
+Although this proof-of-concept simulation shows how variations in the frequency content of the geoelectric field on the line in which GIC is measured can cause the data-estimated system parameters to have a frequency dependence, it is very crude as it uses several approximations that were necessitated by data availability.
+The distance from Kakioka to Memambetsu of $\sim$1400~km is $\sim$14x than the length of the power line of $\sim$100~km on which GIC was measured, and the correlations between the electric field from the two sites are likely to be lower than the fields measured at points separated by a distance closer to the length of the power line.
+In addition, the discontinuous change in the electric field implicit in the model does not satisfy Maxwell's equations and thus is not physically realizable.
+A more realistic simulation would use electric field measurements that were both near a line on which GIC was measured and separated from each other by a distance that is comparable to the length of the line.
 
@@
 %The violation of the spectral uniformity assumption may contribute to the explanation of the finding that Model~4 provides significantly better estimates of GIC than Model~1.
-On average, the magnetic field is more uniform spectrally than the electric field and so the parameters of Model~4 may be capturing both a MT response and spectral averaging effect.
+On average, the magnetic field is more uniform spectrally than the electric field and so the parameters of Model~4 may be capturing both an MT response and spectral averaging effect.
 
 %This explanation may also explain the finding that Model 4 provides significantly better predictions than Model 2 at long periods (but not at short periods).
-The magnetic field spectra typically exhibits significantly less spatial variability so the computed model parameters are able to capture a spatial averaging effect in addition to the effect of the ground impedance.
+The magnetic field spectra typically exhibit significantly less spatial variability so the computed model parameters are able to capture a spatial averaging effect in addition to the effect of the ground impedance.
 
@@
 
-The primary implications of the results presented are (1) Significantly improved predictions of GICs using data-derived system parameters $a$ and $b$ can be made if they are allowed to be frequency dependent and (2) A direct comparison of data-derived $a$ and $b$ parameters and those computed using power system configuration information will be problematic if the data are not consistent with the assumptions of Equation~\ref{model1} as was the case for the data set considered.
+The primary implications of the results presented are (1) significantly improved predictions of GICs using data-derived system parameters $a$ and $b$ can be made if they are allowed to be frequency dependent and (2) a direct comparison of data-derived $a$ and $b$ parameters and those computed using power system configuration information will be problematic if the data are not consistent with the assumptions of Equation~\ref{model1}.
 
 Although the use of frequency independent $a$ and $b$ parameter values estimated from data give a reasonable out-of-sample prediction model, it is not known how much data-estimated parameter values will differ from those directly calculated.
-(Because the details of the power system under consideration in this work are not available, we were not able make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
+(Because the details of the power system under consideration in this work are not available, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
 However, we have shown using a data set of contemporaneous GIC and $\mathbf{E}$ measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
@@
 For this particular data set, we have found that such a model provides the best estimates of GIC in terms of the PE and MSE metrics.
-Local ground magnetic field estimates along with GIC measurements could be used to derive a transfer function model that is updated when the system configuration changes, and this model could be used to estimate GIC under extreme or simulated geomagnetic conditions.
+Local ground magnetic field estimates along with GIC measurements could be used to derive a transfer function model that is updated when the system configuration changes and this model could be used to estimate GIC under extreme or simulated geomagnetic conditions.
 A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and are more readily available.
-A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect $\mathbf{E}$ to $\mathbf{B}$ and GIC to $\mathbf{E}$ and frequency dependent artifacts due to violations of the assumption that that the field is spectrally uniform.
+A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect $\mathbf{E}$ to $\mathbf{B}$ and GIC to $\mathbf{E}$ and frequency dependent artifacts due to violations of the assumption that the field is spectrally uniform.
 
@@
 For the data set considered, significantly better GIC estimates can be made if frequency dependent system parameters are available.
-When the electric field $\mathbf{E}'$ is use in Model~1, the PE is X.
+When the electric field $\mathbf{E}'$ is used in Model~1, the PE is X.
 \todo{Calculate X}
 
-%~\ref{pulkkinen2010} used a variation of this approach, but used an assumed conductivity model to estimate the frequency independent system parameters.
+%~\ref{pulkkinen2010} used a variation of this approach but used an assumed conductivity model to estimate the frequency independent system parameters.
 
@@
 %Ideally, frequency dependent system parameters could be calculated analytically for comparison with those estimated using data in this work.
-Because the details of the power system under consideration in this work are not available, we were not able make this comparison.
+Because the details of the power system under consideration in this work are not available, we were not able to make this comparison.
 Note that frequency independent system parameters calculated using the approach in this work may not be directly comparable to those found from a network analysis for several reasons: (1) At periods less than $\sim 30$~s, the signal-to-noise ratio is low and the parameter estimates are expected to be biased; as shown in the Appendix, the transfer function estimates are highly sensitive to the method used for their calculation at (2) Also at periods less than $\sim 30$~s, effects of the instrument response may be significant.
@@
 We would like to thank the Hokkaido Electric Power Co.
-for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the data.
+for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the GIC data and providing feedback on the manuscript.
 %\clearpage
@@
 
+\iffalse
+
 \begin{figure}[h]
@@
 \end{figure}
+
+\fi
 
@@
 
-\appendix
-\section{Appendix}
-
-%Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.
-
-In the body of the work, a basic ordinary least squares method was used to compute transfer functions.
-Of course, there are many problems that one must address before the parameter estimates from a linear regression model can be used for a physical interpretation because the estimates may be biased, have high uncertainty, and be sensitive to outliers.
-In addition, to make inferences about ground conductivity, corrections must be made for distortion effects and violations of the plane wave assumption \citep{Simpson2005}.
-
-Many robust alternatives to the basic least-squares for the calculation of transfer functions have been proposed and are strongly advocated by their developers to reduce uncertainty and sensitivity to outliers.
-For example, \cite{Egbert2011} notes that the least-squares method often fails and \cite{Chave1987} that failures are sometimes spectacular but more often insidious.
-However, ~\cite{Chave2017} has noted that the commonly used robust regression models are also problematic because the error model assumed for the calculation of robust transfer functions is often invalid due to the pervasiveness of non-gaussian residuals, and as a result it is not unconditionally unbiased.
-
-There are also many approaches that are taken in pre-processing the data before regression is performed, for example, windowing in the time domain prior to the calculation of the fourier transform, windowing in the frequency domain to down-weight values far from the evaluation frequency or to down-weight frequencies that have low coherence, weighting values by an empirically estimated variance, pre-whitening, and using multi-taper methods to estimate spectra.
-
-In general, the approach taken in the MT literature is to select a given variant of least squares and a set of pre-processing steps and to report the transfer functions along with an uncertainty (several approaches to estimating uncertainty exist as well).
-Absent an understanding of which particular regression alternative to use, the appropriate pre-processing approach, and the appropriate uncertainty estimate method, and an understanding of the appropriate adjustable parameters of each option and how they interact, the approach taken here is to investigate the influence of a few of the common regression alternatives, pre-processing steps, and stack-averaging methods on the transfer function estimate.
-
-The primary conclusion is that different methods generally give the statistically similar results in both the prediction metrics and parameter estimates except when the signal to noise ratio is low (below approximately $30$ seconds).
-Although parameter estimates were plotted below $30$ seconds because they were a part of the prediction models and they do contribute somewhat to the prediction efficiency, the estimates should be regarded as highly uncertain due to large error bars and a large sensitivity to the method used to calculate them.
-
-The various tests that were performed to determine the influence of each method are summarized below and the associated results are shown in Figure~\ref{model_2_and_4_variations}.
-
-\begin{enumerate}
-
-\item Robust regression was used instead of linear regression when computing the transfer function estimates at each $\omega_e$ (using a custom implementation with the $\psi$-function referenced in the next item and also using the \texttt{robustfit} function in MATLAB with either the bi-square and Huber weighting options).
-The motivation for the use of robust regression is to reduce the influence of outliers on the parameter estimates; ordinary least squares parameter estimates have a high sensitivity to outliers.
-When using robust regression, PEs and MSEs shown in Table~1 were the same to two decimal places and the confidence intervals were smaller by $\sim 2$\%.
-Figures~\ref{a_variations}-~\ref{b_variations} shows that the robust and used method give near identical estimates to the least square regression estimates.
-
-\item Each 1-day segment was pre-whitened prior to calculation of the transfer functions in order to reduce the spectral range and to remove long-period drifts.
-This was done by computing the pre-whitening filter using a 10th-order Yule-Walker filter using $E_y$ for Model~2 and $B_x$ for Model~4 and applying this filter to all time series used in the calculation of the transfer function.
-The out-of-sample PEs were the same to within two decimal places and the MSE ratios varied by $0.05$.
-Figures~\ref{a_variations}-~\ref{b_variations} show that pre-whitening has a larger impact on the estimates than robust regression.
-
-\item A Parzen window was used to weight the fourier-transformed values instead of a rectangular window.
-This reduces the influence of values used in the transfer function parameter estimate as a function of distance from the evaluation frequency.
-The out-of-sample PEs for Model 3 and Model 4 were higher by $0.02$ and $0.01$, respectively and the MSE ratios for Models 3 and 4 were higher by $0.4$ and $0.6$, respectively.
-Figures~\ref{a_variations}-~\ref{b_variations} show that the pre-whitening has about the same amount of influence as the use of a Parzen window.
-
-\item Five methods for averaging were used to compute magnitudes at each evaluation frequency of a complex-valued model parameter $p(D)$, where $D$ is estimate made using a 1-day of data.
-If $u(D)$ and $v(D)$ are the real and imaginary phase values for Models 2-4, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation).
-The out-of-sample PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$.
-Neither averaging method was yielded consistently higher or lower values.
-Figure~\ref{model_2_and_4_variations} shows that most of the sensitivity is below $T=20$ s, as expected, due to the signal-to-noise being near equal to 1 below this period.
-
-\end{enumerate}
-
-Several addition experiments were performed that we summarize with text only.
-
-\begin{enumerate}
-\item Instead computing an average transfer function by a simple average of the 35 transfer functions, weighted averages were considered.
-The weighting included using the in-sample coherence or prediction efficiency (ad-hoc approaches to give higher weight to better fitting intervals) and weighting the collection of values used to compute the model parameters at a given evaluation frequency by a smoothed estimate of the PSD of the predictand (to account for the fact that the signal to noise ratio depends on frequency as evident in ~\ref{SN} and noted by \citep{Egbert1997}; accounting for this can reduce the variance in the parameter estimates).
-An approach that is more similar to that used in \citep{Egbert1997} was also used - prior to regression, the the values were weighted with a weight that is inversely proportional to an estimate of of the spectra of the vector magnitude of the input.
-
-\item At a given frequency, we have found that there is a correlation between the parameter estimate and the PSD of the input, output, and prediction error that varies between +0.6 and -0.2, which is one of the many ways in which the data violate the assumptions for OLS estimates to be the best linear unbiased estimator.
-To determine how this violation could effect the parameter estimates, at each frequency, the top 17 and lowest 18 transfer functions were averaged and plotted as a function of frequency.
-The parameter estimates from the high input PSD intervals tended to be higher, but within the range of the displayed error bars.
-
-\item Instead of calculating 35 transfer functions and then averaging, compute one transfer function - for each evaluation frequency, combine the 35 fourier spectra values near that evaluation frequency and use them for one regression.
-\end{enumerate}
-
-%In all cases, the parameter estimates below $\sim 20$ seconds using the alternative methods were consistently lower, but the offset was of magnitude that was similar to the variation observed between the alternative methods.
-The largest variations were found in relative to the method used to averaging the 35 transfer functions.
-In addition, how the averages were made when computing the magnitudes mattered - if the magnitudes at a given frequency were calculated by taking the modulus of the mean of the 35 real and mean of the 35 imaginary components vs the mean of the modulus of the 35 transfer functions.
-The same was true for the median.
-
 \clearpage
-
-\begin{figure}[h]
-\begin{minipage}[t]{.48\linewidth}
-\includegraphics[width=\textwidth]{figures/plot_options_comparison_a.pdf}
-\label{plot_options_comparison_a}
-\end{minipage}
-\begin{minipage}[t]{.48\linewidth}
-\includegraphics[width=\textwidth]{figures/plot_options_comparison_b.pdf}
-\end{minipage}
-\begin{minipage}[t]{.48\linewidth}
-\includegraphics[width=\textwidth]{figures/plot_options_comparison_zx.pdf}
-\label{plot_options_comparison_a}
-\end{minipage}
-\begin{minipage}[t]{.48\linewidth}
-\includegraphics[width=\textwidth]{figures/plot_options_comparison_zy.pdf}
-\end{minipage}
-\caption{Caption}
-\label{model_2_and_4_variations}
-\end{figure}
-
-\clearpage
-
 \bibliography{paper.bib}
-
-% Please use ONLY \citet and \citep for reference citations.
-% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
-%% Example \citet and \citep:
-%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
-%  \citet{Ghel00}, and \citet{Leit74}.
-
-%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
-%  \citep{Ghel00, Leit74}.
 

### 2019-04-08 figures/update.sh
update figs

@@ -4,14 +4,16 @@ cp ../../figures/predictions/plot_model_predictions-MeanModel-2006-08-06.pdf .
 
 DIR=../../figures/combined
 cp $DIR/plot_model_summary_aobo_histograms-options-1.pdf .
-cp $DIR/plot_model_summary_H-options-1.pdf .
+#cp $DIR/plot_model_summary_H-options-1.pdf .
 cp $DIR/plot_model_summary_Phi-options-1.pdf .
 cp $DIR/plot_model_summary_Z-options-1.pdf .
+cp $DIR/plot_model_summary_Phi_MT-options-1.pdf .
+cp $DIR/plot_model_summary_Z_MT-options-1.pdf .
 cp $DIR/plot_model_summary_SN-options-1.pdf .
-cp $DIR/plot_options_comparison_a.pdf .
-cp $DIR/plot_options_comparison_b.pdf .
-cp $DIR/plot_options_comparison_zx.pdf .
-cp $DIR/plot_options_comparison_zy.pdf .
+#cp $DIR/plot_options_comparison_a.pdf .
+#cp $DIR/plot_options_comparison_b.pdf .
+#cp $DIR/plot_options_comparison_zx.pdf .
+#cp $DIR/plot_options_comparison_zy.pdf .
 
 cp $DIR/../20060805/plot_raw_All_20060805.pdf .
 git commit -a -m "update figs"

### 2019-04-08 paper.tex

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \iffalse
@@
 \fi
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_model_summary_Z_MT-options-1.pdf}
+\caption{Frequency domain transfer functions for the parameters in Model~3.}
+\label{Z_MT}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_model_summary_Phi_MT-options-1.pdf}
+\caption{Frequency domain phase values for the parameters in Model~3.}
+\label{Phi_MT}
+\end{figure}
+
 \clearpage
 
-\clearpage
 \bibliography{paper.bib}

### 2019-04-08 paper.tex

@@
 \usepackage[colorinlistoftodos]{todonotes}
+
 %\usepackage{framebox}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{figures/plot_model_summary_H-options-1.pdf}
-\caption{Impulse responses for the parameters in Models 1, 2, and 4.
-The circles at $t$ $=0$ correspond to the impulse response for Model 1; the red/blue lines are the impulse responses for Model 2 and 4.
-The values of $\overline{a}_o$ and $\overline{b}_o$ reduced by a factor of 10.}
-\label{H}
-\end{figure}
-
-\fi
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{figure}[h]
@@
 \label{Z}
-\end{figure}
+%\end{figure}
 
-\begin{figure}[h]
+\vspace{4em}
+
+%\begin{figure}[h]
 \centering
@@
 \label{Z_MT}
-\end{figure}
+%\end{figure}
 
-\begin{figure}[h]
+\vspace{4em}
+
+%\begin{figure}[h]
 \centering

### 2019-04-10 paper.tex
Updates from Overleaf

@@
 This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
-In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the effect of transients that exist in the prediction at the start of the interval for the amount of time that the causal part of the impulse response is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.
+In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the prediction at the start of the interval for the amount of time that the causal part of the impulse response is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.
 
@@
 Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the highest.
+Note that the large errors visible at the start and end of the time intervals are due to the transients discussed above and the first and last 10~minutes of each interval were omitted from metrics calculations.
 
@@
 The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120$^o$ while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.
+
+Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3.
+Note that electric and magnetic field data from Memambetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$ (obtained from the IRIS DMC EMTF database \cite{Kelbert2011}).
+When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs 0.78).
+
+%A preliminary investigation suggests that the cause is due to several factors: we have found using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function (using all of the modifications described in Section~3) matches the \cite{Fujii2015} transfer function to within $\pm X$\% except at low periods, which is likely due to the fact that we did not use a remote reference.
+After 2006, our calculated transfer functions differ significantly.
+In addition, we have found that the calculated transfer function is highly sensitive to whether or not stacking was used.
+Due to the large gaps between and the varying length of data segments, transfer functions were computed using one day of data and then averaged.
+We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the result is significantly different than if 21 transfer functions were computed and then averaged.
+
+%As a result, the transfer function parameters shown in Figure~\ref{Z_MT} and~\ref{phi_MT} should be considered a
 
@@
 Each set of parameters was computed 35 times using one day of 1-second cadence data.
-The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 73 A/(V/km) and $b_o$ = -72.0 A/(V/km).}
+The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 73 A/(V/km) and $b_o$ = -72 A/(V/km).}
 \label{histogram}

### 2019-06-26 paper.tex
Updates from Overleaf

--- 2019-04-08 paper.tex
@@
 \usepackage{apacite}
-\usepackage{url} %this package should fix any errors with URLs in refs.
+\usepackage{url} % should fix any errors with URLs in refs.
 \usepackage{amsmath}
@@
 \usepackage[colorinlistoftodos]{todonotes}
-
-%\usepackage{framebox}
-
-%%%%%%%
-%\usepackage{trackchanges}
-% uncomment the line above to use the TrackChanges package to mark revisions if needed.
-% The trackchanges package adds five new LaTeX commands:
-%
-%  \note[editor]{The note}
-%  \annote[editor]{Text to annotate}{The note}
-%  \add[editor]{Text to add}
-%  \remove[editor]{Text to remove}
-%  \change[editor]{Text to remove}{Text to add}
-%
-% complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%https://www.overleaf.com/project/5c3620127117e2522d9e0c30
 
@@
 
-% Not working
-% \listoftodos
-
-\title{A Test of the Frequency Independence Assumption of Power System Parameters used in Geomagnetically Induced Current Estimates}
+\title{A Test of the Frequency Independence Assumption of Power System Coefficients used in Geomagnetically Induced Current Estimates}
 
@@
 \begin{keypoints}
-\item Power system parameters derived from a time series of GIC and geoelectric field measurements are frequency dependent.
-\item A GIC prediction model with frequency dependent system parameters significantly outperforms a frequency independent model.
-\item A model derived using GIC/geomagnetic field data provides better predictions than a GIC/geoelectric field-based model.
+\item Power system coefficients derived from a time series of GIC and geoelectric field measurements are frequency dependent.
+\item A GIC model with frequency-dependent system coefficients provides significantly better estimates of GIC than a frequency independent model.
+\item Data-derived power system coefficients may differ significantly with those computed using power system configuration information.
 \end{keypoints}
@@
 \begin{abstract}
-A common assumption used when estimating geomagnetically induced currents (GICs) in a power system given a nearby direct measurement or indirect estimate of the horizontal geoelectric field components $E_x(t)$ and $E_y(t)$ on Earth's surface is that the system is resistive.
-That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.
-A first test of this assumption is made using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.
-Frequency-dependent network parameters $a(\omega)$ and $b(\omega)$ computed from data in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$,  are shown to be frequency dependent and this model provides a significantly better estimates of the measured GIC over Model~1.
-A third model, which uses the electric field estimated by convolving a magnetotelluric transfer function with the geomagnetic field, provides significantly better estimates over Model~2.
-It is suggested that the measurement-derived frequency dependence of the system parameters is explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters.
+A common assumption used when estimating geomagnetically induced currents (GICs) in a power system given a time series of nearby direct measurements or indirect estimates of the horizontal geoelectric field components $E_x(t)$ and $E_y(t)$ on Earth's surface is that the system is resistive.
+That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency-independent power system coefficients.
+A first test of this assumption uses GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.
+A second model (Model 2) is obtained using a frequency domain generalization of Model 1: $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$.
+The coefficients $a(\omega)$ and $b(\omega)$ are computed from data and are shown to be frequency dependent and this model provides significantly better estimates of the measured GIC than Model~1.
+Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system coefficients may be explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters.
+It is also shown that further improvements over Model~2 can be made using a model with the geomagnetic field as an input.
 \end{abstract}
@@
 Geomagnetically induced currents (GICs) are electric currents in conducting systems that are due to electric fields near the Earth's surface.
-Time-varying electric and magnetic fields at Earth's surface are driven by time variations in ionospheric currents on time scales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
+Time-varying electric and magnetic fields at Earth's surface are driven by time variations in ionospheric currents on time scales of minutes to hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
 Of particular interest are currents induced in electric power systems because they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012,Gaunt2014}.
@@
 A common assumption made in estimates of GIC in an electric power system using either a measured or estimated time series of the horizontal geoelectric field at Earth's surface, $\mathbf{E}(t)$, is that the power system is resistive or quasi-DC in the sense that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.
-In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression when contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.,][]{Boteler2014a,Boteler2014b}.
-This quasi-DC assumption has been used or stated in many GIC-related works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015,Bonner2017}.
-
-%\cite{Oyedokun2013a} developed a method to account for the finite inductive response time of transformers by taking the GIC predicted based on the quasi-DC assumption and then applying a piecewise-linear L/R filter derived from bench-scale transformer measurements, with a time constant dependent on load, to obtain an alternative estimate of GIC.
-
-%\todo[inline]{What can I conclude from \cite{Oyedokun2013a}? Mostly the thesis documents visual differences in time series traces with and without accounting for transformers.}
+In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression when contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.,][]{Boteler2014a,Boteler2014b,Boteler2017}.
+This quasi-DC assumption has been used or stated in many GIC-related works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015}.
 
 In this work, we use a unique dataset in which time series measurements of $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$, were made near a site where GIC was being measured in an electric power system.
-We first estimate the system parameters $a_o$ and $b_o$ from data using conventional methods, in which they are modeled as frequency independent.
-Then, we use a model in which the system parameters are allowed to be frequency dependent, compute the frequency dependence, and compare this model's ability to represent GIC with the traditional frequency-independent model.
+We first estimate the system coefficients $a_o$ and $b_o$ from data using conventional methods, in which they are modeled as frequency independent.
+Then, we use a model in which the system coefficients are allowed to be frequency dependent, compute the frequency dependence, and compare this model's ability to represent GIC with the traditional frequency-independent model.
 
@@
 
-The 1-second-cadence GIC data used span a total of 35 days across 10 intervals that start and end at midnight Japan Standard Time (JST): 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22.
+The 1-second-cadence GIC data used span a total of 35 days across 10 intervals that start and end at midnight Japan Standard Time (JST) and are listed in Table~\ref{intervals}.
 These data have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
@@
 
-One-second-cadence $\mathbf{E}(t)$ and $\mathbf{B}(t)$ measurements from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
+\begin{table}
+\caption{Date intervals used for analysis}
+\centering
+\begin{tabular}{l l}
+\hline
+Start Date & End Date\\
+\hline
+2006-04-03 & 2006-04-10\\
+2006-07-25 & 2006-07-29\\
+2006-08-05 & 2006-08-08\\
+2006-08-19 & 2006-08-21\\
+2006-11-07 & 2006-11-12\\
+2006-11-28 & 2006-11-29\\
+2006-12-01 & 2006-12-14\\
+2006-12-15 & 2006-12-15\\
+2007-11-18 & 2007-11-20\\
+2007-11-22 & 2007-11-22\\
+\hline
+\end{tabular}
+\label{intervals}
+\end{table}
+
+One-second-cadence electric and magnetic field measurements from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
 
@@
 
-Spikes in the electric field measurements were assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1~mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
+Spikes in the electric field measurements were assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1~mV/km at time $t$, and the values $E_i(t-2)$ through $E_i(t+4)$ were replaced using linear interpolation of the values at $E_i(t-3)$ and $E_i(t + 5)$.
 In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$.
-For all of the time intervals considered, the fraction of interpolated $\mathbf{E}$ values ranged from 0--0.1\%.
+For all of the time intervals considered, the fraction of interpolated electric field values ranged from 0--0.1\%.
 
 The only modification made to the magnetic field measurements was the replacement of 36 data points on 2006-08-06 with linearly interpolated values.
-The 36 values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.
+The 36 data points for $B_x$ and $B_y$ have the same value and are much larger than the surrounding values.
 
-The GIC data contains non-physical spikes that are followed by a relaxation time of 40-80 s.
-These spikes were removed by identifying times $t_s$ when the change in the 1~s absolute value of $GIC(t)$ was greater than $0.05$~Amps and then values in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
+The GIC data contains non-physical spikes that are followed by 40-80~s of anomalous values that appear as an exponential decay back to pre-spike values.
+These spikes were removed by identifying times $t$ when the change in the 1~s absolute value of $GIC(t)$ was greater than $0.05$~Amps and then values in a window $GIC(t-2)$ through $GIC(t+100)$ were replaced using a linear interpolation of the values at $GIC(t-3)$ and $GIC(t + 101)$.
 In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
@@
 \noindent
-the matrix equation $\underline{\mathbf{GIC}} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is solved in a least squares sense, where $\underline{\mathbf{GIC}}$ is a zero mean 86400x1 matrix of 1 day observations with rows of $GIC(t)$; $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ with zero mean columns.
+the matrix equation $\underline{\mathbf{GIC}} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is solved in a least squares sense, where $\underline{\mathbf{GIC}}$ is a zero mean 86400x1 matrix with rows of $GIC(t)$; $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ with zero mean columns.
 The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ over a 1-day interval.
@@
 
-In Model 2, the system parameters $a(\omega)$ and $b(\omega)$ are frequency dependent according to
+In Model~2, the system coefficients $a(\omega)$ and $b(\omega)$ are frequency dependent according to
 
@@
 \noindent
-and are computed using contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$.
-The method of parameter estimation and generating predictions for Models 2-4 are given at the end of this section and fourier transformed variables are indicated by explicitly listing their dependence on $\omega$.
+and are estimated using the Fourier transforms of  $GIC(t)$ and $\mathbf{E}(t)$; Frequency-dependent variables are indicated by explicitly listing their dependence on $\omega$.
+The method of parameter estimation and the generation of time-domain GIC estimates for Models~2-4 are given at the end of this section.
 
-In Model~3, the parameters $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field, $\mathbf{B}$, and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ and
+Motivated by the results described in the following section in which Model~2 provides a significantly better representation of GIC than Model~1, we have also considered two additional modeling approaches that can be used to specify GIC.
+In these two models, a direct estimate of the geoelectric field is not needed; instead, the driver is the geomagnetic field, which is much more frequently measured, and as a result, these models may be of practical interest.
+
+In Model~3, the coefficients $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field, $\mathbf{B}$, and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ and
 
@@
 where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. 
-That is, instead of using the measured electric field directly as in Model 2, a magnetotelluric transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.
-
-%Substitution gives
-
-%\begin{linenomath*}
-%\begin{equation*}
-%G_{E'}(\omega) = %\big[a(\omega)Z_{xx}(\omega) + b(\omega)Z_{yx}(\omega) \big] B_x(\omega) + \big[ a(\omega)Z_{xy}(\omega) + b(\omega)Z_{yy}(\omega) \big] B_y(\omega)\,.
-%\end{equation*}
-%\end{linenomath*}
+That is, instead of using the measured electric field directly as in Model~2, a magnetotelluric transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.
 
@@
 \noindent
-In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$ and no electric field measurements are used; this model is expected to produce predictions of GIC that are similar to that obtained from Model 3 because they both use the same input variable.
+In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input variable.
 
-Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $f_e$ from the DFT (discrete fourier transform) of 24 hours of 1-s cadence data.
-The 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if the $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
+Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ in a band of frequencies around $f_e$ from the DFT (discrete Fourier transform) of 24 hours of 1-s cadence data.
+Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
 The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation window has a lower bound of $1/86400$~Hz and highest evaluation window has an upper bound of $0.5$~Hz.
@@
 The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}.
-For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ are solved for using $E_x(\omega) = Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_{y}(\omega)$; the $yx$ and $yy$ components are solved for using $E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
+For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ are solved for using
 
-Transfer function parameters were computed for each of the segments.
+\begin{linenomath*}
+\begin{equation*}
+E_x(\omega) = Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_{y}(\omega)
+\end{equation*}
+\end{linenomath*}
+
+\noindent
+and the $yx$ and $yy$ components are solved for using
+
+\begin{linenomath*}
+\begin{equation*}
+E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)
+\end{equation*}
+\end{linenomath*}
+
+Transfer function coefficients were computed for each of the segments.
 The average transfer function magnitude and phase at a given evaluation frequency was obtained by averaging the segment values at that evaluation frequency.
 
-To convert the frequency domain transfer function parameters estimated at the evaluation frequencies to impulse responses, their real and imaginary components are linearly interpolated individually onto the original DFT frequency grid (with spacing of $1/86400$~Hz), and then the inverse DFT is computed.
-The interpolated transfer function parameters were set to zero at frequencies higher than $0.25$~Hz and the transfer function estimate at $f=0$ was set to zero before interpolation.
+To interpolate the frequency-domain transfer function coefficients estimated at the evaluation frequencies, their real and imaginary components are linearly interpolated individually onto the original DFT frequency grid (with a spacing of $1/86400$~Hz), and then the inverse DFT is computed.
+The interpolated transfer function coefficients were set to zero at frequencies higher than $0.25$~Hz and the transfer function estimate at $f=0$ was set to zero before interpolation.
 
-Prediction time series were calculated by convolution of the interpolated transfer function parameters with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(f)=a^I(f)E_x(f)+b^I(f)E_y(f)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(f)$ was calculated to give the prediction $G_E(t)$ on a 1~s time grid).
+GIC time series were calculated by multiplication of the interpolated transfer function coefficients with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(f)=a^I(f)E_x(f)+b^I(f)E_y(f)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(f)$ was calculated to give the estimate of $G_E(t)$ on a 1~s time grid).
 
@@
 When using these additional steps individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
-Transfer function parameter values found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)
+Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods higher than $\sim30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)
 
@@
 
-The out-of-sample prediction metrics for each model on day $D$ were determined by computing the average model parameters using the remaining 34 days of data and then using these averaged parameters to generate a prediction time series and associated metrics on day $D$.
+Out-of-sample estimation metrics for each model on day $D$ were determined by computing the average model coefficients using the remaining 34 days of data and then using these averaged coefficients to generate GIC time series and associated metrics on day $D$.
 This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
-In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the effect of transients that exist in the prediction at the start of the interval for the amount of time that the causal part of the impulse response is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.
+In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the estimate at the start of the interval for the amount of time that the causal part of the impulse response (found by Fourier inversion of the transfer function) is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.
 
-The performance of each model in predicting GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.
+The performance of each model in estimating GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.
+The PE and MSE ratios are shown in Table~\ref{results}.
 
-The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the data that predicted by the model.
+The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted and it represents the fraction of the variance in the data that is predicted by the model.
 
-The average PE and MSE ratios were computed by averaging the out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the out-of-sample signal-to-noise ratios at that frequency.
+The average PE and MSE ratios were computed by averaging the out-of-sample PE and MSE ratios.
 The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the out-of-sample prediction metrics.
-(The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.) The spectra for the signal and prediction error for each of the segments at a given evaluation frequency was obtained by averaging the raw spectra using the same frequency windows used for the regression.
+(The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.)
+
+The average signal-to-noise ratio at a given frequency is the average of the out-of-sample signal-to-noise ratios at that frequency.
+The spectra for the signal and prediction error used in the signal-to-noise ratio calculation were obtained by averaging the raw spectra in the same frequency windows used for the regression.
 
@@
 \caption{Out-of-sample model performance metrics for models $m=1-4$.
-MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1.
+MSE$_1$/MSE$_m$ is the ratio of the MSE for Model~1 to that of Model $m$.
 The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
@@
 \hline
-$m$\hspace{1em} Model & PE & 95\% CI & MSE$_m$/MSE$_1$ & 95\% CI\\
+$m$\hspace{1em} Model & PE & 95\% CI & MSE$_1$/MSE$_m$ & 95\% CI\\
 \hline
@@
 \end{tabular}
+\label{results}
 \end{table}
 
-Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the highest.
+Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest.
+Note that the large errors visible at the start and end of the time intervals are due to the transients discussed above and the first and last 10~minutes of each interval were omitted from metrics calculations.
 
-The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
+The primary conclusions of this work follow from the results shown in Table~~\ref{results}: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
 
-To determine the statistical significance, $10^5$ bootstrap averages of the PEs of each model were created and their differences were computed.
-From this, the null hypothesis that the PE of Model~2 is equal to that of Model~1 can be rejected at a significance level of less than $10^{-5}$ because the bootstrap average of Model~2 was not found to be lower than that for Model~1.
+To determine the statistical significance of the differences, $10^5$ bootstrap averages of the segment PEs of each model were created and their differences were computed.
+From this, the null hypothesis that the PE of Model~2 is equal to that of Model~1 can be rejected at a significance level of less than $10^{-5}$ because no bootstrap average PE difference from Model~2 was found to be lower than that for Model~1.
 The null hypothesis that the PE of Model~3 is equal to that of Model~2 can also be rejected at the same significance level.
@@
 
-%Although inferior, the Model~1 does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64.
-The top panel of  Figure~\ref{predictions} shows that its predicted GIC generally follows the overall trend of the measured GIC.
-
 Figure~\ref{SN} shows the average signal-to-noise (i.e., signal to prediction error) ratios as a function of period for the four models.
 At each evaluation frequency, the SN ratio is the ratio of the average of the out-of-sample SN ratios of the segments and the error bars represent a 95\% CIs based on 1000 bootstrap samples of the segment SN ratios.
-
-Consistent with the results in Table~1, The SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies. However, ordering of the SN ratio for Models~2-4 is dependent on period.
+Consistent with the results in Table~\ref{results}, the SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies.
+However, ordering of the SN ratio for Models~2-4 is dependent on period.
 
@@
 For these periods, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a higher SN ratio than Model~3, which uses an indirect estimate of the electric field.
-Model~4 has a slightly higher SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the parameters in $\mathbf{Z}$.
-Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors (1) the measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement - the result is that implicit in the model is an assumption that the electric field is uniform over the length of the line on which GIC is measured and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
+Model~4 has a slightly higher SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the data-estimated coefficients in $\mathbf{Z}$.
+
+Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors: (1) The measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement made 9~km away from the line - it follows that there is an implicit assumption in the model that the electric field is uniform over the length of the line on which GIC is measured; and (2) The error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
 Issues related to this are discussed in the following section.
 
-Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio.
+Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~\ref{results}, with Model~1 having the lowest and Model~4 having the highest SN ratio.
 
-The average parameters and 95\% CI for Model 1 are $a_o = 73$ $[62,83]$ A/(V/km) and $b_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the parameters for the segments is shown in Figure~\ref{histogram}.
+The average coefficients and 95\% CI for Model 1 are $a_o = 73$ $[62,83]$ A/(V/km) and $b_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}.
 (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).
 
-%However, for the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.
-
-The average parameters for Models~2 and~4 are shown in Figure~\ref{H}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.
-
-%Figure~\ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model~2 and $\mathbf{B}$ for Model~4 are similar and are different from zero at more than one time value (the impulse response for Model~3 is similar to that of Model~2 and is not shown).
-By definition, the impulse response for Model~1 is zero except at $t=0$.
-Note that the sign of the $E_y$ impulse response at $t=0$ for Model~1 is opposite that for Model~2 (at $t=0$, $b_o < 0$ while $h_b(t=0)>0$).
-(An explanation for the acausality of the impulse response is given in \cite{Egbert1992,Kelbert2017}).
+The average coefficients for Models~2 and~4 are shown in Figure~\ref{Z}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.
 
 Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition).
-For this system, GIC more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+
 The ratio of $|\overline{b}_o|/|\overline{a}_o|$ is approximately 1 for Model~1 and this can be compared with $|b(\omega)|/|a(\omega)|$ calculated using the lines for $|a(\omega)|$ and $|b(\omega)|$ shown in Figure~\ref{Z}.
 The ratio varies from 2.7 at $T=30$~s to 1.6 at $T=12$~h.
-At $T=12$~h, the highest period considered, the phases of $a(\omega)$ and $b(\omega)$ shown in Figure~\ref{Phi} differ by 180 degrees and the signs are consistent with $\overline{a}_o$ and $\overline{b}_o$.
+At $T=12$~h, the highest period considered, the phases of $a(\omega)$ and $b(\omega)$ shown in Figure~\ref{Phi} differ by 180 degrees and the signs are consistent with those of $\overline{a}_o$ and $\overline{b}_o$.
 
-Figure~\ref{Phi} shows the transfer function phases (that for Model~1 is 0$^o$ for positive parameters and 180$^o$ for negative parameters by definition).
+Figure~\ref{Phi} shows the transfer function phases (that for Model~1 is 0$^o$ for positive coefficients and 180$^o$ for negative coefficients by definition).
 The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120$^o$ while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.
 
-%\cite{Pulkkinen2010} noted that one could estimate the ratio of $a_o/b_o$ using GIC and local geomagnetic field measurements even if $\mathbf{Z}(\omega)$ was unknown if one assumed the ground conductivity was 1-D so that $Z\equiv Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$ so that Equation~\ref{model2} can be written
-
-%\begin{linenomath*}
-%\begin{equation}
-%G(\omega) = a_oZ(\omega)B_y(\omega)/\mu_o - b_oZ(\omega)B_x(\omega)/\mu_o\,.
-%\label{pulkkinen2010}
-%\end{equation}
-%\end{linenomath*}
-
-%\noindent
-%If the above equations are solved for with unknowns $Z_a\equiv a_oZ$ and $Z_b\equiv b_oZ$,  then $a_o/b_o = Z_a(\omega)/Z_b(\omega)$.
-\cite{Pulkkinen2010} assumed that this ratio was frequency independent and provided a closed-form equation for $a_o/b_o$ by multiplying Equation~\ref{pulkkinen2010} by the complex conjugates of $B_x(\omega)$ and $B_y(\omega)$ so as to form two equations from which $Z$ can be eliminated.
-With the additional assumption that the cross- and auto-spectra of all variables have the same functional form, they showed that the ratio can be written in a way that requires only evaluation of correlations among $G(t)$, $B_x(t)$, and $B_y(t)$.
+Figure~\ref{Z_MT} and~\ref{Phi_MT} show the coefficients for $\mathbf{Z}$ in Model~3.
+Note that electric and magnetic field data from Memambetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
+When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs.
+0.78).
+(The \cite{Fujii2015} transfer function was obtained from the IRIS DMC EMTF database [\cite{Kelbert2011}]).
+Although not critical to the issues addressed in this work, a preliminary investigation suggests that the cause is due to several factors.
+We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015}.
+After 2006, our calculated transfer functions differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%.
+In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used.
+Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function coefficients were computed using one-day segments of data and then averaged.
+We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% higher than if 21 transfer functions were computed and then averaged.
 
@@
 
-There are at least three possible explanations for the frequency dependence of the data-derived system parameters: (1) the quasi-DC approximation is invalid, (2) it is an artifact of the measurement process, and (3) the electric field is not spatially and spectrally constant across the region of the power system.
+There are at least three possible explanations a frequency dependence in data-derived system coefficients: (1) the quasi-DC approximation is invalid, (2) it is an artifact of the measurement process, and (3) the electric field is not spatially and spectrally constant across the region of the power system.
 
-(1) When the quasi-DC approximation is used, \cite{Albertson1981} is usually cited; their justification for the quasi-DC approximation is given the Appendix where it is stated that ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects.
-The correction factor varies from 0.95 to 1.0 depending on conductor size.'' (Carson's equations  \citep{Carson1926,Grigsby2007} are equations for the impedance associated with a wire above an infinite and uniform half-space conducting slab.) \cite{Lehtinen1985} justifies the use of the quasi-DC approximation using the argument that GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency.
-Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala2005}, the reactance will be $\sim$10 times smaller or less than the resistance at GIC frequencies of $\sim$1~Hz and lower.
+(1) When the quasi-DC approximation is used, \cite{Albertson1981} is usually cited; their justification for the quasi-DC approximation is given in an Appendix as ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects.
+The correction factor varies from 0.95 to 1.0 depending on conductor size.'' (Carson's equations are equations for the impedance associated with a wire above an infinite and uniform half-space conducting slab [\cite{Carson1926,Grigsby2007}] and positive sequence resistance values are the per phase line resistances in a balanced 3-phase power system.)
+
+\cite{Lehtinen1985} justifies the use of the quasi-DC approximation using the argument that GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency.
+Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is typically $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala2005}, the reactance will be $\sim$10 times smaller or less than the resistance at GIC frequencies of $\sim$1~Hz and lower.
 Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
+(However, \cite{Boteler1994} noted that for certain systems, the $L/R$ time constant may be large enough to warrant inclusion in DC flow calculations; this topic has more recently been pursued by \cite{Oyedokun2013a}.)
 
 (2) A second consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence.
-However, the response characteristics of electric and magnetic field instruments are essentially frequency independent relative the observed frequency dependence for periods in the range considered of $10-10^{5}$~s \citep{Simpson2005} and the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range.
+However, the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range and the response characteristics of the magnetic and electric field instruments are essentially frequency independent relative to the observed transfer function frequency dependence for periods in the range of $30-10^{5}$~s (Oowada et al.
+1998, Fujii 2019 personal communication.)
 
@@
 In both Model~1 and Model~2, the input electric field is assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field.
-If it is not, the parameters derived from the data in both models may differ significantly from those calculated using system configuration and resistance information.
-However, the data-derived parameters will not have a frequency dependence that results from a variation in the direction of the electric field along the line.
+If it is not, the coefficients derived from the data in both models may differ significantly from those calculated using system configuration and resistance information.
+However, the data-derived coefficients will not have a frequency dependence that results from a variation in the direction of the electric field along the line.
 
@@
 $G_s$ represents a simulated GIC in which 20\% of the line has a different electric field.
-(The zero-lag correlation is -0.40 for $E^{\mbox{\tiny M}}_x(t)$ and $E^{\mbox{\tiny K}}_x(t)$ and 0.64 for $E^{\mbox{\tiny M}}_y(t)$ and $E^{\mbox{\tiny K}}_y(t)$).
+(The zero-lag cross correlation is -0.40 for $E^{\mbox{\tiny M}}_x(t)$ and $E^{\mbox{\tiny K}}_x(t)$ and 0.64 for $E^{\mbox{\tiny M}}_y(t)$ and $E^{\mbox{\tiny K}}_y(t)$).
 The model equation
@@
 \noindent
-was then used to compute $a(\omega)$ and $b(\omega)$ and they were both found to exhibit a frequency dependence similar to that shown in Figure~\ref{Z} in that they both have an increase in magnitude by a factor of 2-6 for periods in the range of $2$~s to $12$~hr.
+was then used to compute $a(\omega)$ and $b(\omega)$ and they were both found to exhibit a frequency dependence similar to that shown in Figure~\ref{Z} in that they both have an increase in magnitude with period, although in this experiment the increase was a factor 2-6 for periods in the range of $2$~s to $12$~hr.
 
-Although this proof-of-concept simulation shows how variations in the frequency content of the geoelectric field on the line in which GIC is measured can cause the data-estimated system parameters to have a frequency dependence, it is very crude as it uses several approximations that were necessitated by data availability.
-The distance from Kakioka to Memambetsu of $\sim$1400~km is $\sim$14x than the length of the power line of $\sim$100~km on which GIC was measured, and the correlations between the electric field from the two sites are likely to be lower than the fields measured at points separated by a distance closer to the length of the power line.
-In addition, the discontinuous change in the electric field implicit in the model does not satisfy Maxwell's equations and thus is not physically realizable.
+Although this proof-of-concept simulation shows how variations in the frequency content of the geoelectric field on the line in which GIC is measured can cause the data-estimated system coefficients to have a frequency dependence, it is very crude as it uses several approximations that were necessitated by the availability of data.
+The distance from Kakioka to Memambetsu is $\sim$1400~km, which is $\sim$14x that of the length of the power line ($\sim$100~km) on which GIC was measured, and the correlations between the electric field from the two sites are likely to be lower than the fields measured at points separated by a distance closer to the length of the power line.
+In addition, the discontinuous change in the electric field implicit in the model implies that there would be a part of the line that can have a non-zero net charge and thus is not physically realizable for typical conductors.
 A more realistic simulation would use electric field measurements that were both near a line on which GIC was measured and separated from each other by a distance that is comparable to the length of the line.
-
-%In the case that $E^{MMB}_x(t) = E^{KAK}_x(t)$, our calculation would give a frequency independent $a$.
-
-%The violation of the spectral uniformity assumption may contribute to the explanation of the finding that Model~4 provides significantly better estimates of GIC than Model~1.
-On average, the magnetic field is more uniform spectrally than the electric field and so the parameters of Model~4 may be capturing both an MT response and spectral averaging effect.
-
-%This explanation may also explain the finding that Model 4 provides significantly better predictions than Model 2 at long periods (but not at short periods).
-The magnetic field spectra typically exhibit significantly less spatial variability so the computed model parameters are able to capture a spatial averaging effect in addition to the effect of the ground impedance.
-
-%One way to test this hypothesis is to consider time intervals in which the spatial variability in the spectra of $\mathbf{E}$ is low -- the frequency dependence of $a(\omega)$ and $b(\omega)$ should be less than average.
-The frequency dependence of $a(\omega)$ and $b(\omega)$ should depend on the average spatial gradient in $\mathbf{E}_x(\omega)$ and $\mathbf{E}_y(\omega)$, respectively.
-(A reasonable test of this is not possible with this dataset because the nearest electric field instrument is $\sim$10x farther away from MMB, in Kakioka, Japan, than the length of the power line on which GIC was measured.)
-
-%The finding that at long periods, the ratio of $a(\omega)/b(\omega)$ approaches $|\overline{b}_o/|/|\overline{a}_o|$
 
@@
 
-The primary implications of the results presented are (1) significantly improved predictions of GICs using data-derived system parameters $a$ and $b$ can be made if they are allowed to be frequency dependent and (2) a direct comparison of data-derived $a$ and $b$ parameters and those computed using power system configuration information will be problematic if the data are not consistent with the assumptions of Equation~\ref{model1}.
+The primary implications of the results presented are (1) significantly improved estimates of GICs using data-derived system coefficients $a$ and $b$ can be made if they are allowed to be frequency dependent and (2) a direct comparison of data-derived $a$ and $b$ coefficients and those computed using power system configuration information will be problematic if the data are not consistent with the assumptions of Equation~\ref{model1}.
 
@@
 (Because the details of the power system under consideration in this work are not available, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
-However, we have shown using a data set of contemporaneous GIC and $\mathbf{E}$ measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
+However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
 
@@
 A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and are more readily available.
-A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect $\mathbf{E}$ to $\mathbf{B}$ and GIC to $\mathbf{E}$ and frequency dependent artifacts due to violations of the assumption that the field is spectrally uniform.
-
-%Based on the results in this work, the model of the form $GIC(t)$ $=a_oE_x(t)+b_oE_y(t)$, where the electric field is either measured or estimated using a transfer function at a given point near where GIC was measured, should be regarded as an approximation similar to that assumed when an equation of the form $GIC(t)$ $=c_oB_x(t)+d_oB_y(t)$ or $GIC(t)$ $=c_odB_x(t)/dt+d_odB_y(t)/dt$ is used, where $a_o, b_o, c_o, $and $d_o$ are constants (e.g., {\color{red} \textbf{[REFS]}}).
-
-%Several methods have been used in the literature for estimating or predicting GIC depending on the data or models that are available.
-
-%In addition, we have found that the estimates of the system parameters $a_o$ and $b_o$ are highly sensitive to the interval of data used, and have a high degree of uncertainty.
-
-%When only a MT impedance tensor that connects $\mathbf{E}$ to $\mathbf{B}$ is available, a model of the form of Model~3, with the electric field estimated using $\mathbf{E}=\mathbf{Z}\mathbf{B}$ has been used~\citep{Marshall2013,Weigel2017,Bonner2017,Kelbert2017}.
-For the data set considered, significantly better GIC estimates can be made if frequency dependent system parameters are available.
-When the electric field $\mathbf{E}'$ is used in Model~1, the PE is X.
-\todo{Calculate X}
-
-%~\ref{pulkkinen2010} used a variation of this approach but used an assumed conductivity model to estimate the frequency independent system parameters.
-
-%The frequency dependence of $a(\omega)$ and $b(\omega)$ shown in Figure~\ref{model2} is not consistent with a simple model of an ideal resistor $R$ in series with an ideal inductor $L$, for which the transfer function magnitude is $R^{-1}(1+\omega^2\tau^2)^{-1}$ and phase is $\tan^{-1}\omega\tau$, where $\tau=L/R$.
-
-%\todo{All - Need help in discussing the origin of the frequency dependence of the system parameters.}
-
-%Ideally, frequency dependent system parameters could be calculated analytically for comparison with those estimated using data in this work.
-Because the details of the power system under consideration in this work are not available, we were not able to make this comparison.
-Note that frequency independent system parameters calculated using the approach in this work may not be directly comparable to those found from a network analysis for several reasons: (1) At periods less than $\sim 30$~s, the signal-to-noise ratio is low and the parameter estimates are expected to be biased; as shown in the Appendix, the transfer function estimates are highly sensitive to the method used for their calculation at (2) Also at periods less than $\sim 30$~s, effects of the instrument response may be significant.
-\cite{Fujii2015} has noted that the 1-s geoelectric field data from the Memambetsu Observatory were low-pass filtered with a 3.3-s cutoff-period prior to 2010; (3) The GIC observed in a power system is based on the integration of the electric field along a power line.
-All of the models assume that the electric field is constant along the line where the GIC is measured, but many works have shown that the electric field is expected to have variations on length scales comparable to the line considered ($\sim$100~km).
-\todo{REFS}
+A disadvantage is the transfer function coefficients are difficult to interpret physically because they contain components of the transfer functions that connect the electric field to the magnetic field and GIC to the electric field and frequency-dependent artifacts due to violations of the assumption that the field is spectrally uniform.
 
 \acknowledgments
-Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
+Magnetic and electric field measurements made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency Data Portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
 We would like to thank the Hokkaido Electric Power Co.
 for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the GIC data and providing feedback on the manuscript.
+We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response, Trevor Gaunt and Emanuel Benabeau for review and comments on part (1) of Section 5, and E.
+Josh Rigler for substantial feedback and review.
 %\clearpage
@@
 \includegraphics[width=\textwidth]{figures/plot_raw_All_20060805.pdf}
-\caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
-(Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$9 km from the Memanbetsu substation.
-(Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
+\caption{(Top) Geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
+(Middle) Geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB), which is $\sim$9 km from the Memanbetsu substation.
+(Bottom) GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
 The subscripts $x$ and $y$ correspond to Geographic North and East, respectively.}
@@
 \caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval.
-Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled by a factor of 2.}
+Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
 \label{predictions}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1.pdf}
-\caption{Distribution of parameters in Model 1.
-Each set of parameters was computed 35 times using one day of 1-second cadence data.
-The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 73 A/(V/km) and $b_o$ = -72.0 A/(V/km).}
+\caption{Distribution of coefficients in Model~1.
+Each set of coefficients was computed 35 times using one day of 1-second cadence data.
+The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 73 A/(V/km) and $b_o$ = -72 A/(V/km).}
 \label{histogram}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1.pdf}
-\caption{Frequency domain transfer functions for the parameters in Models 1, 2, and 4.
+\caption{Frequency domain transfer functions for the coefficients in Models 1, 2, and 4.
 The frequency domain transfer functions for $a_o$ and $b_o$ in Model~1 are constant and equal to $a_o$ and $b_o$.}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
-\caption{Frequency domain phase values for the parameters in Models 1, 2, and 4.
-The phase for $a_o$ and $b_o$ are constant either 0 or 180 degrees depending on the sign of their value, with positive values having a phase of 0.}
+\caption{Frequency domain phase values for the coefficients in Models 1, 2, and 4.
+The phase for $a_o$ and $b_o$ are constant and either 0$^o$ or 180$^o$ depending on the sign of their value, with positive values having a phase of 0$^o$.}
 \label{Phi}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Z_MT-options-1.pdf}
-\caption{Frequency domain transfer functions for the parameters in Model~3.}
+\caption{Frequency domain transfer functions for the coefficients in Model~3.}
 \label{Z_MT}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi_MT-options-1.pdf}
-\caption{Frequency domain phase values for the parameters in Model~3.}
+\caption{Frequency domain phase values for the coefficients in Model~3.}
 \label{Phi_MT}
@@
 \end{document}
-
-More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-% If you are using line numbers, remember to surround
-% equations with \begin{linenomath*}...\end{linenomath*}
-
-%  To add line numbers to lines in equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}

### 2019-06-26 paper-submitted-1.tex
Updates from Overleaf

--- 2019-04-08 paper.tex
@@
 % complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%https://www.overleaf.com/project/5c3620127117e2522d9e0c30
 
@@
 
+%Weigel: http://orcid.org/0000-0002-9521-5228
+%Cilliers: http://orcid.org/0000-0003-3175-5134
+
 \affiliation{1}{Space Weather Lab, George Mason University}
@@
 \item Power system parameters derived from a time series of GIC and geoelectric field measurements are frequency dependent.
-\item A GIC prediction model with frequency dependent system parameters significantly outperforms a frequency independent model.
-\item A model derived using GIC/geomagnetic field data provides better predictions than a GIC/geoelectric field-based model.
+\item A GIC prediction model with frequency-dependent system parameters provides significantly better predictions than a frequency independent model.
+\item A model derived using GIC and geomagnetic field data provides better predictions than a GIC and geoelectric field-based model.
 \end{keypoints}
@@
 \begin{abstract}
-A common assumption used when estimating geomagnetically induced currents (GICs) in a power system given a nearby direct measurement or indirect estimate of the horizontal geoelectric field components $E_x(t)$ and $E_y(t)$ on Earth's surface is that the system is resistive.
-That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.
+A common assumption used when estimating geomagnetically induced currents (GICs) in a power system given a time series of nearby direct measurements or indirect estimates of the horizontal geoelectric field components $E_x(t)$ and $E_y(t)$ on Earth's surface is that the system is resistive.
+That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency-independent network parameters.
 A first test of this assumption is made using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.
-Frequency-dependent network parameters $a(\omega)$ and $b(\omega)$ computed from data in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$,  are shown to be frequency dependent and this model provides a significantly better estimates of the measured GIC over Model~1.
+A second model (Model 2) is obtained using a frequency domain generalization of Model 1: $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$.
+The network parameters $a(\omega)$ and $b(\omega)$ are computed from data and are shown to be frequency dependent and this model provides significantly better estimates of the measured GIC than Model 1.
 A third model, which uses the electric field estimated by convolving a magnetotelluric transfer function with the geomagnetic field, provides significantly better estimates over Model~2.
-It is suggested that the measurement-derived frequency dependence of the system parameters is explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters.
+Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system parameters may be explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters.
+
+%Network parameters $a(\omega)$ and $b(\omega)$ computed from data in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$, are shown to be frequency dependent and this model provides a significantly better estimates of the measured GIC over Model~1.
+
+%It is suggested that the measurement-derived frequency dependence of the system parameters is explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters.
+
 \end{abstract}
@@
 A common assumption made in estimates of GIC in an electric power system using either a measured or estimated time series of the horizontal geoelectric field at Earth's surface, $\mathbf{E}(t)$, is that the power system is resistive or quasi-DC in the sense that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.
-In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression when contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.,][]{Boteler2014a,Boteler2014b}.
+In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression when contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.,][]{Boteler2014a,Boteler2014b,Boteler2017}.
 This quasi-DC assumption has been used or stated in many GIC-related works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015,Bonner2017}.
@@
 
-Spikes in the electric field measurements were assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1~mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$.
+Spikes in the electric field measurements were assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1~mV/km at time $t$, and the values $E_i(t-2)$ through $E_i(t+4)$ were replaced using linear interpolation of the values at $E_i(t-3)$ and $E_i(t + 5)$.
 In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$.
-For all of the time intervals considered, the fraction of interpolated $\mathbf{E}$ values ranged from 0--0.1\%.
+For all of the time intervals considered, the fraction of interpolated electric field values ranged from 0--0.1\%.
 
@@
 
-The GIC data contains non-physical spikes that are followed by a relaxation time of 40-80 s.
-These spikes were removed by identifying times $t_s$ when the change in the 1~s absolute value of $GIC(t)$ was greater than $0.05$~Amps and then values in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$.
+The GIC data contains non-physical spikes that are followed by 40-80~s of anomalous values that appear as an exponential decay back to pre-spike values.
+These spikes were removed by identifying times $t$ when the change in the 1~s absolute value of $GIC(t)$ was greater than $0.05$~Amps and then values in a window $GIC(t-2)$ through $GIC(t+100)$ were replaced using a linear interpolation of the values at $GIC(t-3)$ and $GIC(t + 101)$.
 In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
@@
 Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $f_e$ from the DFT (discrete fourier transform) of 24 hours of 1-s cadence data.
-The 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if the $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
+Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
 The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation window has a lower bound of $1/86400$~Hz and highest evaluation window has an upper bound of $0.5$~Hz.
@@
 The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}.
-For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ are solved for using $E_x(\omega) = Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_{y}(\omega)$; the $yx$ and $yy$ components are solved for using $E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
+For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ are solved for using
+
+\begin{linenomath*}
+\begin{equation*}
+E_x(\omega) = Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_{y}(\omega)
+\end{equation*}
+\end{linenomath*}
+
+\noindent
+and the $yx$ and $yy$ components are solved for using
+
+\begin{linenomath*}
+\begin{equation*}
+E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)
+\end{equation*}
+\end{linenomath*}
 
@@
 
-To convert the frequency domain transfer function parameters estimated at the evaluation frequencies to impulse responses, their real and imaginary components are linearly interpolated individually onto the original DFT frequency grid (with spacing of $1/86400$~Hz), and then the inverse DFT is computed.
+To interpolate the frequency-domain transfer function parameters estimated at the evaluation frequencies, their real and imaginary components are linearly interpolated individually onto the original DFT frequency grid (with a spacing of $1/86400$~Hz), and then the inverse DFT is computed.
 The interpolated transfer function parameters were set to zero at frequencies higher than $0.25$~Hz and the transfer function estimate at $f=0$ was set to zero before interpolation.
@@
 When using these additional steps individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
-Transfer function parameter values found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)
+Transfer function parameter values found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods higher than $\sim30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)
 
@@
 
-The out-of-sample prediction metrics for each model on day $D$ were determined by computing the average model parameters using the remaining 34 days of data and then using these averaged parameters to generate a prediction time series and associated metrics on day $D$.
+Out-of-sample prediction metrics for each model on day $D$ were determined by computing the average model parameters using the remaining 34 days of data and then using these averaged parameters to generate a prediction time series and associated metrics on day $D$.
 This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
-In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the effect of transients that exist in the prediction at the start of the interval for the amount of time that the causal part of the impulse response is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.
+In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the prediction at the start of the interval for the amount of time that the causal part of the impulse response is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.
 
@@
 
-The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the data that predicted by the model.
+The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted and it represents the fraction of the variance in the data that predicted by the model.
 
@@
 \end{tabular}
+\label{results}
 \end{table}
@@
 
-To determine the statistical significance, $10^5$ bootstrap averages of the PEs of each model were created and their differences were computed.
-From this, the null hypothesis that the PE of Model~2 is equal to that of Model~1 can be rejected at a significance level of less than $10^{-5}$ because the bootstrap average of Model~2 was not found to be lower than that for Model~1.
+To determine the statistical significance of the differences, $10^5$ bootstrap averages of the segment PEs of each model were created and their differences were computed.
+From this, the null hypothesis that the PE of Model~2 is equal to that of Model~1 can be rejected at a significance level of less than $10^{-5}$ because no bootstrap average PE difference from Model~2 was found to be lower than that for Model~1.
 The null hypothesis that the PE of Model~3 is equal to that of Model~2 can also be rejected at the same significance level.
@@
 
-Consistent with the results in Table~1, The SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies. However, ordering of the SN ratio for Models~2-4 is dependent on period.
+Consistent with the results in Table~1, the SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies. However, ordering of the SN ratio for Models~2-4 is dependent on period.
 
@@
 Model~4 has a slightly higher SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the parameters in $\mathbf{Z}$.
-Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors (1) the measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement - the result is that implicit in the model is an assumption that the electric field is uniform over the length of the line on which GIC is measured and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
+
+Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors: (1) The measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement made 9~km away from the line - it follows that there is an implicit assumption in the model that the electric field is uniform over the length of the line on which GIC is measured; and (2) The error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
 Issues related to this are discussed in the following section.
@@
 
-The average parameters for Models~2 and~4 are shown in Figure~\ref{H}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.
+The average parameters for Models~2 and~4 are shown in Figure~\ref{Z}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.
 
@@
 For this system, GIC more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+
 The ratio of $|\overline{b}_o|/|\overline{a}_o|$ is approximately 1 for Model~1 and this can be compared with $|b(\omega)|/|a(\omega)|$ calculated using the lines for $|a(\omega)|$ and $|b(\omega)|$ shown in Figure~\ref{Z}.
 The ratio varies from 2.7 at $T=30$~s to 1.6 at $T=12$~h.
-At $T=12$~h, the highest period considered, the phases of $a(\omega)$ and $b(\omega)$ shown in Figure~\ref{Phi} differ by 180 degrees and the signs are consistent with $\overline{a}_o$ and $\overline{b}_o$.
+At $T=12$~h, the highest period considered, the phases of $a(\omega)$ and $b(\omega)$ shown in Figure~\ref{Phi} differ by 180 degrees and the signs are consistent with those of $\overline{a}_o$ and $\overline{b}_o$.
 
@@
 The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120$^o$ while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.
+
+Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3.
+Note that electric and magnetic field data from Memambetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
+When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs.
+0.78).
+(The \cite{Fujii2015} transfer function was obtained from the IRIS DMC EMTF database [\cite{Kelbert2011}]).
+Although not critical to the issues addressed in this work, a preliminary investigation suggests that the cause is due to several factors.
+We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015}.
+After 2006, our calculated transfer functions differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%.
+In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used.
+Due to the large gaps between and the varying length of continuous data segments used in this work, transfer function parameters were computed using one-day segments of data and then averaged.
+We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% higher than if 21 transfer functions were computed and then averaged.
+
+%As a result, the transfer function parameters are shown in Figure~\ref{Z_MT} and~\ref{phi_MT} should be considered a
 
@@
 
-There are at least three possible explanations for the frequency dependence of the data-derived system parameters: (1) the quasi-DC approximation is invalid, (2) it is an artifact of the measurement process, and (3) the electric field is not spatially and spectrally constant across the region of the power system.
+There are at least three possible explanations a frequency dependence in data-derived system parameters: (1) the quasi-DC approximation is invalid, (2) it is an artifact of the measurement process, and (3) the electric field is not spatially and spectrally constant across the region of the power system.
 
-(1) When the quasi-DC approximation is used, \cite{Albertson1981} is usually cited; their justification for the quasi-DC approximation is given the Appendix where it is stated that ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects.
-The correction factor varies from 0.95 to 1.0 depending on conductor size.'' (Carson's equations  \citep{Carson1926,Grigsby2007} are equations for the impedance associated with a wire above an infinite and uniform half-space conducting slab.) \cite{Lehtinen1985} justifies the use of the quasi-DC approximation using the argument that GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency.
-Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala2005}, the reactance will be $\sim$10 times smaller or less than the resistance at GIC frequencies of $\sim$1~Hz and lower.
+(1) When the quasi-DC approximation is used, \cite{Albertson1981} is usually cited; their justification for the quasi-DC approximation is given in an Appendix as ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects.
+The correction factor varies from 0.95 to 1.0 depending on conductor size.'' (Carson's equations are equations for the impedance associated with a wire above an infinite and uniform half-space conducting slab [\cite{Carson1926,Grigsby2007}].)
+
+\cite{Lehtinen1985} justifies the use of the quasi-DC approximation using the argument that GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency.
+Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is typically $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala2005}, the reactance will be $\sim$10 times smaller or less than the resistance at GIC frequencies of $\sim$1~Hz and lower.
 Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
+(However, \cite{Boteler1994} noted that for certain systems, the $L/R$ time constant may be large enough to warrant inclusion in DC flow calculations; this topic has more recently been pursued by \cite{Oyedokun2013a}.)
 
 (2) A second consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence.
-However, the response characteristics of electric and magnetic field instruments are essentially frequency independent relative the observed frequency dependence for periods in the range considered of $10-10^{5}$~s \citep{Simpson2005} and the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range.
+However, the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range and the response characteristics of the magnetic and electric field instruments are essentially frequency independent relative to the observed transfer function frequency dependence for periods in the range of $30-10^{5}$~s (Oowada et al.
+1998, Fujii 2019 personal communication.)
+
+%However, the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range and the response characteristics of the magnetic and electric field instruments are essentially frequency independent relative to the observed transfer function frequency dependence for periods in the range of $30-10^{5}$~s \citep{Oowada1998,Fujii2015}.
+
+%The response characteristics of the electric field instrument measuring at a 1~s cadence (EC85 by Kokusai Electronics Corporation) is not precisely known, but comparison with that of the 0.1~s instrument, for which the response is known and frequency independent above $\sim 30$~s, indicates that the 1~s instrument response is frequency independent above $\sim 30$~s.
 
@@
 \noindent
-was then used to compute $a(\omega)$ and $b(\omega)$ and they were both found to exhibit a frequency dependence similar to that shown in Figure~\ref{Z} in that they both have an increase in magnitude by a factor of 2-6 for periods in the range of $2$~s to $12$~hr.
+was then used to compute $a(\omega)$ and $b(\omega)$ and they were both found to exhibit a frequency dependence similar to that shown in Figure~\ref{Z} in that they both have an increase in magnitude with period, although in this experiment the increase was a factor 2-6 for periods in the range of $2$~s to $12$~hr.
 
 Although this proof-of-concept simulation shows how variations in the frequency content of the geoelectric field on the line in which GIC is measured can cause the data-estimated system parameters to have a frequency dependence, it is very crude as it uses several approximations that were necessitated by data availability.
-The distance from Kakioka to Memambetsu of $\sim$1400~km is $\sim$14x than the length of the power line of $\sim$100~km on which GIC was measured, and the correlations between the electric field from the two sites are likely to be lower than the fields measured at points separated by a distance closer to the length of the power line.
-In addition, the discontinuous change in the electric field implicit in the model does not satisfy Maxwell's equations and thus is not physically realizable.
+The distance from Kakioka to Memambetsu is $\sim$1400~km, which is $\sim$14x that of the length of the power line ($\sim$100~km) on which GIC was measured, and the correlations between the electric field from the two sites are likely to be lower than the fields measured at points separated by a distance closer to the length of the power line.
+In addition, the discontinuous change in the electric field implicit in the model implies that there would be a part of the line that can have a non-zero net charge and thus is not physically realizable for typical conductors.
 A more realistic simulation would use electric field measurements that were both near a line on which GIC was measured and separated from each other by a distance that is comparable to the length of the line.
@@
 (Because the details of the power system under consideration in this work are not available, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
-However, we have shown using a data set of contemporaneous GIC and $\mathbf{E}$ measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
+However, we have shown using a data set of contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
 
@@
 A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and are more readily available.
-A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect $\mathbf{E}$ to $\mathbf{B}$ and GIC to $\mathbf{E}$ and frequency dependent artifacts due to violations of the assumption that the field is spectrally uniform.
+A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect the electric field to the magnetic field and GIC to the electric field and frequency-dependent artifacts due to violations of the assumption that the field is spectrally uniform.
 
@@
 
-%When only a MT impedance tensor that connects $\mathbf{E}$ to $\mathbf{B}$ is available, a model of the form of Model~3, with the electric field estimated using $\mathbf{E}=\mathbf{Z}\mathbf{B}$ has been used~\citep{Marshall2013,Weigel2017,Bonner2017,Kelbert2017}.
+%When only an MT impedance tensor that connects $\mathbf{E}$ to $\mathbf{B}$ is available, a model of the form of Model~3, with the electric field estimated using $\mathbf{E}=\mathbf{Z}\mathbf{B}$ has been used~\citep{Marshall2013,Weigel2017,Bonner2017,Kelbert2017}.
 For the data set considered, significantly better GIC estimates can be made if frequency dependent system parameters are available.
@@
 for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the GIC data and providing feedback on the manuscript.
+We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response.
 %\clearpage
@@
 \caption{Frequency domain phase values for the parameters in Models 1, 2, and 4.
-The phase for $a_o$ and $b_o$ are constant either 0 or 180 degrees depending on the sign of their value, with positive values having a phase of 0.}
+The phase for $a_o$ and $b_o$ are constant and either 0$^o$ or 180$^o$ depending on the sign of their value, with positive values having a phase of 0$^o$.}
 \label{Phi}

### 2019-07-01 paper.tex
Updates from Overleaf

@@
 Geomagnetically induced currents (GICs) are electric currents in conducting systems that are due to electric fields near the Earth's surface.
-Time-varying electric and magnetic fields at Earth's surface are driven by time variations in ionospheric currents on time scales of minutes to hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
+Time-varying electric and magnetic fields at Earth's surface are driven by time variations in ionospheric currents on time scales of seconds to hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}.
 Of particular interest are currents induced in electric power systems because they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012,Gaunt2014}.
@@
 The results presented here are for the 1~Hz low-pass-filtered GIC measurements
-
-\begin{table}
-\caption{Date intervals used for analysis}
-\centering
-\begin{tabular}{l l}
-\hline
-Start Date & End Date\\
-\hline
-2006-04-03 & 2006-04-10\\
-2006-07-25 & 2006-07-29\\
-2006-08-05 & 2006-08-08\\
-2006-08-19 & 2006-08-21\\
-2006-11-07 & 2006-11-12\\
-2006-11-28 & 2006-11-29\\
-2006-12-01 & 2006-12-14\\
-2006-12-15 & 2006-12-15\\
-2007-11-18 & 2007-11-20\\
-2007-11-22 & 2007-11-22\\
-\hline
-\end{tabular}
-\label{intervals}
-\end{table}
 
@@
 Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ in a band of frequencies around $f_e$ from the DFT (discrete Fourier transform) of 24 hours of 1-s cadence data.
-Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
-The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation window has a lower bound of $1/86400$~Hz and highest evaluation window has an upper bound of $0.5$~Hz.
+Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of approximately $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
+The periods associated with the evaluation frequencies are listed in Table~\ref{evaluationperiods}.
+The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation frequency window has a lower bound of $1/86400$~Hz and highest evaluation frequency window has an upper bound of $0.375$~Hz.
 
@@
 
-The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted and it represents the fraction of the variance in the data that is predicted by the model.
+The mean-squared-error on day $D$ is
+\begin{equation*}
+\mbox{MSE}_m = \frac{1}{N}\sum_{t=t_o}^{t_f} \big(GIC(t)-GIC_m(t)\big)^2 \,,
+\end{equation*}
+\noindent
+where $GIC_m(t)$ is the model prediction of model $m$ given in Table~\ref{results} at time step $t$, $t_o=600$, and $t_f=85800$, so that sum is over one day of one-second-cadence measurements excluding the first and last 10~minutes.
 
-The average PE and MSE ratios were computed by averaging the out-of-sample PE and MSE ratios.
-The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the out-of-sample prediction metrics.
+The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is calculated for day $D$ using 
+\begin{equation*}
+\mbox{PE}_m = 1 - \frac{\sum_{t=t_o}^{t_f} \big(GIC(t)-GIC_m(t)\big)^2}{\sum_{t=t_o}^{t_f} \big(GIC(t)-\overline{GIC(t)}\big)^2}\,,
+\end{equation*}
+\noindent
+where $\overline{GIC(t)}$ is an average over the same time interval.
+PE is a measure of the skill of the model relative to the variance in the data that are predicted and it represents the fraction of the variance in the data that is predicted by the model.
+
+The average PE and the MSE ratios were computed by averaging the out-of-sample PEs and MSE ratios.
+The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using the bootstrap method \citep{Zoubir1998}: 1000 samples of the out-of-sample prediction metrics were used to compute the CIs.
 (The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.)
 
-The average signal-to-noise ratio at a given frequency is the average of the out-of-sample signal-to-noise ratios at that frequency.
-The spectra for the signal and prediction error used in the signal-to-noise ratio calculation were obtained by averaging the raw spectra in the same frequency windows used for the regression.
+The signal-to-noise ratio on day $D$ at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC$ to the smoothed power spectrum value of the prediction error at that evaluation frequency.
+The smoothed spectrum was obtained by averaging the raw spectrum in the same frequency windows used for the regression.
+The average signal-to-noise ratio at a given evaluation frequency is the average of the 35 out-of-sample signal-to-noise ratios at that frequency.
 
@@
 
-The primary conclusions of this work follow from the results shown in Table~~\ref{results}: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
+The primary conclusions of this work follow from the results shown in Table~\ref{results}: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
 
@@
 Figure~\ref{SN} shows the average signal-to-noise (i.e., signal to prediction error) ratios as a function of period for the four models.
-At each evaluation frequency, the SN ratio is the ratio of the average of the out-of-sample SN ratios of the segments and the error bars represent a 95\% CIs based on 1000 bootstrap samples of the segment SN ratios.
+At each evaluation frequency, the SN ratio is the ratio of the average of the out-of-sample SN ratios and the error bars represent a 95\% CIs based on 1000 bootstrap samples of the segment SN ratios.
 Consistent with the results in Table~\ref{results}, the SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies.
@@
 The ratio varies from 2.7 at $T=30$~s to 1.6 at $T=12$~h.
-At $T=12$~h, the highest period considered, the phases of $a(\omega)$ and $b(\omega)$ shown in Figure~\ref{Phi} differ by 180 degrees and the signs are consistent with those of $\overline{a}_o$ and $\overline{b}_o$.
 
@@
 The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120$^o$ while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.
+At $T=12$~h, the phases of $a(\omega)$ and $b(\omega)$ differ by 180 degrees and the associated signs are consistent with those of $\overline{a}_o$ and $\overline{b}_o$.
+
+%At $T=12$~h, the phases of $|a(\omega)|$ and $|b(\omega)|$ match that of $|\overline{b}_o|$ and $|\overline{a}_o|$ and the ratio of  $|b(\omega)|/|a(\omega)|$ of 1.6 is the closest it gets to that of $|\overline{b}_o|/|\overline{a}_o|$ of approximately 1.
+However, the absolute values of $a(\omega)$ and $b(\omega)$ are significantly larger than $a_o$ and $b_o$
+
+From Figure~\ref{Z} and Figure~\ref{Phi}, it follows that the 
+In Figure~\ref{Phi}, at $T=12$~h, the phase difference between $a(\omega)$ and $b(\omega)$ of 180 degrees is the same as that for $a_o$ and $b_o$.
+In addition the ratio of $|b(\omega)|/|a(\omega)| =1.6$ is nearest to the frequency independent value predicted by Model~1 of $|\overline{b}_o|/|\overline{a}_o| = 1$.
 
@@
 We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% higher than if 21 transfer functions were computed and then averaged.
+This result is unexpected, and extensive successful verification tests of our MT calculations along with the fact that this effect has observed by the co-author, who used different MT software on different data sets, suggests the explanation is not due to an error in the code used her to compute transfer functions.
+We expect to consider this in a future investigation as it suggests that stacking can be a significant source of transfer function bias and note that the most likely explanation involves non-stationarity in the noise.
 
@@
 In both Model~1 and Model~2, the input electric field is assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field.
-If it is not, the coefficients derived from the data in both models may differ significantly from those calculated using system configuration and resistance information.
+Spatial variations in the electric field measured on Earth’s surface can be due to horizontal variations in the incident electric field and horizontal variations in sub-surface conductivity, including those at land-ocean interfaces.
+If a spatial variation exists along the line in which GIC was measured, the coefficients derived from the data in both models may differ significantly from those calculated using system configuration and resistance information.
 However, the data-derived coefficients will not have a frequency dependence that results from a variation in the direction of the electric field along the line.
@@
 
+\section{Appendix}
+
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+
+\begin{table}[h]
+\caption{Date intervals used for analysis}
+\centering
+\begin{tabular}{l l}
+\hline
+Start Date & End Date\\
+\hline
+2006-04-03 & 2006-04-10\\
+2006-07-25 & 2006-07-29\\
+2006-08-05 & 2006-08-08\\
+2006-08-19 & 2006-08-21\\
+2006-11-07 & 2006-11-12\\
+2006-11-28 & 2006-11-29\\
+2006-12-01 & 2006-12-14\\
+2006-12-15 & 2006-12-15\\
+2007-11-18 & 2007-11-20\\
+2007-11-22 & 2007-11-22\\
+\hline
+\end{tabular}
+\label{intervals}
+\end{table}
+
+\begin{table}
+\caption{Periods associated with evaluation frequency windows used for regression and smoothing spectra.
+\# is the evaluation frequency number, $N$ is the number of DFT points in the window, and $T_e=1/f_e$.
+The window range is $T_l$ through $T_h$.}
+\centering
+\begin{tabular}{l l l l l}
+\hline \\
+\# & $N$ & $T_e$ [s] & $T_l$ [s] & $T_h$ [s] \\
+\hline \\
+1 & 2 & 43200 & 28800 & 86400 \\
+2 & 2 & 28800 & 21600 & 43200 \\
+3 & 3 & 21600 & 14400 & 43200 \\
+4 & 4 & 14400 & 9600 & 28800 \\
+5 & 5 & 10800 & 7200 & 21600 \\
+6 & 6 & 7854.5 & 5400 & 14400 \\
+7 & 8 & 5760 & 3927.3 & 10800 \\
+8 & 12 & 3927.3 & 2618.2 & 7854.5 \\
+9 & 16 & 2880 & 1920 & 5760 \\
+10 & 22 & 2009.3 & 1350 & 3927.3 \\
+11 & 31 & 1440 & 960 & 2880 \\
+12 & 43 & 1016.5 & 680.3 & 2009.3 \\
+13 & 61 & 720 & 480 & 1440 \\
+14 & 85 & 511.2 & 341.5 & 1016.5 \\
+15 & 120 & 361.5 & 241.3 & 720 \\
+16 & 170 & 255.6 & 170.4 & 511.2 \\
+17 & 240 & 180.8 & 120.5 & 361.5 \\
+18 & 338 & 128 & 85.4 & 255.6 \\
+19 & 478 & 90.5 & 60.3 & 180.8 \\
+20 & 676 & 64 & 42.7 & 128 \\
+21 & 956 & 45.2 & 30.2 & 90.5 \\
+22 & 1351 & 32 & 21.3 & 64 \\
+23 & 1910 & 22.6 & 15.1 & 45.2 \\
+24 & 2701 & 16 & 10.7 & 32 \\
+25 & 3819 & 11.3 & 7.5 & 22.6 \\
+26 & 5401 & 8 & 5.3 & 16 \\
+27 & 7638 & 5.7 & 3.8 & 11.3 \\
+28 & 10801 & 4 & 2.7 & 8 \\
+\hline \\
+\end{tabular}
+\label{evaluationperiods}
+\end{table}
+
+\clearpage
+
 \bibliography{paper.bib}

### 2019-07-02 paper.tex
Updates from Overleaf

@@
 
-Motivated by the results described in the following section in which Model~2 provides a significantly better representation of GIC than Model~1, we have also considered two additional modeling approaches that can be used to specify GIC.
-In these two models, a direct estimate of the geoelectric field is not needed; instead, the driver is the geomagnetic field, which is much more frequently measured, and as a result, these models may be of practical interest.
+Motivated by the results described in the following section, in which Model~2 provides a significantly better representation of GIC than Model~1, we have also considered two additional models that can be used to specify GIC.
+In these two models, a direct measurement of the geoelectric field is not used; instead, the driver is the geomagnetic field, which is much more frequently measured, and as a result, these models may be of practical interest.
 
@@
 The periods associated with the evaluation frequencies are listed in Table~\ref{evaluationperiods}.
-The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation frequency window has a lower bound of $1/86400$~Hz and highest evaluation frequency window has an upper bound of $0.375$~Hz.
+The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation frequency band has a lower bound of $1/86400$~Hz and highest evaluation frequency band has an upper bound of $0.375$~Hz.
 
@@
 When using these additional steps individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
-Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods higher than $\sim30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)
+Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods longer than $\sim30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)
 
@@
 
-The mean-squared-error on day $D$ is
+The mean-squared-error for model $m$ on day $D$ is
 \begin{equation*}
@@
 \noindent
-where $\overline{GIC(t)}$ is an average over the same time interval.
-PE is a measure of the skill of the model relative to the variance in the data that are predicted and it represents the fraction of the variance in the data that is predicted by the model.
+where $\overline{GIC(t)}$ is an average over the same time interval used in the summations.
+The PE is a measure of the skill of the model relative to the variance in the data that are predicted and it represents the fraction of the variance in the data that is predicted by the model.
 
@@
 The signal-to-noise ratio on day $D$ at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC$ to the smoothed power spectrum value of the prediction error at that evaluation frequency.
-The smoothed spectrum was obtained by averaging the raw spectrum in the same frequency windows used for the regression.
+The smoothed spectrum was obtained by averaging the raw spectrum in the same frequency bands used for the regression.
 The average signal-to-noise ratio at a given evaluation frequency is the average of the 35 out-of-sample signal-to-noise ratios at that frequency.
@@
 This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high-frequency fluctuations in the prediction error for Model~2 are the smallest among the models.
-For these periods, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a higher SN ratio than Model~3, which uses an indirect estimate of the electric field.
-Model~4 has a slightly higher SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the data-estimated coefficients in $\mathbf{Z}$.
+For these periods, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a larger SN ratio than Model~3, which uses an indirect estimate of the electric field.
+Model~4 has a slightly larger SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the data-estimated coefficients in $\mathbf{Z}$.
 
@@
 
-Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~\ref{results}, with Model~1 having the lowest and Model~4 having the highest SN ratio.
+Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~\ref{results}, with Model~1 having the lowest and Model~4 having the largest SN ratio.
 
@@
 
-From Figure~\ref{Z} and Figure~\ref{Phi}, it follows that the 
-In Figure~\ref{Phi}, at $T=12$~h, the phase difference between $a(\omega)$ and $b(\omega)$ of 180 degrees is the same as that for $a_o$ and $b_o$.
+%From Figure~\ref{Z} and Figure~\ref{Phi}, it follows that the In Figure~\ref{Phi}, at $T=12$~h, the phase difference between $a(\omega)$ and $b(\omega)$ of 180 degrees is the same as that for $a_o$ and $b_o$.
 In addition the ratio of $|b(\omega)|/|a(\omega)| =1.6$ is nearest to the frequency independent value predicted by Model~1 of $|\overline{b}_o|/|\overline{a}_o| = 1$.
 
-Figure~\ref{Z_MT} and~\ref{Phi_MT} show the coefficients for $\mathbf{Z}$ in Model~3.
+Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3.
 Note that electric and magnetic field data from Memambetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
@@
 Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function coefficients were computed using one-day segments of data and then averaged.
-We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% higher than if 21 transfer functions were computed and then averaged.
+We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged.
 This result is unexpected, and extensive successful verification tests of our MT calculations along with the fact that this effect has observed by the co-author, who used different MT software on different data sets, suggests the explanation is not due to an error in the code used her to compute transfer functions.
@@
 (3) The equation of Model~1 assumes that the electric field is spatially uniform at all points in the systems.
-The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the earth potential at the grounding points of the system, and simulations of GIC flow \citep{Albertson1981,Lehtinen1985} typically start with an assumed electric field along the lines to calculate these potential differences.
-The potential differences are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system.
+(An equation of this form can be computed in principle for an electric field with an arbitrary spatial distribution but most often the electric field is assumed to be spatially uniform).
+The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the electric field at all points of the system, and simulations of GIC flow typically start with an assumed electric field along the lines to calculate input voltages \citep{Albertson1981,Lehtinen1985, Horton2012}.
+The voltages are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system.
 In both Model~1 and Model~2, the input electric field is assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field.
-Spatial variations in the electric field measured on Earth’s surface can be due to horizontal variations in the incident electric field and horizontal variations in sub-surface conductivity, including those at land-ocean interfaces.
+Spatial variations in the electric field measured on Earth's surface can be due to horizontal variations in the incident electric field and horizontal variations in sub-surface conductivity, including those at land-ocean interfaces.
 If a spatial variation exists along the line in which GIC was measured, the coefficients derived from the data in both models may differ significantly from those calculated using system configuration and resistance information.
@@
 A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and are more readily available.
-A disadvantage is the transfer function coefficients are difficult to interpret physically because they contain components of the transfer functions that connect the electric field to the magnetic field and GIC to the electric field and frequency-dependent artifacts due to violations of the assumption that the field is spectrally uniform.
+A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect the electric field to the magnetic field and GIC to the electric field and frequency-dependent artifacts due to violations of the assumption that the field is spectrally uniform.
 
@@
 \begin{table}
-\caption{Periods associated with evaluation frequency windows used for regression and smoothing spectra.
+\caption{Periods associated with evaluation frequency bands used for regression and smoothing spectra.
 \# is the evaluation frequency number, $N$ is the number of DFT points in the window, and $T_e=1/f_e$.

### 2019-07-02 paper-diff.sh
add paper-diff.sh

@@ -1,0 +1,1 @@
+latexdiff paper-submitted-1.tex paper.tex > paper-diff.tex; git commit paper-diff.tex -m "Update diff"; git push

### 2019-07-10 paper.tex
Updates from Overleaf

@@
 That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency-independent power system coefficients.
-A first test of this assumption uses GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.
+A first test of this assumption is made using GIC measurements in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.
 A second model (Model 2) is obtained using a frequency domain generalization of Model 1: $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$.
 The coefficients $a(\omega)$ and $b(\omega)$ are computed from data and are shown to be frequency dependent and this model provides significantly better estimates of the measured GIC than Model~1.
-Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system coefficients may be explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters.
-It is also shown that further improvements over Model~2 can be made using a model with the geomagnetic field as an input.
+Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system coefficients may be explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system components.
+It is also shown that further improvements over Model~2 can be made using models with the geomagnetic field as an input.
 \end{abstract}
@@
 Then, we use a model in which the system coefficients are allowed to be frequency dependent, compute the frequency dependence, and compare this model's ability to represent GIC with the traditional frequency-independent model.
+Two additional models are considered that require only the geomagnetic field as an input and they are shown to be significantly better at estimating GIC than the models that use only the geoelectric field as an input.
 
@@
 The 1-second-cadence GIC data used span a total of 35 days across 10 intervals that start and end at midnight Japan Standard Time (JST) and are listed in Table~\ref{intervals}.
-These data have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
+These data along with a map showing the location of the transformer and the connected 187~kV power line have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
 These time intervals are those for which GIC data was made available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.
 The GIC dataset includes 1~s raw data and 1~Hz low-pass-filtered raw measurements.
-The results presented here are for the 1~Hz low-pass-filtered GIC measurements
+The results presented here are for the 1~Hz low-pass-filtered GIC measurements.
 
@@
 
-Data for the time span of 2006-08-04T15:00 through 2006-08-08T15:00 are shown Figure~\ref{sample}.
+Data for the time span of 2006-08-04T15:00Z through 2006-08-08T15:00Z are shown Figure~\ref{sample}.
 The GIC data were provided as 1-day files that start at midnight JST and the times shown in Figure~\ref{sample} are Universal Time.
@@
 In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$.
-For all of the time intervals considered, the fraction of interpolated electric field values ranged from 0--0.1\%.
+For the time intervals considered, the fraction of interpolated electric field values ranged from 0--0.1\%.
 
@@
 In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
-For all of the time intervals considered, the fraction of interpolated GIC values ranged from 1--4\%.
+For the time intervals considered, the fraction of interpolated GIC values ranged from 1--4\%.
 
@@
 
-In Model~2, the system coefficients $a(\omega)$ and $b(\omega)$ are frequency dependent according to
+In Model~2, the system coefficients $a(\omega)$ and $b(\omega)$ are frequency dependent according to (frequency-dependent variables are indicated by explicitly listing their dependence on $\omega$)
 
@@
 \noindent
-and are estimated using the Fourier transforms of  $GIC(t)$ and $\mathbf{E}(t)$; Frequency-dependent variables are indicated by explicitly listing their dependence on $\omega$.
-The method of parameter estimation and the generation of time-domain GIC estimates for Models~2-4 are given at the end of this section.
+and were estimated using the Fourier transforms of  $GIC(t)$ and $\mathbf{E}(t)$; the method of coefficient estimation and the generation of time-domain GIC estimates for this and the following models are given at the end of this section.
 
@@
 \noindent
-In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input variable.
+In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input measurements.
 
 Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ in a band of frequencies around $f_e$ from the DFT (discrete Fourier transform) of 24 hours of 1-s cadence data.
-Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of approximately $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
+Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
 The periods associated with the evaluation frequencies are listed in Table~\ref{evaluationperiods}.
@@
 
-Transfer function coefficients were computed for each of the segments.
+Transfer functions were computed for each of the 1-day segments.
 The average transfer function magnitude and phase at a given evaluation frequency was obtained by averaging the segment values at that evaluation frequency.
 
-To interpolate the frequency-domain transfer function coefficients estimated at the evaluation frequencies, their real and imaginary components are linearly interpolated individually onto the original DFT frequency grid (with a spacing of $1/86400$~Hz), and then the inverse DFT is computed.
-The interpolated transfer function coefficients were set to zero at frequencies higher than $0.25$~Hz and the transfer function estimate at $f=0$ was set to zero before interpolation.
+To interpolate the frequency-domain transfer function components estimated at the evaluation frequencies, their real and imaginary components were linearly interpolated individually onto the original DFT frequency grid (with a spacing of $1/86400$~Hz), and then the inverse DFT was computed.
+The interpolated transfer function coefficients were set to zero at frequencies higher than $0.375$~Hz and the transfer function estimate at $f=0$ was set to zero before interpolation.
 
@@
 Several modifications to this basic procedure were considered.
-Robust regression \citep{Egbert1986} with a Huber weighting constant of $1.345$ and an upper limit of 50 iterations and bi-square weighting with a constant of $4.685$ (both constants provide 95\% efficiency for normally distributed errors), both with no hard-cutoff step; time domain pre-whitening and tapering; and weighting the frequencies around each evaluation frequency using Parzen window weights.
+Robust regression \citep{Egbert1986} with a Huber weighting constant of $1.345$ and an upper limit of 50 iterations and bi-square weighting with a constant of $4.685$ (both constants provide 95\% efficiency for gaussian-distributed errors), both with no hard-cutoff step; time domain pre-whitening and tapering; and weighting the frequencies around each evaluation frequency using Parzen window weights.
 When using these additional steps individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
-Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods longer than $\sim30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)
+Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods longer than $\sim$30~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to a low signal-to-noise ratio.)
 
-\section{Results}
+\section{Model Evaluation}
 
-Out-of-sample estimation metrics for each model on day $D$ were determined by computing the average model coefficients using the remaining 34 days of data and then using these averaged coefficients to generate GIC time series and associated metrics on day $D$.
+Out-of-sample estimation quality metrics for each model on day $D$ were determined by computing the average model coefficients using the remaining 34 days of data and then using these averaged coefficients to generate GIC time series and associated metrics on day $D$.
 This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
@@
 
-The performance of each model in estimating GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.
-The PE and MSE ratios are shown in Table~\ref{results}.
+The performance of each model in estimating GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio (SN).
 
@@
 
+\section{Results}
+
+The PE and MSE ratios are shown in Table~\ref{results} and Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest.
+Note that the large errors visible at the start and end of the time intervals are due to the transients discussed above and the first and last 10~minutes of each interval were omitted from metrics calculations.
+
 \begin{table}
@@
 
-Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest.
-Note that the large errors visible at the start and end of the time intervals are due to the transients discussed above and the first and last 10~minutes of each interval were omitted from metrics calculations.
-
 The primary conclusions of this work follow from the results shown in Table~\ref{results}: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
 
-To determine the statistical significance of the differences, $10^5$ bootstrap averages of the segment PEs of each model were created and their differences were computed.
+To determine the statistical significance of the PE differences, $10^5$ bootstrap averages of the segment PEs of each model were created and their differences were computed.
 From this, the null hypothesis that the PE of Model~2 is equal to that of Model~1 can be rejected at a significance level of less than $10^{-5}$ because no bootstrap average PE difference from Model~2 was found to be lower than that for Model~1.
@@
 (The \cite{Fujii2015} transfer function was obtained from the IRIS DMC EMTF database [\cite{Kelbert2011}]).
-Although not critical to the issues addressed in this work, a preliminary investigation suggests that the cause is due to several factors.
+Although not critical to the conclusions in this work, a preliminary investigation suggests several explanations.
 We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015}.
@@
 In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used.
-Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function coefficients were computed using one-day segments of data and then averaged.
+Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function parameters were computed using one-day segments of data and then averaged.
 We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged.
-This result is unexpected, and extensive successful verification tests of our MT calculations along with the fact that this effect has observed by the co-author, who used different MT software on different data sets, suggests the explanation is not due to an error in the code used her to compute transfer functions.
-We expect to consider this in a future investigation as it suggests that stacking can be a significant source of transfer function bias and note that the most likely explanation involves non-stationarity in the noise.
+This result is unexpected, and extensive successful verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code was MT software distributed with MT hardware), suggests the explanation is not due to an error in the code used her to compute transfer functions.
+We expect to consider this in a future investigation as it suggests that stack averaging can be a significant source of transfer function bias and note that the most likely explanation involves non-stationarity in the noise.
 
@@
 (Bottom) GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
-The subscripts $x$ and $y$ correspond to Geographic North and East, respectively.}
+The subscripts $x$ and $y$ correspond to Geographic North and East, respectively, and the date labels are for Universal Time.}
 \label{sample}

### 2019-07-16 update.sh
cp map.pdf

@@ -1,0 +1,18 @@
+git pull
+
+DIR=../../figures/combined
+
+cp ../../map/map.pdf .
+cp $DIR/../20060805/plot_raw_All_20060805.pdf .
+cp ../../figures/predictions/plot_model_predictions-MeanModel-2006-08-06.pdf .
+
+cp $DIR/plot_model_summary_aobo_histograms-options-1.pdf .
+cp $DIR/plot_model_summary_Phi-options-1.pdf .
+cp $DIR/plot_model_summary_Z-options-1.pdf .
+cp $DIR/plot_model_summary_Phi_MT-options-1.pdf .
+cp $DIR/plot_model_summary_Z_MT-options-1.pdf .
+cp $DIR/plot_model_summary_SN-options-1.pdf .
+
+
+git commit -a -m "update figs"
+git push

### 2019-07-16 paper.tex
Updates from Overleaf

@@
 
-One-second-cadence electric and magnetic field measurements from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal.
+One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910 N, geodetic longitude: 144.189 E) were obtained from the Japan Meteorological Agency data portal.
 
@@
 \section{Models and Methods}
+\label{section:Models_and_Methods}
 
@@
 \section{Model Evaluation}
+\label{section:Model_Evaluation}
 
@@
 This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
-In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the estimate at the start of the interval for the amount of time that the causal part of the impulse response (found by Fourier inversion of the transfer function) is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.
+In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the estimate at the start of the interval for the amount of time that the causal part of the transfer function impulse responses (computed by Fourier inversion of the interpolated transfer function components) is non-zero and at the end for the amount of time that the acausal part of the impulse responses are non-zero.
 
@@
 \noindent
-where $GIC_m(t)$ is the model prediction of model $m$ given in Table~\ref{results} at time step $t$, $t_o=600$, and $t_f=85800$, so that sum is over one day of one-second-cadence measurements excluding the first and last 10~minutes.
+where $GIC_m(t)$ is the model prediction of model $m$ at time step $t$, $t_o=600$, and $t_f=85800$, so that sum is over one day of one-second-cadence measurements excluding the first and last 10~minutes.
+The MSE ratio for model $m$ is MSE$_1$/MSE$_m$.
 
@@
 
-The signal-to-noise ratio on day $D$ at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC$ to the smoothed power spectrum value of the prediction error at that evaluation frequency.
-The smoothed spectrum was obtained by averaging the raw spectrum in the same frequency bands used for the regression.
+The signal-to-noise ratio on day $D$ at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC(\omega_e)$ to the smoothed power spectrum value of the prediction error at that evaluation frequency.
+A smoothed spectrum was obtained by averaging the raw spectrum in the same frequency bands used for the regression described in Section~\ref{section:Models_and_Methods}.
 The average signal-to-noise ratio at a given evaluation frequency is the average of the 35 out-of-sample signal-to-noise ratios at that frequency.
@@
 
-The PE and MSE ratios are shown in Table~\ref{results} and Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest.
-Note that the large errors visible at the start and end of the time intervals are due to the transients discussed above and the first and last 10~minutes of each interval were omitted from metrics calculations.
+The PE and MSE ratios are shown in Table~\ref{results}.
+Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest.
+Note that the large errors visible at the start and end of the time intervals are due to the transients discussed in \ref{section:Model_Evaluation} and the first and last 10~minutes of each interval were omitted from metrics calculations.
 
 \begin{table}
-\caption{Out-of-sample model performance metrics for models $m=1-4$.
-MSE$_1$/MSE$_m$ is the ratio of the MSE for Model~1 to that of Model $m$.
+\caption{Out-of-sample model performance metrics for Model $m$.
+MSE$_1$/MSE$_m$ is the MSE ratio of Model~1 with respect to that of Model $m$.
 The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
@@
 Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3.
-Note that electric and magnetic field data from Memambetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
+Note that electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
 When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs.
@@
 
-To test the effect of different points on the line having different electric field spectra on the parameter estimates, we created a simulated GIC signal using electric field measurements made in Memambetsu (indicated with a superscript M) and Kakioka (superscript K), Japan:
+To test the effect of different points on the line having different electric field spectra on the parameter estimates, we created a simulated GIC signal using electric field measurements made in Memanbetsu (indicated with a superscript M) and Kakioka (superscript K), Japan:
 
@@
 Although this proof-of-concept simulation shows how variations in the frequency content of the geoelectric field on the line in which GIC is measured can cause the data-estimated system coefficients to have a frequency dependence, it is very crude as it uses several approximations that were necessitated by the availability of data.
-The distance from Kakioka to Memambetsu is $\sim$1400~km, which is $\sim$14x that of the length of the power line ($\sim$100~km) on which GIC was measured, and the correlations between the electric field from the two sites are likely to be lower than the fields measured at points separated by a distance closer to the length of the power line.
+The distance from Kakioka to Memanbetsu is $\sim$1400~km, which is $\sim$14x that of the length of the power line ($\sim$100~km) on which GIC was measured, and the correlations between the electric field from the two sites are likely to be lower than the fields measured at points separated by a distance closer to the length of the power line.
 In addition, the discontinuous change in the electric field implicit in the model implies that there would be a part of the line that can have a non-zero net charge and thus is not physically realizable for typical conductors.
@@
 \label{sample}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/map.pdf}
+\caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval.
+Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
+\label{predictions}
 \end{figure}

### 2019-07-16 paper.tex
Updates from Overleaf

@@
 The 1-second-cadence GIC data used span a total of 35 days across 10 intervals that start and end at midnight Japan Standard Time (JST) and are listed in Table~\ref{intervals}.
-These data along with a map showing the location of the transformer and the connected 187~kV power line have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
+These data have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
 These time intervals are those for which GIC data was made available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.
-The GIC dataset includes 1~s raw data and 1~Hz low-pass-filtered raw measurements.
-The results presented here are for the 1~Hz low-pass-filtered GIC measurements.
+The GIC dataset includes 1~s raw data and 1~Hz low-pass-filtered raw measurements and the results presented are for the 1~Hz low-pass-filtered GIC measurements.
+
+GIC was measured in a grounded neutral point of a Y-connected three-phase transformer connected to the 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
+The 187 kV line extends from the Ashro to Memanbetsu substation, which are both grounded.
+Memanbetsu is an end point of the line.
 
 One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910 N, geodetic longitude: 144.189 E) were obtained from the Japan Meteorological Agency data portal.
+The location of the observatory and substation are shown in Figure~\ref{map}.
 
@@
 for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the GIC data and providing feedback on the manuscript.
-We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response, Trevor Gaunt and Emanuel Benabeau for review and comments on part (1) of Section 5, and E.
-Josh Rigler for substantial feedback and review.
+We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response, Trevor Gaunt and Emanuel Benabeau for review and comments on part (1) of Section 5, and Josh Rigler for substantial feedback and review.
 %\clearpage
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{figures/map.pdf}
+\caption{Locations of measurement sites for GIC, at the Memanbetsu substation, and the geoelectric and geomagnetic field measurement site at Memambetsu Magnetic Observatory.
+The distance from the observatory to the substation is $\sim$9 km.}
+\label{map}
+\end{figure}
 
@@
 \caption{(Top) Geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
-(Middle) Geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB), which is $\sim$9 km from the Memanbetsu substation.
-(Bottom) GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
+(Middle) Geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB).
 The subscripts $x$ and $y$ correspond to Geographic North and East, respectively, and the date labels are for Universal Time.}
 \label{sample}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{figures/map.pdf}
-\caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval.
-Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
-\label{predictions}
 \end{figure}

### 2019-07-17 figures/update.sh
update figs

@@ -1,20 +1,0 @@
-git pull
-
-cp ../../figures/predictions/plot_model_predictions-MeanModel-2006-08-06.pdf .
-
-DIR=../../figures/combined
-cp $DIR/plot_model_summary_aobo_histograms-options-1.pdf .
-#cp $DIR/plot_model_summary_H-options-1.pdf .
-cp $DIR/plot_model_summary_Phi-options-1.pdf .
-cp $DIR/plot_model_summary_Z-options-1.pdf .
-cp $DIR/plot_model_summary_Phi_MT-options-1.pdf .
-cp $DIR/plot_model_summary_Z_MT-options-1.pdf .
-cp $DIR/plot_model_summary_SN-options-1.pdf .
-#cp $DIR/plot_options_comparison_a.pdf .
-#cp $DIR/plot_options_comparison_b.pdf .
-#cp $DIR/plot_options_comparison_zx.pdf .
-#cp $DIR/plot_options_comparison_zy.pdf .
-
-cp $DIR/../20060805/plot_raw_All_20060805.pdf .
-git commit -a -m "update figs"
-git push

### 2019-07-17 update.sh
scripts

@@ -1,18 +1,0 @@
-git pull
-
-DIR=../../figures/combined
-
-cp ../../map/map.pdf .
-cp $DIR/../20060805/plot_raw_All_20060805.pdf .
-cp ../../figures/predictions/plot_model_predictions-MeanModel-2006-08-06.pdf .
-
-cp $DIR/plot_model_summary_aobo_histograms-options-1.pdf .
-cp $DIR/plot_model_summary_Phi-options-1.pdf .
-cp $DIR/plot_model_summary_Z-options-1.pdf .
-cp $DIR/plot_model_summary_Phi_MT-options-1.pdf .
-cp $DIR/plot_model_summary_Z_MT-options-1.pdf .
-cp $DIR/plot_model_summary_SN-options-1.pdf .
-
-
-git commit -a -m "update figs"
-git push

### 2019-07-17 github-update.sh
update figs

@@ -1,5 +1,7 @@
 git pull
 
+cd figures
+
 DIR=../../figures/combined
 
 cp ../../map/map.pdf .

### 2019-07-17 github-update.sh
update figs

@@ -1,19 +1,17 @@
 git pull
 
-cd figures
-
 DIR=../../figures/combined
 
-cp ../../map/map.pdf .
-cp $DIR/../20060805/plot_raw_All_20060805.pdf .
-cp ../../figures/predictions/plot_model_predictions-MeanModel-2006-08-06.pdf .
+cp ../../map/map.pdf figures
+cp $DIR/../20060805/plot_raw_All_20060805.pdf figures
+cp ../../figures/predictions/plot_model_predictions-MeanModel-2006-08-06.pdf figures
 
-cp $DIR/plot_model_summary_aobo_histograms-options-1.pdf .
-cp $DIR/plot_model_summary_Phi-options-1.pdf .
-cp $DIR/plot_model_summary_Z-options-1.pdf .
-cp $DIR/plot_model_summary_Phi_MT-options-1.pdf .
-cp $DIR/plot_model_summary_Z_MT-options-1.pdf .
-cp $DIR/plot_model_summary_SN-options-1.pdf .
+cp $DIR/plot_model_summary_aobo_histograms-options-1.pdf figures
+cp $DIR/plot_model_summary_Phi-options-1.pdf figures
+cp $DIR/plot_model_summary_Z-options-1.pdf figures
+cp $DIR/plot_model_summary_Phi_MT-options-1.pdf figures
+cp $DIR/plot_model_summary_Z_MT-options-1.pdf figures
+cp $DIR/plot_model_summary_SN-options-1.pdf figures
 
 
 git commit -a -m "update figs"

### 2019-07-17 paper.tex

@@
 
-One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910 N, geodetic longitude: 144.189 E) were obtained from the Japan Meteorological Agency data portal.
+One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910$^o$ N, geodetic longitude: 144.189$^o$ E) were obtained from the Japan Meteorological Agency data portal.
 The location of the observatory and substation are shown in Figure~\ref{map}.
@@
 Magnetic and electric field measurements made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency Data Portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
+The geomagnetic coordinates in Figure~\ref{map} were computed using code from \url{https://www.ngdc.noaa.gov/geomag/geom_util/apex.shtml}.
 We would like to thank the Hokkaido Electric Power Co.
@@
 \caption{Locations of measurement sites for GIC, at the Memanbetsu substation, and the geoelectric and geomagnetic field measurement site at Memambetsu Magnetic Observatory.
-The distance from the observatory to the substation is $\sim$9 km.}
+The red line corresponds to the path of the 187 kV power line.
+The distance from the observatory to the substation is $\sim$9 km; dotted lines are Apex geomagnetic latitudes \cite{Richmond1995} in 2015.}
 \label{map}

### 2019-07-18 paper.tex
Updates from Overleaf

@@
 
-Motivated by the results described in the following section, in which Model~2 provides a significantly better representation of GIC than Model~1, we have also considered two additional models that can be used to specify GIC.
+Motivated by the results described in the following section, in which Model~2 provides a significantly better representation of GIC than Model~1, we have also considered two additional models that can be used to estimate GIC.
 In these two models, a direct measurement of the geoelectric field is not used; instead, the driver is the geomagnetic field, which is much more frequently measured, and as a result, these models may be of practical interest.
@@
 The periods associated with the evaluation frequencies are listed in Table~\ref{evaluationperiods}.
-The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation frequency band has a lower bound of $1/86400$~Hz and highest evaluation frequency band has an upper bound of $0.375$~Hz.
+The regression for each $f_e$ uses all DFT frequencies with values in a band with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest band has a lower bound of $1/86400$~Hz and highest band has an upper bound of $0.375$~Hz.
 
@@
 When using these additional steps individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
-Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods longer than $\sim$30~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to a low signal-to-noise ratio.)
+Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods longer than $\sim$30~s; At periods shorter than $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to a low signal-to-noise ratio.)
 
@@
 Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest.
-Note that the large errors visible at the start and end of the time intervals are due to the transients discussed in \ref{section:Model_Evaluation} and the first and last 10~minutes of each interval were omitted from metrics calculations.
+Note that the large errors visible at the start and end of the time intervals are due to the transients discussed in Section~\ref{section:Model_Evaluation} and the first and last 10~minutes of each interval were omitted from metrics calculations.
 
@@
 
-The primary conclusions of this work follow from the results shown in Table~\ref{results}: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
+The primary conclusions of this work follow from the results shown in Table~\ref{results}: the frequency dependent model, Model~2, is significantly better at estimating GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
 
@@
 
-Below $600$~s (10~min), Model~2 has the largest SN ratio.
+Below $600$~s, Model~2 has the largest SN ratio.
 This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high-frequency fluctuations in the prediction error for Model~2 are the smallest among the models.
@@
 
-Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors: (1) The measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement made 9~km away from the line - it follows that there is an implicit assumption in the model that the electric field is uniform over the length of the line on which GIC is measured; and (2) The error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
+Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors: (1) the measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement made 9~km away from the line - it follows that there is an implicit assumption in the model that the electric field is uniform over the length of the line on which GIC is measured; and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
 Issues related to this are discussed in the following section.
@@
 
-The average coefficients for Models~2 and~4 are shown in Figure~\ref{Z}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.
+The average coefficients for Models~2 and~4 are shown in Figure~\ref{Z}--\ref{Phi}. The error bars at each period correspond to the standard error of the 35 values used to compute the average.
 
@@
 We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged.
-This result is unexpected, and extensive successful verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code was MT software distributed with MT hardware), suggests the explanation is not due to an error in the code used her to compute transfer functions.
+This result is unexpected, and extensive successful verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code used was LEMI MT V1.4), suggests the explanation is not due to an error in the code used here to compute transfer functions.
 We expect to consider this in a future investigation as it suggests that stack averaging can be a significant source of transfer function bias and note that the most likely explanation involves non-stationarity in the noise.
@@
 The red line corresponds to the path of the 187 kV power line.
-The distance from the observatory to the substation is $\sim$9 km; dotted lines are Apex geomagnetic latitudes \cite{Richmond1995} in 2015.}
+The distance from the observatory to the substation is $\sim$9 km and dotted lines are Apex geomagnetic latitudes \citep{Richmond1995} in 2015.}
 \label{map}
@@
 Each set of coefficients was computed 35 times using one day of 1-second cadence data.
-The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 73 A/(V/km) and $b_o$ = -72 A/(V/km).}
+The circle or dot on the horizontal axis indicates the average of a histogram: $a_o$ = 73 A/(V/km) and $b_o$ = -72 A/(V/km).}
 \label{histogram}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Z_MT-options-1.pdf}
-\caption{Frequency domain transfer functions for the coefficients in Model~3.}
+\caption{Frequency domain transfer functions for the $\mathbf{Z}$ coefficients used in Model~3.}
 \label{Z_MT}
@@
 \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi_MT-options-1.pdf}
-\caption{Frequency domain phase values for the coefficients in Model~3.}
+\caption{Frequency domain phase values for the $\mathbf{Z}$ coefficients used in Model~3.}
 \label{Phi_MT}
@@
 \caption{Periods associated with evaluation frequency bands used for regression and smoothing spectra.
-\# is the evaluation frequency number, $N$ is the number of DFT points in the window, and $T_e=1/f_e$.
-The window range is $T_l$ through $T_h$.}
+\# is the evaluation frequency number, $N$ is the number of DFT points in the band, and $T_e\equv 1/f_e$.
+The band range is from $T_l$ through $T_h$.}
 \centering

### 2019-07-22 paper.tex
Updates from Overleaf

@@
 \section{Discussion}
+\label{discussion}
 
@@
 for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the GIC data and providing feedback on the manuscript.
-We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response, Trevor Gaunt and Emanuel Benabeau for review and comments on part (1) of Section 5, and Josh Rigler for substantial feedback and review.
+We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response, Trevor Gaunt and Emanuel Benabeau for review and comments on part (1) of Section\ref{discussion}, and Josh Rigler for substantial feedback and review.
 %\clearpage

### 2019-07-24 paper.tex
Updates from Overleaf

@@
 \section{Results}
+\label{results}
 
@@
 
-The average coefficients and 95\% CI for Model 1 are $a_o = 73$ $[62,83]$ A/(V/km) and $b_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}.
-(Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).
+The average coefficients and 95\% CI for Model 1 are $\overlin{a}_o = 73.0$ $[62,83]$ A/(V/km) and $\overline{b}_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}.
+(Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14 to 2006-12-15 JST time interval).
+
+Note that the values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction (\cite{Pulkkinen2010} and \cite{Watari2015} also found values of $a_o>0$ and $b_o<0$).
+However, as shown in Figure~\ref{map}, the 187 kV line which terminates at the Memanbetsu substation runs in the northeast direction, and hence the dominant contribution to the measured GIC would be from currents flowing in this line.
+If the electric field is the same at all points on the line, GIC in the grounded transformers at Memanbetsu substation should be the largest when the electric field points in the northeast direction.
 
@@
 For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+Note that transfer function shown in Figure~\ref{Z} differs significantly from that of a layered slab conductivity model for which $|Z_{xy}|=|Z_{yx}|$.
 
@@
 If a spatial variation exists along the line in which GIC was measured, the coefficients derived from the data in both models may differ significantly from those calculated using system configuration and resistance information.
-However, the data-derived coefficients will not have a frequency dependence that results from a variation in the direction of the electric field along the line.
+However, the data-derived coefficients will not have a frequency dependence that results only from the variation in the direction of the electric field along the line.
 
@@
 
+Note that in this discussion, we have attempted to address independently the contribution of spatial variations in the electric field magnitude and direction and spatial variations in the electric field spectra, but generally these variations do not occur independently as they are both caused by horizontal variations in the ground conductivity.
+Thus, the finding that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line (for both Model~1 and Model~2) is likely to be consistent with the frequency dependence of $a$ and $b$ found in Model~2.
+Testing this claim would require a full network analysis an knowledge of the ground conductivity throughout the network.
+
+%We have argued above that a spatial variation in the spectra of the electric field can explain the frequency dependence of $a$ and $b$. There is also evidence that the electric field varies along
+
+%A consideration of the ratio of $b_o/a_o$ suggests that the ground conductivity varies significantly along the length of the line.
+As shown in Figure~\ref{map}, the 187~kV line runs in a direction that is approximately 45$^o$ east of north.
+Assuming negligible GIC contributions from other unknown parts of the network and an electric field that is uniform along the length of the line, the equation for Model~1, $GIC(t) = a_oE_x(t) + b_oE_y(t)$, where $x$ is northward and $y$ is eastward,
+
+%one would expect the values of $a_o$ and $b_o$ to be such that $tan^{1}b_o/a_o$ is approximately 45$^o$ east of north .
+
 \section{Conclusions}
@@
 Although the use of frequency independent $a$ and $b$ parameter values estimated from data give a reasonable out-of-sample prediction model, it is not known how much data-estimated parameter values will differ from those directly calculated.
-(Because the details of the power system under consideration in this work are not available, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
+(Due to the limited availability of details of the power system and the ground conductivity along the line, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
 However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.

### 2019-07-25 paper.tex

@@
 
-Note that the values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction (\cite{Pulkkinen2010} and \cite{Watari2015} also found values of $a_o>0$ and $b_o<0$).
-However, as shown in Figure~\ref{map}, the 187 kV line which terminates at the Memanbetsu substation runs in the northeast direction, and hence the dominant contribution to the measured GIC would be from currents flowing in this line.
+Note that the values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction, which is approximately perpendicular to the 187~kV line that runs in the northeast direction.
+\cite{Pulkkinen2010} used magnetic field data at Memanbetsu and GIC data for a 3-day interval starting on 2005-12-26 and found a $b_o/a_o$ ratio of $-1.9/50$, corresponding to an angle of 2 degrees west of north; The $b_o/a_o = -7.4/38.1$ ratio documented in \cite{Watari2015} using data from a 2-day interval starting on 2006-12-14 that corresponds to an angle of 11 degrees west of north.
+
+%However, as shown in Figure~\ref{map}, the 187 kV line which terminates at the Memanbetsu substation runs in the northeast direction, and hence the dominant contribution to the measured GIC is expected be from currents flowing in this line.
 If the electric field is the same at all points on the line, GIC in the grounded transformers at Memanbetsu substation should be the largest when the electric field points in the northeast direction.
@@
 We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015}.
-After 2006, our calculated transfer functions differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%.
+After 2006, our calculated transfer function magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%.
 In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used.

### 2019-07-31 paper.tex
Updates from Overleaf

@@
 
-One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910$^o$ N, geodetic longitude: 144.189$^o$ E) were obtained from the Japan Meteorological Agency data portal.
+One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910$^{\circ}$ N, geodetic longitude: 144.189$^{\circ}$ E) were obtained from the Japan Meteorological Agency data portal.
 The location of the observatory and substation are shown in Figure~\ref{map}.
@@
     3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
-    4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
+    4\hspace{1em} $G_{B}(t) = z_xB_x(t) + z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
     \hline
@@
 
-Note that the values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction, which is approximately perpendicular to the 187~kV line that runs in the northeast direction.
-\cite{Pulkkinen2010} used magnetic field data at Memanbetsu and GIC data for a 3-day interval starting on 2005-12-26 and found a $b_o/a_o$ ratio of $-1.9/50$, corresponding to an angle of 2 degrees west of north; The $b_o/a_o = -7.4/38.1$ ratio documented in \cite{Watari2015} using data from a 2-day interval starting on 2006-12-14 that corresponds to an angle of 11 degrees west of north.
+The values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) and Model~1 imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction, which is approximately perpendicular to the 187~kV line that runs in the northeast direction.
+\cite{Pulkkinen2010} used magnetic field data at Memanbetsu and GIC data for a 3-day interval starting on 2005-12-26 and found a $b_o/a_o$ ratio of $-1.9/50$, corresponding to an angle of 2$^{\circ}$ west of north; The $b_o/a_o = -7.4/38.1$ ratio documented in \cite{Watari2015} using data from a 2-day interval starting on 2006-12-14 that corresponds to an angle of 11$^{\circ}$ west of north.
+
+%The values of $a_o$ and $b_o$ are highly sensitive to the method number of points used in the regression.
+The values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) were found using 86400 points to compute 35 pairs of $a_o$ and $b_o$ values.
+If 864 points are used to compute 3500 pairs of $a_o$ and $b_o$ values, $\overline{a}_o=+25.7$ and $\overline{b}_o=+15.2$ A/(V/km) and Model~1 predicts maximum GIC when the electric field is approximately 30 degrees east of north, in the approximate direction of the 187 kV line.
 
@@
 The ratio varies from 2.7 at $T=30$~s to 1.6 at $T=12$~h.
+Figure~\ref{Phi} shows the transfer function coefficient phase angles (that for Model~1 is 0$^{\circ}$ for positive coefficients and 180$^{\circ}$ for negative coefficients by definition).
+The phases for $a(\omega)$ and $b(\omega)$ are frequency dependent.
+At $T=12$~h, the phases of $a(\omega)$ and $b(\omega)$ differ by 180$^{\circ}$ and the associated signs are consistent with those of $\overline{a}_o > 0$ and $\overline{b}_o < 0$.
+At $T=12$~h, the values of $b(\omega)$ and $a(\omega)$ and Model~2 correspond to an electric field of 60$^{\circ}$ west of north being associated with the largest GIC.
 
-Figure~\ref{Phi} shows the transfer function phases (that for Model~1 is 0$^o$ for positive coefficients and 180$^o$ for negative coefficients by definition).
-The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120$^o$ while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.
-At $T=12$~h, the phases of $a(\omega)$ and $b(\omega)$ differ by 180 degrees and the associated signs are consistent with those of $\overline{a}_o$ and $\overline{b}_o$.
-
-%At $T=12$~h, the phases of $|a(\omega)|$ and $|b(\omega)|$ match that of $|\overline{b}_o|$ and $|\overline{a}_o|$ and the ratio of  $|b(\omega)|/|a(\omega)|$ of 1.6 is the closest it gets to that of $|\overline{b}_o|/|\overline{a}_o|$ of approximately 1.
-However, the absolute values of $a(\omega)$ and $b(\omega)$ are significantly larger than $a_o$ and $b_o$
-
-%From Figure~\ref{Z} and Figure~\ref{Phi}, it follows that the In Figure~\ref{Phi}, at $T=12$~h, the phase difference between $a(\omega)$ and $b(\omega)$ of 180 degrees is the same as that for $a_o$ and $b_o$.
-In addition the ratio of $|b(\omega)|/|a(\omega)| =1.6$ is nearest to the frequency independent value predicted by Model~1 of $|\overline{b}_o|/|\overline{a}_o| = 1$.
+Both Model~1 and 2 predict a maximum GIC when the electric field angle is significantly different from the angle of the 187~kV line.
+This can be caused by part other parts of the power network (the details of which are unavailable) making significant contributions to GIC and variations in ground conductivity.
 
@@
 
-Note that in this discussion, we have attempted to address independently the contribution of spatial variations in the electric field magnitude and direction and spatial variations in the electric field spectra, but generally these variations do not occur independently as they are both caused by horizontal variations in the ground conductivity.
-Thus, the finding that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line (for both Model~1 and Model~2) is likely to be consistent with the frequency dependence of $a$ and $b$ found in Model~2.
-Testing this claim would require a full network analysis an knowledge of the ground conductivity throughout the network.
-
-%We have argued above that a spatial variation in the spectra of the electric field can explain the frequency dependence of $a$ and $b$. There is also evidence that the electric field varies along
-
-%A consideration of the ratio of $b_o/a_o$ suggests that the ground conductivity varies significantly along the length of the line.
-As shown in Figure~\ref{map}, the 187~kV line runs in a direction that is approximately 45$^o$ east of north.
-Assuming negligible GIC contributions from other unknown parts of the network and an electric field that is uniform along the length of the line, the equation for Model~1, $GIC(t) = a_oE_x(t) + b_oE_y(t)$, where $x$ is northward and $y$ is eastward,
-
-%one would expect the values of $a_o$ and $b_o$ to be such that $tan^{1}b_o/a_o$ is approximately 45$^o$ east of north .
+In this discussion, we have attempted to address independently the contribution of spatial variations in the electric field magnitude and direction and spatial variations in the electric field spectra, but generally these variations do not occur independently as they are both caused by horizontal variations in the ground conductivity.
+Thus, the finding that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line for both Model~1 and Model~2 is likely to be consistent with the frequency dependence of $a$ and $b$ found in Model~2.
+However, testing this claim would require a full network analysis an knowledge of the ground conductivity throughout the network.
 
@@
 (Middle) Geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB).
-The subscripts $x$ and $y$ correspond to Geographic North and East, respectively, and the date labels are for Universal Time.}
+(Bottom) GIC measurements at Memanbetsu substation.
+The subscripts $x$ and $y$ correspond to Geographic North and East, respectively, the date labels are for Universal Time, and the average have been subtracted from each time series.}
   \label{sample}
@@
   \includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-2006-08-06.pdf}
-  \caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval.
+  \caption{Out-of-sample model predictions (top) and predictions errors for each model in a selected 1-day interval.
 Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
@@
   \caption{Frequency domain phase values for the coefficients in Models 1, 2, and 4.
-The phase for $a_o$ and $b_o$ are constant and either 0$^o$ or 180$^o$ depending on the sign of their value, with positive values having a phase of 0$^o$.}
+The phase for $a_o$ and $b_o$ are constant and either 0$^{\circ}$ or 180$^{\circ}$ depending on the sign of their value, with positive values having a phase of 0$^{\circ}$.}
   \label{Phi}

### 2019-07-31 paper.tex
New fig filenames

@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_raw_All_20060805.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_raw_All_20060805-v1.pdf}
   \caption{(Top) Geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-2006-08-06.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_GIC_predictions-MeanModel-2006-08-06-v0-o0.pdf}
   \caption{Out-of-sample model predictions (top) and predictions errors for each model in a selected 1-day interval.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_SN-options-1.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_SN-options-1-v0-o0.pdf}
   \caption{Out-of-sample signal-to-noise (signal to prediction error) ratios at the evaluation frequencies for the four models.}
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1-v0-o0.pdf}
   \caption{Distribution of coefficients in Model~1.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1-v0-o0.pdf}
   \caption{Frequency domain transfer functions for the coefficients in Models 1, 2, and 4.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1-v1-o0.pdf}
   \caption{Frequency domain phase values for the coefficients in Models 1, 2, and 4.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_Z_MT-options-1.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_Z_MT-options-1-v0-o0.pdf}
   \caption{Frequency domain transfer functions for the $\mathbf{Z}$ coefficients used in Model~3.}
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi_MT-options-1.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi_MT-options-1-v1-o1.pdf}
   \caption{Frequency domain phase values for the $\mathbf{Z}$ coefficients used in Model~3.}

### 2019-07-31 figures/update.sh
new figs

@@ -1,0 +1,15 @@
+git pull
+
+DIR=../../../figures
+
+cp $DIR/20060805/plot_raw_All_20060805-v1.pdf .
+cp $DIR/predictions/plot_GIC_predictions-MeanModel-2006-08-06-v0-o0.pdf .
+cp $DIR/summary/plot_model_summary_SN-options-1-v0-o0.pdf .
+cp $DIR/summary/plot_model_summary_aobo_histograms-options-1-v0-o0.pdf .
+cp $DIR/summary/plot_model_summary_Z-options-1-v0-o0.pdf .
+cp $DIR/summary/plot_model_summary_Phi-options-1-v1-o0.pdf .
+cp $DIR/summary/plot_model_summary_Phi_MT-options-1-v1-o0.pdf .
+cp $DIR/summary/plot_model_summary_Z_MT-options-1-v0-o0.pdf .
+
+#git commit -a -m "update figs"
+#git push

### 2019-07-31 paper-diff.sh
Update for filename change

@@ -1,1 +1,1 @@
-latexdiff paper-submitted-1.tex paper.tex > paper-diff.tex; git commit paper-diff.tex -m "Update diff"; git push
+latexdiff paper-submitted-1.tex paper-submitted-2.tex > paper-diff.tex; git commit paper-diff.tex -m "Update diff"; git push

### 2019-07-31 paper-submitted-2.tex

--- 2019-07-31 paper.tex
@@
 Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system coefficients may be explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system components.
-It is also shown that further improvements over Model~2 can be made using models with the geomagnetic field as an input.
+It is also shown that further improvements over Model~2 can be made using frequency-dependent models with the geomagnetic field as an input.
 \end{abstract}
@@
 Then, we use a model in which the system coefficients are allowed to be frequency dependent, compute the frequency dependence, and compare this model's ability to represent GIC with the traditional frequency-independent model.
-Two additional models are considered that require only the geomagnetic field as an input and they are shown to be significantly better at estimating GIC than the models that use only the geoelectric field as an input.
+Two additional frequency-dependent models are considered that require only the geomagnetic field as an input and they are shown to be significantly better at estimating GIC than the models that use only the geoelectric field as an input.
 
@@
 
-Data for the time span of 2006-08-04T15:00Z through 2006-08-08T15:00Z are shown Figure~\ref{sample}.
+Data for the time span of 2006-08-04T15:00Z through 2006-08-08T15:00Z are shown in Figure~\ref{sample}.
 The GIC data were provided as 1-day files that start at midnight JST and the times shown in Figure~\ref{sample} are Universal Time.
@@
 These spikes were removed by identifying times $t$ when the change in the 1~s absolute value of $GIC(t)$ was greater than $0.05$~Amps and then values in a window $GIC(t-2)$ through $GIC(t+100)$ were replaced using a linear interpolation of the values at $GIC(t-3)$ and $GIC(t + 101)$.
-In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.
+In the example interval shown in Figure~\ref{sample}, 493 spikes were removed and approximately 15 spikes that were were not removed by the automated algorithm are still visible.
 For the time intervals considered, the fraction of interpolated GIC values ranged from 1--4\%.
@@
 Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
-(However, \cite{Boteler1994} noted that for certain systems, the $L/R$ time constant may be large enough to warrant inclusion in DC flow calculations; this topic has more recently been pursued by \cite{Oyedokun2013a}.)
+(However, \cite{Boteler1994} noted that for certain systems, the $L/R$ time constant of the power network may be large enough to warrant inclusion in GIC flow calculations; this topic has more recently been pursued by \cite{Oyedokun2013a}.)
 
@@
 
-In this discussion, we have attempted to address independently the contribution of spatial variations in the electric field magnitude and direction and spatial variations in the electric field spectra, but generally these variations do not occur independently as they are both caused by horizontal variations in the ground conductivity.
+In the discussion of (3) above, we have attempted to address independently the contribution of spatial variations in the electric field magnitude and direction and spatial variations in the electric field spectra, but generally these variations do not occur independently as they are both caused by horizontal variations in the ground conductivity.
 Thus, the finding that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line for both Model~1 and Model~2 is likely to be consistent with the frequency dependence of $a$ and $b$ found in Model~2.
-However, testing this claim would require a full network analysis an knowledge of the ground conductivity throughout the network.
+However, testing this claim would require a full network analysis and knowledge of the ground conductivity throughout the network.
 
@@
 Local ground magnetic field estimates along with GIC measurements could be used to derive a transfer function model that is updated when the system configuration changes and this model could be used to estimate GIC under extreme or simulated geomagnetic conditions.
-A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and are more readily available.
+A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and more readily available than geoelectric field measurements.
 A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect the electric field to the magnetic field and GIC to the electric field and frequency-dependent artifacts due to violations of the assumption that the field is spectrally uniform.
@@
 \acknowledgments
-Magnetic and electric field measurements made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency Data Portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
+Magnetic and electric field measurements made at the Kakioka Magnetic Observatory were obtained from the Japan Meteorological Agency Data Portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}.
 The geomagnetic coordinates in Figure~\ref{map} were computed using code from \url{https://www.ngdc.noaa.gov/geomag/geom_util/apex.shtml}.
@@
 for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the GIC data and providing feedback on the manuscript.
-We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response, Trevor Gaunt and Emanuel Benabeau for review and comments on part (1) of Section\ref{discussion}, and Josh Rigler for substantial feedback and review.
+We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response, Trevor Gaunt and Emanuel Benabeau for review and comments on part (1) of Section~\ref{discussion}, and Josh Rigler for substantial feedback and review.
 %\clearpage
@@
 (Bottom) GIC measurements at Memanbetsu substation.
-The subscripts $x$ and $y$ correspond to Geographic North and East, respectively, the date labels are for Universal Time, and the average have been subtracted from each time series.}
+The subscripts $x$ and $y$ correspond to Geographic North and East, respectively, the date labels are for Universal Time, and the average has been subtracted from each time series.}
   \label{sample}
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi_MT-options-1-v1-o1.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi_MT-options-1-v1-o0.pdf}
   \caption{Frequency domain phase values for the $\mathbf{Z}$ coefficients used in Model~3.}

### 2019-08-01 paper-submitted-2.tex
Updates from Overleaf

@@
 \item Power system coefficients derived from a time series of GIC and geoelectric field measurements are frequency dependent.
-\item A GIC model with frequency-dependent system coefficients provides significantly better estimates of GIC than a frequency independent model.
-\item Data-derived power system coefficients may differ significantly with those computed using power system configuration information.
+-\item Data-derived power system coefficients may differ significantly with those computed using power system configuration information.
 \end{keypoints}
@@
 A second model (Model 2) is obtained using a frequency domain generalization of Model 1: $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$.
-The coefficients $a(\omega)$ and $b(\omega)$ are computed from data and are shown to be frequency dependent and this model provides significantly better estimates of the measured GIC than Model~1.
+The coefficients $a(\omega)$ and $b(\omega)$ are shown to be frequency dependent and this model provides significantly better estimates of the measured GIC than Model~1.
 Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system coefficients may be explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system components.
@@
 Then, we use a model in which the system coefficients are allowed to be frequency dependent, compute the frequency dependence, and compare this model's ability to represent GIC with the traditional frequency-independent model.
-Two additional frequency-dependent models are considered that require only the geomagnetic field as an input and they are shown to be significantly better at estimating GIC than the models that use only the geoelectric field as an input.
+Two additional frequency-dependent models are considered that require only the geomagnetic field as an input and they are shown to be significantly better at estimating GIC than the models that use the geoelectric field as an input.
 
@@
 GIC was measured in a grounded neutral point of a Y-connected three-phase transformer connected to the 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co.
-The 187 kV line extends from the Ashro to Memanbetsu substation, which are both grounded.
-Memanbetsu is an end point of the line.
+The 187 kV line extends from the Ashro substation to the Memanbetsu substation, which are both grounded.
+Memanbetsu is an end point of the 187 kV line.
 
@@
 The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ over a 1-day interval.
-\citep[][provided the mathematically equivalent closed-form equations for solving the matrix equation.]{Pulkkinen2007} The notation $\overline{a}_o$ and $\overline{b}_o$ will be used for the average of the $a_o$ and $b_o$ values computed using the 35 1-day window segments of available data.
+\citep[][provided the mathematically equivalent analytic equations for solving the matrix equation.]{Pulkkinen2007} The notation $\overline{a}_o$ and $\overline{b}_o$ will be used for the average of the $a_o$ and $b_o$ values computed using 35 1-day-window-segment regressions.
 
@@
 \noindent
-and were estimated using the Fourier transforms of  $GIC(t)$ and $\mathbf{E}(t)$; the method of coefficient estimation and the generation of time-domain GIC estimates for this and the following models are given at the end of this section.
+and were estimated using the Fourier transforms of  $GIC(t)$, $E_x(t)$, and $E_y(t)$; the method of coefficient estimation and the generation of time-domain GIC estimates for this and the following models are given at the end of this section.
 
 Motivated by the results described in the following section, in which Model~2 provides a significantly better representation of GIC than Model~1, we have also considered two additional models that can be used to estimate GIC.
-In these two models, a direct measurement of the geoelectric field is not used; instead, the driver is the geomagnetic field, which is much more frequently measured, and as a result, these models may be of practical interest.
+In these two models, a direct measurement of the geoelectric field is not used; instead, the driver is the geomagnetic field, which is much more often measured, and as a result, these models may be of practical interest.
 
-In Model~3, the coefficients $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field, $\mathbf{B}$, and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ and
+In Model~3, previously computed coefficients $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field, $\mathbf{B}$, and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ and
 
@@
 \noindent
-where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$.
-That is, instead of using the measured electric field directly as in Model~2, a magnetotelluric transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.
+where, explicitly, the electric field components are $E'_x(\omega)$ = $Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_y(\omega)$ and $E'_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
+That is, instead of using the measured electric field directly as in Model~2, a magnetotelluric transfer function that relates $\mathbf{E}(\omega)$ to $\mathbf{B}(\omega)$ is used to provide an estimate of the electric field based on the magnetic field.
 
@@
 \noindent
-In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input measurements.
+In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$, $B_x(t)$, and $B_y(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input measurements.
 
@@
 The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}.
+
 For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ are solved for using
@@
 
-To interpolate the frequency-domain transfer function components estimated at the evaluation frequencies, their real and imaginary components were linearly interpolated individually onto the original DFT frequency grid (with a spacing of $1/86400$~Hz), and then the inverse DFT was computed.
+To interpolate the frequency-domain transfer function components that are estimated at the evaluation frequencies onto the original DFT frequency grid (with a spacing of $1/86400$~Hz), their real and imaginary components were linearly interpolated individually.
 The interpolated transfer function coefficients were set to zero at frequencies higher than $0.375$~Hz and the transfer function estimate at $f=0$ was set to zero before interpolation.
 
-GIC time series were calculated by multiplication of the interpolated transfer function coefficients with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(f)=a^I(f)E_x(f)+b^I(f)E_y(f)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(f)$ was calculated to give the estimate of $G_E(t)$ on a 1~s time grid).
+GIC time series were calculated by multiplication of the interpolated transfer function coefficients with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(\omega)=a^I(\omega)E_x(\omega)+b^I(\omega)E_y(\mega)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(\omega)$ was calculated to give the estimate of $G_E(t)$ on a 1~s time grid).
 
 Several modifications to this basic procedure were considered.
-Robust regression \citep{Egbert1986} with a Huber weighting constant of $1.345$ and an upper limit of 50 iterations and bi-square weighting with a constant of $4.685$ (both constants provide 95\% efficiency for gaussian-distributed errors), both with no hard-cutoff step; time domain pre-whitening and tapering; and weighting the frequencies around each evaluation frequency using Parzen window weights.
-When using these additional steps individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
-Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods longer than $\sim$30~s; At periods shorter than $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to a low signal-to-noise ratio.)
+Robust regression \citep{Egbert1986} with either a Huber weighting using a constant of $1.345$ or bi-square weighting using a constant of $4.685$ (both with no hard-cutoff step).
+The commonly used constants provide 95\% efficiency for gaussian-distributed errors.
+Models were also computed using time domain pre-whitening (using a 10th-order Yule-Walker model computed using the model output), time domain tapering (using a Parzen window), and weighting the frequencies around each evaluation frequency using Parzen window weights.
+When using these additional procedures individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
+Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal-to-noise ratio is above $\sim$3 (corresponding to periods longer than $\sim$30~s); at periods shorter than $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to a low signal-to-noise ratio.)
 
@@
 \noindent
-where $GIC_m(t)$ is the model prediction of model $m$ at time step $t$, $t_o=600$, and $t_f=85800$, so that sum is over one day of one-second-cadence measurements excluding the first and last 10~minutes.
+where $GIC_m(t)$ is the estimate from model $m$ at time step $t$;  $t_o=600$ and $t_f=85800$, so that sum is over one day of one-second-cadence measurements excluding the first and last 10~minutes.
 The MSE ratio for model $m$ is MSE$_1$/MSE$_m$.
@@
 where $\overline{GIC(t)}$ is an average over the same time interval used in the summations.
-The PE is a measure of the skill of the model relative to the variance in the data that are predicted and it represents the fraction of the variance in the data that is predicted by the model.
+The PE is a measure of the skill of the model with respect to the variance in the predictand and it represents the fraction of the variance in the data that is predicted by the model (when PE~$\ge 0)$.
 
 The average PE and the MSE ratios were computed by averaging the out-of-sample PEs and MSE ratios.
-The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using the bootstrap method \citep{Zoubir1998}: 1000 samples of the out-of-sample prediction metrics were used to compute the CIs.
-(The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.)
+The 95\% confidence interval, CI, on the PE and MSE ratios were calculated using the bootstrap method \citep{Zoubir1998}: 1000 random samples (with replacement) of 17 of the 35 out-of-sample prediction metrics were used to compute the CIs.
+The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.
 
@@
 
-The PE and MSE ratios are shown in Table~\ref{results}.
-Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest.
+The PE and MSE ratios are shown in Table~\ref{resultstable}.
+Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{resultstable}, i.e., Model~1 has the lowest PE and Model~4 has the highest.
 Note that the large errors visible at the start and end of the time intervals are due to the transients discussed in Section~\ref{section:Model_Evaluation} and the first and last 10~minutes of each interval were omitted from metrics calculations.
@@
   \end{tabular}
-  \label{results}
+  \label{resultstable}
 \end{table}
 
-The primary conclusions of this work follow from the results shown in Table~\ref{results}: the frequency dependent model, Model~2, is significantly better at estimating GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
+The primary conclusions of this work follow from the results shown in Table~\ref{resultstable}: the frequency dependent model, Model~2, is significantly better at estimating GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
 
@@
 At each evaluation frequency, the SN ratio is the ratio of the average of the out-of-sample SN ratios and the error bars represent a 95\% CIs based on 1000 bootstrap samples of the segment SN ratios.
-Consistent with the results in Table~\ref{results}, the SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies.
+Consistent with the results in Table~\ref{resultstable}, the SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies.
 However, ordering of the SN ratio for Models~2-4 is dependent on period.
@@
 
-Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~\ref{results}, with Model~1 having the lowest and Model~4 having the largest SN ratio.
+Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~\ref{resultstable}, with Model~1 having the lowest and Model~4 having the largest SN ratio.
 
@@
 For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).
-Note that transfer function shown in Figure~\ref{Z} differs significantly from that of a layered slab conductivity model for which $|Z_{xy}|=|Z_{yx}|$.
 
@@
 The ratio varies from 2.7 at $T=30$~s to 1.6 at $T=12$~h.
+
 Figure~\ref{Phi} shows the transfer function coefficient phase angles (that for Model~1 is 0$^{\circ}$ for positive coefficients and 180$^{\circ}$ for negative coefficients by definition).
@@
 
-Both Model~1 and 2 predict a maximum GIC when the electric field angle is significantly different from the angle of the 187~kV line.
-This can be caused by part other parts of the power network (the details of which are unavailable) making significant contributions to GIC and variations in ground conductivity.
-
+Both Model~1 and~2 predict a maximum GIC when the electric field angle is significantly different from the angle of the 187~kV line.
+As discussed in the following section, this can be caused by other parts of the power network (the details of which are unavailable) making significant contributions to GIC and variations in ground conductivity.
 Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3.
-Note that electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
+Note that transfer function shown in Figure~\ref{Z_MT} differs significantly from that of a layered slab conductivity model for which $|Z_{xy}|=|Z_{yx}|$.
+That electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
 When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs.
 0.78).
-(The \cite{Fujii2015} transfer function was obtained from the IRIS DMC EMTF database [\cite{Kelbert2011}]).
 Although not critical to the conclusions in this work, a preliminary investigation suggests several explanations.
-We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015}.
-After 2006, our calculated transfer function magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%.
+We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015} and the fact that we did not use a remote reference in our calculations.
+After 2006 (January through May tested) our calculated transfer function magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%.
 In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used.
@@
 We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged.
-This result is unexpected, and extensive successful verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code used was LEMI MT V1.4), suggests the explanation is not due to an error in the code used here to compute transfer functions.
-We expect to consider this in a future investigation as it suggests that stack averaging can be a significant source of transfer function bias and note that the most likely explanation involves non-stationarity in the noise.
+This result is unexpected, and results from extensive verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code used was LEMI MT V1.4), suggests the explanation is not due to an error in the code used to compute the transfer functions in this work.
+We expect to consider this in a future investigation as it suggests that stack averaging can be a significant source of transfer function bias and expect that the explanation involves non-stationarity in the noise.
 
@@
 Thus, the finding that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line for both Model~1 and Model~2 is likely to be consistent with the frequency dependence of $a$ and $b$ found in Model~2.
-However, testing this claim would require a full network analysis and knowledge of the ground conductivity throughout the network.
+However, testing this claim would require a full network analysis and knowledge of the ground conductivity throughout the network (\cite{Uyeshima2001} has shown significant variations in the impedance tensor $\mathb{Z}$ on Hokkaido Island).
 
@@
 The geomagnetic coordinates in Figure~\ref{map} were computed using code from \url{https://www.ngdc.noaa.gov/geomag/geom_util/apex.shtml}.
-We would like to thank the Hokkaido Electric Power Co.
+The \cite{Fujii2015} transfer function was obtained from the IRIS DMC EMTF database [\cite{Kelbert2011}].
+We thank the Hokkaido Electric Power Co.
 for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the GIC data and providing feedback on the manuscript.
@@
   \includegraphics[width=\textwidth]{figures/map.pdf}
-  \caption{Locations of measurement sites for GIC, at the Memanbetsu substation, and the geoelectric and geomagnetic field measurement site at Memambetsu Magnetic Observatory.
+  \caption{Locations of measurement sites for GIC, at the Memanbetsu substation, and the geoelectric and geomagnetic field measurement site at Memanbetsu Magnetic Observatory.
 The red line corresponds to the path of the 187 kV power line.

### 2019-08-01 paper-submitted-2.tex
Updates from Overleaf

@@
 \item Power system coefficients derived from a time series of GIC and geoelectric field measurements are frequency dependent.
--\item Data-derived power system coefficients may differ significantly with those computed using power system configuration information.
+\item A GIC model with frequency-dependent system coefficients provides significantly better estimates of GIC than a frequency independent model.
+\item Data-derived power system coefficients may differ significantly with those computed using power system configuration information.
 \end{keypoints}
@@
 Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{resultstable}, i.e., Model~1 has the lowest PE and Model~4 has the highest.
-Note that the large errors visible at the start and end of the time intervals are due to the transients discussed in Section~\ref{section:Model_Evaluation} and the first and last 10~minutes of each interval were omitted from metrics calculations.
+The large errors visible at the start and end of the time intervals are due to the transients discussed in Section~\ref{section:Model_Evaluation} and the first and last 10~minutes of each interval were omitted from metrics calculations.
 
@@
 Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition).
-For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).
+For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).
 
@@
 Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3.
-Note that transfer function shown in Figure~\ref{Z_MT} differs significantly from that of a layered slab conductivity model for which $|Z_{xy}|=|Z_{yx}|$.
+Note that transfer function shown in Figure~\ref{Z_MT} differs significantly from that of a layered slab conductivity model for which $Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$.
 That electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
@@
 Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function parameters were computed using one-day segments of data and then averaged.
-We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged.
+We have found that if a continuous span of measurements over 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged.
 This result is unexpected, and results from extensive verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code used was LEMI MT V1.4), suggests the explanation is not due to an error in the code used to compute the transfer functions in this work.
-We expect to consider this in a future investigation as it suggests that stack averaging can be a significant source of transfer function bias and expect that the explanation involves non-stationarity in the noise.
+This result suggests that stack averaging can be a significant source of transfer function bias and we expect that the explanation involves variations in the signal-to-noise between segments.
 
@@
   \caption{Out-of-sample model predictions (top) and predictions errors for each model in a selected 1-day interval.
-Note that the scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
+The scales for the error time series differ from that for the top panel so that the error amplitudes are scaled up by a factor of 2.}
   \label{predictions}
@@
 Each set of coefficients was computed 35 times using one day of 1-second cadence data.
-The circle or dot on the horizontal axis indicates the average of a histogram: $a_o$ = 73 A/(V/km) and $b_o$ = -72 A/(V/km).}
+The circle or dot on the horizontal axis indicates the average of a histogram and $\overline{a}_o$ = 73 A/(V/km) and $\overline{b}_o$ = -72 A/(V/km).}
   \label{histogram}

### 2019-08-02 paper-submitted-2.tex
Updates from Overleaf

@@
 One-second-cadence electric and magnetic field measurements from Memanbetsu Magnetic Observatory (geodetic latitude: 43.910$^{\circ}$ N, geodetic longitude: 144.189$^{\circ}$ E) were obtained from the Japan Meteorological Agency data portal.
-The location of the observatory and substation are shown in Figure~\ref{map}.
+The locations of the observatory and substations are shown in Figure~\ref{map}.
 
@@
 where, explicitly, the electric field components are $E'_x(\omega)$ = $Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_y(\omega)$ and $E'_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
-That is, instead of using the measured electric field directly as in Model~2, a magnetotelluric transfer function that relates $\mathbf{E}(\omega)$ to $\mathbf{B}(\omega)$ is used to provide an estimate of the electric field based on the magnetic field.
+That is, instead of using the measured electric field directly as in Model~2, a transfer function that relates $\mathbf{E}(\omega)$ to $\mathbf{B}(\omega)$ is used to provide an estimate of the electric field based on the magnetic field.
 
@@
 \noindent
-In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$, $B_x(t)$, and $B_y(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input measurements.
+In this model, the transfer function components $z_x(\omega)$ and $z_y(\omega)$ are determined directly from measurements of $GIC(t)$, $B_x(t)$, and $B_y(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input measurements.
 
-Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ in a band of frequencies around $f_e$ from the DFT (discrete Fourier transform) of 24 hours of 1-s cadence data.
+Estimates of $a(\omega)$ and $b(\omega)$ in Model~2 are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS).
+At each evaluation frequency, a regression is performed using a set $G_E(\omega)$, $E_x(\omega)$, and $E_y(\omega)$ values from the DFT (discrete Fourier transform) of 24 hours of 1-s cadence data in a band of frequencies around $\omega_e$.
 Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used.
@@
 
-The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}.
+The same approach is used for solving for $z_x(\omega_e)$ and $z_y(\omega_e)$ in Equation~\ref{model4}.
 
-For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ are solved for using
+For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ at the evaluation frequencies are solved for using the same method used for Model~2 using the equation
 
@@
   \begin{equation*}
-    E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)
+    E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)\,.
   \end{equation*}
@@
 Transfer functions were computed for each of the 1-day segments.
-The average transfer function magnitude and phase at a given evaluation frequency was obtained by averaging the segment values at that evaluation frequency.
+The average transfer function magnitude and phase at a given evaluation frequency were obtained by averaging the segment values at that evaluation frequency.
 
@@
 
-GIC time series were calculated by multiplication of the interpolated transfer function coefficients with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(\omega)=a^I(\omega)E_x(\omega)+b^I(\omega)E_y(\mega)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(\omega)$ was calculated to give the estimate of $G_E(t)$ on a 1~s time grid).
+GIC time series were calculated by multiplication of the interpolated transfer function coefficients with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(\omega)=a^I(\omega)E_x(\omega)+b^I(\omega)E_y(\omega)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(\omega)$ was calculated to give the estimate of $G_E(t)$ on a 1~s time grid).
 
 Several modifications to this basic procedure were considered.
-Robust regression \citep{Egbert1986} with either a Huber weighting using a constant of $1.345$ or bi-square weighting using a constant of $4.685$ (both with no hard-cutoff step).
-The commonly used constants provide 95\% efficiency for gaussian-distributed errors.
+Robust regression \citep{Egbert1986} with either a Huber weighting using a constant of $1.345$ or bi-square weighting using a constant of $4.685$, both with no hard-cutoff step.
+The commonly used constants provide 95\% efficiency for Gaussian-distributed errors.
 Models were also computed using time domain pre-whitening (using a 10th-order Yule-Walker model computed using the model output), time domain tapering (using a Parzen window), and weighting the frequencies around each evaluation frequency using Parzen window weights.
-When using these additional procedures individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
-Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal-to-noise ratio is above $\sim$3 (corresponding to periods longer than $\sim$30~s); at periods shorter than $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to a low signal-to-noise ratio.)
+When using these additional procedures individually, we find that the conclusions made regarding the differences in each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method.
+Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal-to-noise ratio is above $\sim$3 (corresponding to periods longer than $\sim$30~s); at periods shorter than $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to a low signal-to-noise ratio.
 
@@
 
-The mean-squared-error for model $m$ on day $D$ is
+The mean-squared-error for model $m$ is
 \begin{equation*}
@@
 
-The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is calculated for day $D$ using
+The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is calculated using
 \begin{equation*}
@@
 
-The signal-to-noise ratio on day $D$ at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC(\omega_e)$ to the smoothed power spectrum value of the prediction error at that evaluation frequency.
+The frequency dependent signal-to-noise ratio for each 1-day segment at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC(\omega_e)$ to the smoothed power spectrum value of the prediction error at that evaluation frequency.
 A smoothed spectrum was obtained by averaging the raw spectrum in the same frequency bands used for the regression described in Section~\ref{section:Models_and_Methods}.
@@
     \hline
-    1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
     2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
@@
 This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high-frequency fluctuations in the prediction error for Model~2 are the smallest among the models.
-For these periods, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a larger SN ratio than Model~3, which uses an indirect estimate of the electric field.
+Below $600$~s, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a larger SN ratio than Model~3, which uses an indirect estimate of the electric field.
 Model~4 has a slightly larger SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the data-estimated coefficients in $\mathbf{Z}$.
 
-Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors: (1) the measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement made 9~km away from the line - it follows that there is an implicit assumption in the model that the electric field is uniform over the length of the line on which GIC is measured; and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.
+Above $600$~s, the ordering of the SN is not consistent with these expectations and is possibly explained by two factors: (1) the measured GIC is based on an integrated electric field along the power line, and the electric field used in Model~2 is a point measurement made 9~km away from the line - it follows that there is an implicit assumption in the model that the electric field is uniform over the length of the line on which GIC is measured; and (2) the error associated with this assumption has less of an influence on the errors in a model that uses point measurements of the magnetic field.
 Issues related to this are discussed in the following section.
@@
 
-The average coefficients and 95\% CI for Model 1 are $\overlin{a}_o = 73.0$ $[62,83]$ A/(V/km) and $\overline{b}_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}.
+The average coefficients and 95\% CI for Model 1 are $\overline{a}_o = 73.0$ $[62,83]$ A/(V/km) and $\overline{b}_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}.
 (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14 to 2006-12-15 JST time interval).
 
-The values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) and Model~1 imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points in the northwest direction, which is approximately perpendicular to the 187~kV line that runs in the northeast direction.
-\cite{Pulkkinen2010} used magnetic field data at Memanbetsu and GIC data for a 3-day interval starting on 2005-12-26 and found a $b_o/a_o$ ratio of $-1.9/50$, corresponding to an angle of 2$^{\circ}$ west of north; The $b_o/a_o = -7.4/38.1$ ratio documented in \cite{Watari2015} using data from a 2-day interval starting on 2006-12-14 that corresponds to an angle of 11$^{\circ}$ west of north.
+The values of $\overline{a}_o = 73$ A/(V/km) and $\overline{b}_o = -72$ A/(V/km) and Model~1 imply that for a given electric field magnitude, GIC is largest at the measurement site when the electric field points 45$^{\circ}$ west of north, which is approximately perpendicular to the 187~kV line shown in Figure~\ref{map} that runs in the northeast direction.
+\cite{Pulkkinen2010} used magnetic field data at Memanbetsu and GIC data for a 3-day interval starting on 2005-12-26 and found a $b_o/a_o$ ratio of $-1.9/50$, corresponding to a maximum GIC when the electric field angle is 2$^{\circ}$ west of north; The $b_o/a_o = -7.4/38.1$ ratio documented in \cite{Watari2015} using data from a 2-day interval starting on 2006-12-14 that corresponds to an angle of 11$^{\circ}$ west of north.
 
@@
 
-Figure~\ref{Phi} shows the transfer function coefficient phase angles (that for Model~1 is 0$^{\circ}$ for positive coefficients and 180$^{\circ}$ for negative coefficients by definition).
+Figure~\ref{Phi} shows the transfer function coefficient phase angles.
+For Model~1, the phase angle is 0$^{\circ}$ for positive coefficients and 180$^{\circ}$ for negative coefficients by definition.
 The phases for $a(\omega)$ and $b(\omega)$ are frequency dependent.
@@
 Both Model~1 and~2 predict a maximum GIC when the electric field angle is significantly different from the angle of the 187~kV line.
-As discussed in the following section, this can be caused by other parts of the power network (the details of which are unavailable) making significant contributions to GIC and variations in ground conductivity.
-Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3.
-Note that transfer function shown in Figure~\ref{Z_MT} differs significantly from that of a layered slab conductivity model for which $Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$.
+As discussed in the following section, this can be caused by other parts of the power network (the details of which are unavailable) making significant contributions to GIC and differences in ground conductivity in parts of the network that contribute to GIC.
+
+Figure~\ref{Z_MT} and~\ref{Phi_MT} show the magnitude and phase for the coefficients of $\mathbf{Z}$ in Model~3.
+Note that transfer function coefficients shown in Figure~\ref{Z_MT} differ significantly from that of a layered slab conductivity model for which $Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$.
 That electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
@@
 Although not critical to the conclusions in this work, a preliminary investigation suggests several explanations.
-We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015} and the fact that we did not use a remote reference in our calculations.
-After 2006 (January through May tested) our calculated transfer function magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%.
+We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function coefficients match the \cite{Fujii2015} coefficients to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015} and the fact that we did not use a remote reference in our calculations.
+However, using one-month intervals after 2006 (January through May tested) our calculated transfer function coefficient magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%.
 In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used.
-Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer function parameters were computed using one-day segments of data and then averaged.
-We have found that if a continuous span of measurements over 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged.
+Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer functions were computed using one-day segments of data and then averaged.
+We have found that if a continuous span of measurements over 21 days is used to compute a single transfer function, the transfer function coefficient magnitudes are up to 25\% larger than if 21 transfer functions were computed and then averaged.
 This result is unexpected, and results from extensive verification tests of our MT calculations along with the fact that this effect has observed by both authors using two different MT calculation codes on different data sets (the other code used was LEMI MT V1.4), suggests the explanation is not due to an error in the code used to compute the transfer functions in this work.
@@
 Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is typically $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala2005}, the reactance will be $\sim$10 times smaller or less than the resistance at GIC frequencies of $\sim$1~Hz and lower.
-Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
+Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth to more accurately model skin an proximity effects also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
 (However, \cite{Boteler1994} noted that for certain systems, the $L/R$ time constant of the power network may be large enough to warrant inclusion in GIC flow calculations; this topic has more recently been pursued by \cite{Oyedokun2013a}.)
@@
 (2) A second consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence.
-However, the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range and the response characteristics of the magnetic and electric field instruments are essentially frequency independent relative to the observed transfer function frequency dependence for periods in the range of $30-10^{5}$~s (Oowada et al.
-1998, Fujii 2019 personal communication.)
+However, the GIC instrument (HIOKI 9279 current sensor) is frequency independent to within $\pm 0.2$\% in this period range and the response characteristics of the magnetic and electric field instruments are essentially frequency independent relative to the observed transfer function frequency dependence for periods in the range of $30-10^{5}$~s (\cite{Oowada1998}; Fujii 2019 personal communication.)
 
 (3) The equation of Model~1 assumes that the electric field is spatially uniform at all points in the systems.
-(An equation of this form can be computed in principle for an electric field with an arbitrary spatial distribution but most often the electric field is assumed to be spatially uniform).
+(An equation of this form can be computed in principle for a static electric field with an arbitrary spatial distribution but most often the electric field is assumed to be spatially uniform).
 The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the electric field at all points of the system, and simulations of GIC flow typically start with an assumed electric field along the lines to calculate input voltages \citep{Albertson1981,Lehtinen1985, Horton2012}.
-The voltages are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system.
-In both Model~1 and Model~2, the input electric field is assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field.
+The electric potential differences are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system.
+In both Model~1 and Model~2, the input electric field is typically assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field.
 Spatial variations in the electric field measured on Earth's surface can be due to horizontal variations in the incident electric field and horizontal variations in sub-surface conductivity, including those at land-ocean interfaces.
-If a spatial variation exists along the line in which GIC was measured, the coefficients derived from the data in both models may differ significantly from those calculated using system configuration and resistance information.
-However, the data-derived coefficients will not have a frequency dependence that results only from the variation in the direction of the electric field along the line.
+If a spatial variation exists along the line in which GIC was measured, the coefficients derived from measurements in both models may differ significantly from those calculated using system configuration and resistance information.
+However, the data-derived coefficients will not have a frequency dependence that results only from the variation in the direction of a static electric field along the line.
 
@@
 
-To test the effect of different points on the line having different electric field spectra on the parameter estimates, we created a simulated GIC signal using electric field measurements made in Memanbetsu (indicated with a superscript M) and Kakioka (superscript K), Japan:
+To test the effect of different points on the line having different electric field spectra on the coefficient estimates, we created a simulated GIC signal using electric field measurements made in Memanbetsu (indicated with a superscript M) and Kakioka (superscript K), Japan:
 
@@
 
-Although the use of frequency independent $a$ and $b$ parameter values estimated from data give a reasonable out-of-sample prediction model, it is not known how much data-estimated parameter values will differ from those directly calculated.
+Although the use of frequency independent $a$ and $b$ coefficients estimated from data give a reasonable estimates of GIC, it is not known how much data-estimated coefficient values will differ from those directly calculated.
 (Due to the limited availability of details of the power system and the ground conductivity along the line, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
@@
 When GIC and magnetic field measurements are available, a model of the form of Model~4 can be used.
-For this particular data set, we have found that such a model provides the best estimates of GIC in terms of the PE and MSE metrics.
+For this particular data set, we have found that such a model provides the best estimates of GIC in terms of the PE, MSE, and signal-to-error metrics.
 Local ground magnetic field estimates along with GIC measurements could be used to derive a transfer function model that is updated when the system configuration changes and this model could be used to estimate GIC under extreme or simulated geomagnetic conditions.
 A key advantage of this approach is that long-term geomagnetic field measurements are generally easier to make and more readily available than geoelectric field measurements.
-A disadvantage is the transfer function parameters are difficult to interpret physically because they contain components of the transfer functions that connect the electric field to the magnetic field and GIC to the electric field and frequency-dependent artifacts due to violations of the assumption that the field is spectrally uniform.
+A disadvantage is the transfer function coefficients of Model~4 are difficult to interpret physically because they contain components of the transfer functions that connect the electric field to the magnetic field and GIC to the electric field.
 
@@
 The red line corresponds to the path of the 187 kV power line.
-The distance from the observatory to the substation is $\sim$9 km and dotted lines are Apex geomagnetic latitudes \citep{Richmond1995} in 2015.}
+The distance from the observatory to the Memanbetsu substation is $\sim$9 km and dotted lines are Apex geomagnetic latitudes \citep{Richmond1995} in 2015.}
   \label{map}

### 2019-08-04 paper-submitted-2.tex
Updates from Overleaf

@@
 In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression when contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.,][]{Boteler2014a,Boteler2014b,Boteler2017}.
-This quasi-DC assumption has been used or stated in many GIC-related works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015}.
+This quasi-DC assumption has been used or stated in many GIC-related works, for example, \citet{Pulkkinen2007,Wik2008,Pirjola2009,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015}.
 
@@
 
-In Model~2, the system coefficients $a(\omega)$ and $b(\omega)$ are frequency dependent according to (frequency-dependent variables are indicated by explicitly listing their dependence on $\omega$)
+In Model~2, the system coefficients $a(\omega)$ and $b(\omega)$ are frequency dependent (frequency-dependent variables are indicated by explicitly listing their dependence on $\omega$):
 
@@
   \begin{equation}
-    G_E(\omega) = a(\omega)E_x(\omega) + b(\omega)E_y(\omega)
+    G_E(\omega) = a(\omega)E_x(\omega) + b(\omega)E_y(\omega)\,,
     \label{model2}
@@
 \noindent
-and were estimated using the Fourier transforms of  $GIC(t)$, $E_x(t)$, and $E_y(t)$; the method of coefficient estimation and the generation of time-domain GIC estimates for this and the following models are given at the end of this section.
+and were estimated using the Fourier transforms of $GIC(t)$ and $\mathbf{E}(t)$; the method of coefficient estimation and the generation of time-domain GIC estimates for this and the following models are given at the end of this section.
 
@@
   \begin{equation}
-    G_{E'}(\omega) = a(\omega)E'_x(\omega) + b(\omega)E'_y(\omega)
+    G_{E'}(\omega) = a(\omega)E'_x(\omega) + b(\omega)E'_y(\omega)\,,
     \label{model3}
@@
 \noindent
-where, explicitly, the electric field components are $E'_x(\omega)$ = $Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_y(\omega)$ and $E'_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
+where the electric field components are $E'_x(\omega)$ = $Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_y(\omega)$ and $E'_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.
 That is, instead of using the measured electric field directly as in Model~2, a transfer function that relates $\mathbf{E}(\omega)$ to $\mathbf{B}(\omega)$ is used to provide an estimate of the electric field based on the magnetic field.
@@
 \noindent
-In this model, the transfer function components $z_x(\omega)$ and $z_y(\omega)$ are determined directly from measurements of $GIC(t)$, $B_x(t)$, and $B_y(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input measurements.
+In this model, the transfer function components $z_x(\omega)$ and $z_y(\omega)$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input measurements.
 
@@
 The periods associated with the evaluation frequencies are listed in Table~\ref{evaluationperiods}.
-The regression for each $f_e$ uses all DFT frequencies with values in a band with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest band has a lower bound of $1/86400$~Hz and highest band has an upper bound of $0.375$~Hz.
 
@@
 
-Out-of-sample estimation quality metrics for each model on day $D$ were determined by computing the average model coefficients using the remaining 34 days of data and then using these averaged coefficients to generate GIC time series and associated metrics on day $D$.
+Out-of-sample estimation quality metrics for each model on day $D$ were determined by computing the average model coefficients using the other 34 days of data and then using these average coefficients to generate GIC time series and associated metrics on day $D$.
 This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
@@
 Note that transfer function coefficients shown in Figure~\ref{Z_MT} differ significantly from that of a layered slab conductivity model for which $Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$.
-That electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
+The electric and magnetic field data from Memanbetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$.
 When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs.
@@
 
-(3) The equation of Model~1 assumes that the electric field is spatially uniform at all points in the systems.
-(An equation of this form can be computed in principle for a static electric field with an arbitrary spatial distribution but most often the electric field is assumed to be spatially uniform).
-The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the electric field at all points of the system, and simulations of GIC flow typically start with an assumed electric field along the lines to calculate input voltages \citep{Albertson1981,Lehtinen1985, Horton2012}.
-The electric potential differences are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system.
+(3) The coefficients in Model~1 are determined assuming that the electric field is spatially uniform at all points in the system.
+The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the electric field at all points of the system, and simulations of GIC flow typically start with an assumed electric field along the lines to calculate input voltages \citep{Albertson1981,Lehtinen1985, Pirjola2009}.
+The electric fields are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system.
 In both Model~1 and Model~2, the input electric field is typically assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field.
@@
 Violation of this assumption can lead to an apparent violation of the quasi-static approximation.
-Suppose the electric field and GIC are measured at one end of a long line segment and the electric field is non-zero only at a frequency of $\omega_1$.
+Consider an electric field and GIC that are measured at one end of a long line segment with the electric field being non-zero only at a frequency of $\omega_1$.
 If the electric field is spectrally uniform along the full line segment, then from the quasi-static approximation it follows that the measured GIC will also only be non-zero at a frequency of $\omega_1$.
-If the electric field on the far half of the line is non-zero only at a frequency of $\omega_2$, the observed GIC and electric field will appear to be inconsistent with the quasi-static approximation because GIC will have non-zero variations at both $\omega_1$ and $\omega_2$.
+If the electric field at the other end of the line is non-zero only at a frequency of $\omega_2$, the observed GIC and electric field will appear to be inconsistent with the quasi-static approximation because GIC will have non-zero variations at both $\omega_1$ and $\omega_2$.
 
@@
 In the discussion of (3) above, we have attempted to address independently the contribution of spatial variations in the electric field magnitude and direction and spatial variations in the electric field spectra, but generally these variations do not occur independently as they are both caused by horizontal variations in the ground conductivity.
-Thus, the finding that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line for both Model~1 and Model~2 is likely to be consistent with the frequency dependence of $a$ and $b$ found in Model~2.
-However, testing this claim would require a full network analysis and knowledge of the ground conductivity throughout the network (\cite{Uyeshima2001} has shown significant variations in the impedance tensor $\mathb{Z}$ on Hokkaido Island).
+A time-varying ionospheric current can also cause spatial and spectral variations in the electric field magnitude \citep{Boteler1998}.
+Thus, the finding in Section~\ref{results} that the angle of the electric field that produces maximum GIC differs significantly from the angle of the 187~kV power line for both Model~1 and Model~2 may be consistent with the frequency dependence of $a$ and $b$ found in Model~2.
+However, testing this claim would require a full network analysis, knowledge of the ground conductivity throughout the network, and accounting for the driving ionospheric currents.
+
+%(\cite{Uyeshima2001} has shown significant variations in the impedance tensor $\mathb{Z}$ on Hokkaido Island)
 
@@
 (Due to the limited availability of details of the power system and the ground conductivity along the line, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
-However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Model~2.
+However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Models~2 and 3.
 
@@
   \caption{Periods associated with evaluation frequency bands used for regression and smoothing spectra.
-\# is the evaluation frequency number, $N$ is the number of DFT points in the band, and $T_e\equv 1/f_e$.
+\# is the evaluation frequency number, $N$ is the number of DFT points in the band, and $T_e\equiv 1/f_e$.
 The band range is from $T_l$ through $T_h$.}

### 2019-08-05 paper-submitted-2.tex
Updates from Overleaf

@@
 The coefficients $a(\omega)$ and $b(\omega)$ are shown to be frequency dependent and this model provides significantly better estimates of the measured GIC than Model~1.
-Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system coefficients may be explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system components.
+Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system coefficients may be explained by spatial variations in the spectrum of the geoelectric field over the spatial extent of the power system.
 It is also shown that further improvements over Model~2 can be made using frequency-dependent models with the geomagnetic field as an input.

### 2019-08-05 paper-submitted-2.tex
Updates from Overleaf

@@
 \item Power system coefficients derived from a time series of GIC and geoelectric field measurements are frequency dependent.
-\item A GIC model with frequency-dependent system coefficients provides significantly better estimates of GIC than a frequency independent model.
+\item A GIC model with frequency-dependent system coefficients provides significantly better estimates of GIC than a frequency-independent model.
 \item Data-derived power system coefficients may differ significantly with those computed using power system configuration information.
@@
 A second model (Model 2) is obtained using a frequency domain generalization of Model 1: $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$.
-The coefficients $a(\omega)$ and $b(\omega)$ are shown to be frequency dependent and this model provides significantly better estimates of the measured GIC than Model~1.
+The coefficients $a(\omega)$ and $b(\omega)$ are shown to be frequency-dependent and this model provides significantly better estimates of the measured GIC than Model~1.
 Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system coefficients may be explained by spatial variations in the spectrum of the geoelectric field over the spatial extent of the power system.
@@
 These data have been presented previously in \citet{Watari2009} and \cite{Watari2015}.
-These time intervals are those for which GIC data was made available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.
+These time intervals are those for which GIC data was made available except for three days for which there were obvious problems with either the electric field or GIC measurements.
 The GIC dataset includes 1~s raw data and 1~Hz low-pass-filtered raw measurements and the results presented are for the 1~Hz low-pass-filtered GIC measurements.
@@
 The 187 kV line extends from the Ashro substation to the Memanbetsu substation, which are both grounded.
-Memanbetsu is an end point of the 187 kV line.
+Memanbetsu is an end-point of the 187 kV line.
 
@@
 
-The frequency dependent signal-to-noise ratio for each 1-day segment at a given evaluation frequency is the ratio of of the smoothed power spectrum value of $GIC(\omega_e)$ to the smoothed power spectrum value of the prediction error at that evaluation frequency.
+The frequency-dependent signal-to-noise ratio for each 1-day segment at a given evaluation frequency is the ratio of the smoothed power spectrum value of $GIC(\omega_e)$ to the smoothed power spectrum value of the prediction error at that evaluation frequency.
 A smoothed spectrum was obtained by averaging the raw spectrum in the same frequency bands used for the regression described in Section~\ref{section:Models_and_Methods}.
@@
   \caption{Out-of-sample model performance metrics for Model $m$.
-MSE$_1$/MSE$_m$ is the MSE ratio of Model~1 with respect to that of Model $m$.
+MSE$_1$/MSE$_m$ is the MSE ratio of Model~1 to that of Model $m$.
 The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
@@
 
-The primary conclusions of this work follow from the results shown in Table~\ref{resultstable}: the frequency dependent model, Model~2, is significantly better at estimating GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
+The primary conclusions of this work follow from the results shown in Table~\ref{resultstable}: the frequency-dependent model, Model~2, is significantly better at estimating GIC than the frequency-independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.
 
@@
 However, using one-month intervals after 2006 (January through May tested) our calculated transfer function coefficient magnitudes differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%.
-In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used.
+Also, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used.
 Due to the large gaps between and the varying length of continuous segments of data used in this work, transfer functions were computed using one-day segments of data and then averaged.
@@
 Clarified, this is essentially a statement that the reactance is proportional to frequency and although the reactance is typically $\sim$5 times larger than the resistance at 50~Hz \citep{Purchala2005}, the reactance will be $\sim$10 times smaller or less than the resistance at GIC frequencies of $\sim$1~Hz and lower.
-Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth to more accurately model skin an proximity effects also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
+Given these arguments, the quasi-DC approximation is justified, and an explanation involving a violation of the assumption of a uniform half-space conducting slab Earth using a conductivity that varies with depth to more accurately model skin and proximity effects also seems unlikely to explain the results as it would need to account for at least an order of magnitude of reactance.
 (However, \cite{Boteler1994} noted that for certain systems, the $L/R$ time constant of the power network may be large enough to warrant inclusion in GIC flow calculations; this topic has more recently been pursued by \cite{Oyedokun2013a}.)
@@
 
-Although the use of frequency independent $a$ and $b$ coefficients estimated from data give a reasonable estimates of GIC, it is not known how much data-estimated coefficient values will differ from those directly calculated.
+Although the use of frequency-independent $a$ and $b$ coefficients estimated from data give reasonable estimates of GIC, it is not known how much data-estimated coefficient values will differ from those directly calculated.
 (Due to the limited availability of details of the power system and the ground conductivity along the line, we were not able to make this comparison.) Works that use Model~1 and generally note a high degree of visual consistency between the measured values and predictions using this model (e.g., the top panel of Figure~\ref{predictions}).
-However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency dependent model of the form of Models~2 and 3.
+However, we have shown using contemporaneous GIC and electric field measurements that a significantly better estimation of GIC can be obtained by using a frequency-dependent model of the form of Models~2 and 3.
 

### 2019-09-20 paper-submitted-3-nosubdir.tex
Updates from Overleaf

--- 2019-08-05 paper-submitted-2.tex
@@
 \usepackage{amsmath}
-\usepackage{xcolor}
-\usepackage[colorlinks]{hyperref}
-\usepackage[colorinlistoftodos]{todonotes}
+%\usepackage{xcolor}
+%\usepackage[colorlinks]{hyperref}
+%\usepackage[colorinlistoftodos]{todonotes}
 
@@
 
-\title{A Test of the Frequency Independence Assumption of Power System Coefficients used in Geomagnetically Induced Current Estimates}
+\title{An Evaluation of the Frequency Independence Assumption of Power System Coefficients used in Geomagnetically Induced Current Estimates}
 
@@
 
-Estimates of $a(\omega)$ and $b(\omega)$ in Model~2 are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS).
+Estimates of $a(\omega)$ and $b(\omega)$ in Model~2 are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing an ordinary least squares regression (OLS).
 At each evaluation frequency, a regression is performed using a set $G_E(\omega)$, $E_x(\omega)$, and $E_y(\omega)$ values from the DFT (discrete Fourier transform) of 24 hours of 1-s cadence data in a band of frequencies around $\omega_e$.
@@
 This process was repeated to create a set of 35 out-of-sample prediction metrics for each model.
-In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the estimate at the start of the interval for the amount of time that the causal part of the transfer function impulse responses (computed by Fourier inversion of the interpolated transfer function components) is non-zero and at the end for the amount of time that the acausal part of the impulse responses are non-zero.
+In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients at the start and end of the interval.
 
@@
 
-Figure~\ref{Z_MT} and~\ref{Phi_MT} show the magnitude and phase for the coefficients of $\mathbf{Z}$ in Model~3.
+Figure~\ref{Z_MT} and~\ref{Phi_MT} show the magnitude and phase for the coefficients of $\mathbf{Z}$ used in Model~3.
 Note that transfer function coefficients shown in Figure~\ref{Z_MT} differ significantly from that of a layered slab conductivity model for which $Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$.
@@
 The electric fields are then used along with the DC system component resistances and their connectivity to calculate GIC in the different parts of the system.
-In both Model~1 and Model~2, the input electric field is typically assumed to be spatially uniform along the line in which GIC was measured so that the potential difference is directly proportional to the electric field.
+In both Model~1 and Model~2, the input electric field is typically assumed to be spatially uniform along the line in which GIC was measured so that the induced voltage is directly proportional to the electric field.
 Spatial variations in the electric field measured on Earth's surface can be due to horizontal variations in the incident electric field and horizontal variations in sub-surface conductivity, including those at land-ocean interfaces.
@@
 
-The primary implications of the results presented are (1) significantly improved estimates of GICs using data-derived system coefficients $a$ and $b$ can be made if they are allowed to be frequency dependent and (2) a direct comparison of data-derived $a$ and $b$ coefficients and those computed using power system configuration information will be problematic if the data are not consistent with the assumptions of Equation~\ref{model1}.
+The primary implications of the results presented are (1) significantly improved estimates of GICs using data-derived system coefficients $a$ and $b$ can be made if they are allowed to be frequency dependent and (2) $a$ and $b$ coefficients derived using a traditional frequency-independent model may differ significantly from those computed using power system configuration information.
+
+%a direct comparison of data-derived $a$ and $b$ coefficients and those computed using power system configuration information will be problematic if the data are not consistent with the frequency independent assumption.
 
@@
 for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the GIC data and providing feedback on the manuscript.
-We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response, Trevor Gaunt and Emanuel Benabeau for review and comments on part (1) of Section~\ref{discussion}, and Josh Rigler for substantial feedback and review.
+We also thank Ikuko Fujii for providing clarifications on the geoelectric instrument response, Trevor Gaunt and Emanuel Bernabeau for review and comments on part (1) of Section~\ref{discussion}, and Josh Rigler for substantial feedback and review.
 %\clearpage
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/map.pdf}
+  \includegraphics[width=\textwidth]{map.pdf}
   \caption{Locations of measurement sites for GIC, at the Memanbetsu substation, and the geoelectric and geomagnetic field measurement site at Memanbetsu Magnetic Observatory.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_raw_All_20060805-v1.pdf}
+  \includegraphics[width=\textwidth]{plot_raw_All_20060805-v1.pdf}
   \caption{(Top) Geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_GIC_predictions-MeanModel-2006-08-06-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{plot_GIC_predictions-MeanModel-2006-08-06-v0-o0.pdf}
   \caption{Out-of-sample model predictions (top) and predictions errors for each model in a selected 1-day interval.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_SN-options-1-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{plot_model_summary_SN-options-1-v0-o0.pdf}
   \caption{Out-of-sample signal-to-noise (signal to prediction error) ratios at the evaluation frequencies for the four models.}
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{plot_model_summary_aobo_histograms-options-1-v0-o0.pdf}
   \caption{Distribution of coefficients in Model~1.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{plot_model_summary_Z-options-1-v0-o0.pdf}
   \caption{Frequency domain transfer functions for the coefficients in Models 1, 2, and 4.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1-v1-o0.pdf}
+  \includegraphics[width=\textwidth]{plot_model_summary_Phi-options-1-v1-o0.pdf}
   \caption{Frequency domain phase values for the coefficients in Models 1, 2, and 4.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_Z_MT-options-1-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{plot_model_summary_Z_MT-options-1-v0-o0.pdf}
   \caption{Frequency domain transfer functions for the $\mathbf{Z}$ coefficients used in Model~3.}
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi_MT-options-1-v1-o0.pdf}
+  \includegraphics[width=\textwidth]{plot_model_summary_Phi_MT-options-1-v1-o0.pdf}
   \caption{Frequency domain phase values for the $\mathbf{Z}$ coefficients used in Model~3.}

### 2019-09-20 paper-submitted-3.tex
Updates from Overleaf

--- 2019-09-20 paper-submitted-3-nosubdir.tex
@@
   \centering
-  \includegraphics[width=\textwidth]{map.pdf}
+  \includegraphics[width=\textwidth]{figures/map.pdf}
   \caption{Locations of measurement sites for GIC, at the Memanbetsu substation, and the geoelectric and geomagnetic field measurement site at Memanbetsu Magnetic Observatory.
@@
   \centering
-  \includegraphics[width=\textwidth]{plot_raw_All_20060805-v1.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_raw_All_20060805-v1.pdf}
   \caption{(Top) Geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB).
@@
   \centering
-  \includegraphics[width=\textwidth]{plot_GIC_predictions-MeanModel-2006-08-06-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_GIC_predictions-MeanModel-2006-08-06-v0-o0.pdf}
   \caption{Out-of-sample model predictions (top) and predictions errors for each model in a selected 1-day interval.
@@
   \centering
-  \includegraphics[width=\textwidth]{plot_model_summary_SN-options-1-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_SN-options-1-v0-o0.pdf}
   \caption{Out-of-sample signal-to-noise (signal to prediction error) ratios at the evaluation frequencies for the four models.}
@@
   \centering
-  \includegraphics[width=\textwidth]{plot_model_summary_aobo_histograms-options-1-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1-v0-o0.pdf}
   \caption{Distribution of coefficients in Model~1.
@@
   \centering
-  \includegraphics[width=\textwidth]{plot_model_summary_Z-options-1-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1-v0-o0.pdf}
   \caption{Frequency domain transfer functions for the coefficients in Models 1, 2, and 4.
@@
   \centering
-  \includegraphics[width=\textwidth]{plot_model_summary_Phi-options-1-v1-o0.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1-v1-o0.pdf}
   \caption{Frequency domain phase values for the coefficients in Models 1, 2, and 4.
@@
   \centering
-  \includegraphics[width=\textwidth]{plot_model_summary_Z_MT-options-1-v0-o0.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_Z_MT-options-1-v0-o0.pdf}
   \caption{Frequency domain transfer functions for the $\mathbf{Z}$ coefficients used in Model~3.}
@@
   \centering
-  \includegraphics[width=\textwidth]{plot_model_summary_Phi_MT-options-1-v1-o0.pdf}
+  \includegraphics[width=\textwidth]{figures/plot_model_summary_Phi_MT-options-1-v1-o0.pdf}
   \caption{Frequency domain phase values for the $\mathbf{Z}$ coefficients used in Model~3.}
